### 2021-11-19 SLOSS.Rmd
First commit

---
title: "Single large or Several Small SLOSS "
author: "Leonardo A. Saravia"
date: "9/3/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

needed_packages <- c(
    "tidyverse"
  , "lubridate"
  , "nlrx")

lapply(needed_packages, function(x) { if(!require(x,character.only = TRUE)) install.packages(x)} )

theme_set(theme_bw())
source("R/functions.r")

#### make changes
# Unix default NetLogo installation path (adjust to your needs!):
netlogopath <- file.path("/home/leonardo/NetLogo")
simfolder <- "/home/leonardo/Dropbox/Projects/ariadnaNL"
modelpath <- file.path(simfolder, "ariadnaNL.nlogo")
outpath <- file.path(simfolder,"Simulations")

# If not defined set the JAVA version of your local 
if(Sys.getenv("JAVA_HOME")==""){
  Sys.setenv(JAVA_HOME = "/usr/lib/jvm/java-11-openjdk-amd64")
  ## "/usr/lib/jvm/java-8-oracle"
}

nl <- nl(nlversion = "6.1.1",
         nlpath = netlogopath,
         modelpath = modelpath,
         jvmmem = 2048)

```

## Why do several small patches hold more species that few large patches?

## Simulaciones para ajuste usando Latin Hypercubic Sampling

* Incursion by common, generalist species. 
* Minimun patch size selective extinction and nestedness
* Time since patch creation, accumulation of species losses
* Habitat heterogeneity environmental clumping patchy species distribution
* Dominance of colonization/inmigration
* Population variability spreading the risk

## Lambda = 4

* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )

```{r SLOSS_lambda4_simulations}
require(tidyverse)

source("R/functions.r")

# Neutral communities 
#
mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_NoSelection_pf06_hpf3-59_dd1-3.csv")
sloss_sim <- mdl

names(mdl)
#
# Plot for neutral communities 
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

require(plotly)
plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

# Plot for hierarchical communities 
#
mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_Hierarchical_pf06_hpf3-59_dd1-3.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

#
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

# Plot for BirthSelection communities 
#
mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_BirthSelection_pf06_hpf3-59_dd1-3.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

#
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

# check how many simulations 
mdl1 <-sloss_sim %>% count(habitat_patch_size,birds_dispersal_distance,birds_behavior)

```

## Lambda = 2

* For patch sizes 3 9 without connectivity between patches (dispersal 1.1 - 1.2 ) the patch populations get extinct 
* Here I added another field "Migration0_at_deforestation"

```{r SLOSS_lambda4_simulations}
require(tidyverse)

source("R/functions.r")

# Neutral communities 
#
# No migration after deforestation
mdl <- read_netlogo_simul("Simulations/NoSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

## With migration
#
mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_NoSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

names(mdl)
#
# Plot for neutral communities 
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

# Plot for hierarchical communities 
#
mdl <- read_netlogo_simul("Simulations/Hierarchical_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_Hierarchical_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

#
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

# Plot for BirthSelection communities 
#

mdl <- read_netlogo_simul("Simulations/BirthSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

## with migration

mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_BirthSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)

# Add Migration0_at_deforestation=FALSE to previous simulations 
#
unique(sloss_sim$Migration0_at_deforestation)
sloss_sim <- sloss_sim %>% mutate( Migration0_at_deforestation = if_else(is.na(Migration0_at_deforestation), FALSE, Migration0_at_deforestation ))

#
#
mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>%
  mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()

```

## Not fragmented communities Lambda = 1.7 - 4

* 100x100 world

```{r SLOSS_notfrag_simulations}
require(tidyverse)

source("R/functions.r")

# Not fragmented communities = habitat_patch_size = 100
#
mdl <- read_netlogo_simul("Simulations/NoHiBi_pf00_hpf3_dd1-3_lambda1.7-4.csv") %>% mutate(habitat_patch_size=100)

# New Simulations 
#
mdl <- read_netlogo_simul("Simulations/NoHiBi_pf00_hpf3_dd1-3_lambda1.1-4.csv") %>% mutate(habitat_patch_size=100)

sloss_sim <- bind_rows(sloss_sim,mdl)

# Add Migration0_at_deforestation=FALSE to previous simulations 
#
unique(sloss_sim$Migration0_at_deforestation)
sloss_sim <- sloss_sim %>% mutate( Migration0_at_deforestation = if_else(is.na(Migration0_at_deforestation), FALSE, Migration0_at_deforestation )) 
unique(sloss_sim$habitat_patch_size)
#sloss_sim <- sloss_sim %>% mutate( habitat_patch_size = if_else(habitat_patch_size==0, 100, habitat_patch_size ))

saveRDS(sloss_sim, "Simulations/sloss_sim.rds")

names(mdl)
#
# Plot for neutral communities 
# I filter birth_rate_birds 1.1 because these communities survive only by migration
#
mdl1 <- mdl %>% filter(birth_rate_birds!=1.1) %>% group_by(birds_dispersal_distance, birth_rate_birds,birds_behavior) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=birth_rate_birds)) + geom_point() + geom_line() + scale_color_viridis_d(name="lambda") + facet_wrap( ~birds_behavior )  + theme_bw() + ylab("Number of species") + xlab("Dispersal Distance") 
ggsave("Figures/Species_NoFrag_Dispersal_Lambda.png",width=8,height=6,units="in",dpi=600)

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=birth_rate_birds)) + geom_point() + scale_color_viridis_d() + facet_wrap( ~birds_behavior ) + theme_bw()

mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=count_birds,color=birth_rate_birds)) + geom_point() + geom_line() + scale_color_viridis_d(name="lambda") + facet_wrap( ~birds_behavior )  + theme_bw() + ylab("Number of indiv.") + xlab("Dispersal Distance") 
ggsave("Figures/Indiv_NoFrag_Dispersal_Lambda.png",width=8,height=6,units="in",dpi=600)

mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>%  mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=birth_rate_birds,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()+ facet_wrap( ~birds_behavior )

```

## Add more simulations lambda = 2 with Migration

```{r SLOSS_combined_plots}

mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda2_Migrat.csv")

sloss_sim <- bind_rows(sloss_sim,mdl)

mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda2_NoMigrat.csv") 
sloss_sim <- bind_rows(sloss_sim,mdl)

mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda4_Migrat.csv") 
sloss_sim <- bind_rows(sloss_sim,mdl)

save.image()

#
# 200 x 200 - Started simulation 10:47
#
mdl <- read_netlogo_simul("Simulations/NoHiBi_200_pf06_hpf3-61_dd1-3_lambda2_NoMigrat.csv")
sloss_sim <- bind_rows(sloss_sim,mdl)
sloss_sim <- sloss_sim %>% mutate( world_width=if_else(is.na(world_width),100,world_width))
sloss_sim <- sloss_sim %>% mutate( world_width=if_else(world_width==99,100,world_width))

saveRDS(sloss_sim, "Simulations/sloss_sim.rds")

```

## Plots of combined behaviors and patch 100

```{r SLOSS_combined_plots}
require(tidyverse)

source("R/functions.r")

#mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation | (birth_rate_birds==2 & habitat_patch_size==100)) %>%  
sloss_sim %>% distinct(Migration0_at_deforestation, world_width)

#
# Count number of simulations
#
mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation, world_width==100 ) %>%  
  count(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior)

mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation, world_width==100 ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

#mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3 || birds_dispersal_distance == 1.5) %>%  
mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")

ggsave("Figures/Species_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_shannon_diversity,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Shannon diversity") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")
ggsave("Figures/Shannon_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)

#
# With migration all the time
#

mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation ) %>%  
  count(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior)

#
# Plot without SD
#
mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birth_rate_birds == 2, world_width==100 ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% Migration")

ggsave("Figures/Species_PatchSize_Dispersal_Migration.png",width=8,height=6,units="in",dpi=600)

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_shannon_diversity,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Shannon diversity") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% Migration")
ggsave("Figures/Shannon_PatchSize_Dispersal_Migration.png",width=8,height=6,units="in",dpi=600)

#
# From Lamda 2 to lambda 4 Neutral
#

mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birds_behavior == "NoSelection" ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birth_rate_birds ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Neutral - Degraded 60% - Migration 0.001")
ggsave("Figures/Species_lambda2-4_PatchSize_Dispersal_NoSelection.png",width=8,height=6,units="in",dpi=600)

#
# From Lamda 2 to lambda 4 Hierarchical
#

mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birds_behavior == "Hierarchical" ) %>%  
  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds) %>%
  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )

mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birth_rate_birds ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Hierarchical - Degraded 60% - Migration 0.001")
ggsave("Figures/Species_lambda2-4_PatchSize_Dispersal_Hierarchical.png",width=8,height=6,units="in",dpi=600)

```

### 2021-11-19 R/functions.r

@@ -1,0 +1,74 @@
+
+
+
+
+#' Binomal smooth function for ggplot 
+#'
+#' @param ... 
+#'
+#' @return
+#' @export
+#'
+#' @examples
+binomial_smooth <- function(...) {
+  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...,linetype=2,size=0.5,se=FALSE)
+}
+
+#' Read Netlogo simualtions and clean field names 
+#'
+#' @param fname name of the csv file with NetLogo simulations
+#' @param skip if the simulation is done with Behavior Space tool it has to skip the 6 first lines
+#'
+#' @return a data.frame 
+#' @export
+#'
+#' @examples
+read_netlogo_simul <- function(fname,skip=6){
+  require(readr)
+  mdl <- read_csv(fname,skip=skip)
+  
+  nam <- names(mdl)
+  nam <- gsub("-","_",nam)
+  nam <- gsub("[][]","",nam)
+  nam <- gsub(" ","_",nam)
+  names(mdl) <- nam
+  return(mdl)
+}
+
+#' Estimate the critical point of species living inside and habitat (birds inside forest) 
+#'
+#' @param mdl 
+#'
+#' @return
+#' @export
+#'
+#' @examples
+estim_crit_point_size <- function(mdl,size){
+  if( "world_width" %in% names(mdl) ){
+    if( unique(mdl$world_width)!= size ) {
+      stop( paste("Size different from world view",unique(mdl$world_width)))
+    } 
+  }
+  #
+  # Estimate the probability of birds survival
+  #
+  mdlp <- mdl %>% group_by(birth_rate_birds,birth_rate_forest) %>% mutate(birds_survive=(step==1000)) %>%
+    summarize(n=n(),birds_survive_prob=sum(birds_survive/n)) 
+  
+  # Estimate from which forest_birth_rate we have enough data to calculate the birdsfores_
+  #
+  mm <- mdlp %>% group_by(birth_rate_forest) %>% summarize(ngt0 = sum(birds_survive_prob>0)) %>% filter(ngt0>=10) %>% summarize(min_br=min(birth_rate_forest),max_br=max(birth_rate_forest))
+  
+  mdlp2 <- mdlp %>% filter(birth_rate_forest>=mm$min_br) %>% group_by(birth_rate_forest) %>% do( 
+    {
[…]

### 2022-11-14 RunSims.Rmd
Add new Rmd with simulation runs

--- 2021-11-19 SLOSS.Rmd
@@
 ---
-title: "Single large or Several Small SLOSS "
+title: "Running simulations of MultipleContactNeutralMigrat.nlogo"
 author: "Leonardo A. Saravia"
-date: "9/3/2019"
+date: "8/11/2022"
 output: html_document
@@
   , "lubridate"
+  , "future"
+  , "tictoc"
   , "nlrx")
@@
 
-#### make changes
+#
+# Where the NetLogo model is
+# Setup for nlrx 
+#
+if( Sys.info()['nodename'] =="ls-pro") {
+  simfolder <- "/home/leonardo/Dropbox/Projects/DynamicForestExtinction"
+
+} else if(Sys.info()['nodename'] =="biologia2018") {                     # Server UNGS
+  simfolder <- "/home/leonardo/Dropbox/Projects/DynamicForestExtinction"
+
+}
+
 # Unix default NetLogo installation path (adjust to your needs!):
+#
 netlogopath <- file.path("/home/leonardo/NetLogo")
-simfolder <- "/home/leonardo/Dropbox/Projects/ariadnaNL"
-modelpath <- file.path(simfolder, "ariadnaNL.nlogo")
+modelpath <- file.path(simfolder, "MultipleContactNeutralMigrat.nlogo")
 outpath <- file.path(simfolder,"Simulations")
@@
 
-nl <- nl(nlversion = "6.1.1",
+nl <- nl(nlversion = "6.2",
          nlpath = netlogopath,
@@
 
-## Simulaciones para ajuste usando Latin Hypercubic Sampling
-
-* Incursion by common, generalist species. 
-* Minimun patch size selective extinction and nestedness
-* Time since patch creation, accumulation of species losses
-* Habitat heterogeneity environmental clumping patchy species distribution
-* Dominance of colonization/inmigration
-* Population variability spreading the risk
-
 ## Lambda = 4
@@
 
-```{r SLOSS_lambda4_simulations}
-require(tidyverse)
+```{r SLOSS_lambda4_simulations_lambda4,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=TRUE}
 
-source("R/functions.r")
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+#
+nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "birth-rate-birds" = 4,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.001,
+                                             "Video" = "false",
+                                             "prob-frag" =  0.6,
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"NoSelection\"",
+                                             "replacement-rate" = 0.0
+                                             ))
+
+#
+# Run 10 times this do not set the run variable for output 
+#
+# nseeds * number_of_parameters / split = 10 * 300 / 4 - total number of jobs = nseeds * split = 10*4
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=10)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 4)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+```{r readSimulations, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_ff.csv"),skip=0) %>% 
+  mutate(mean_free_path=str_sub(mean_free_path,2,-2)) %>% 
+  separate(mean_free_path, c("non_degraded_mfp","degraded_mfp"), sep=" ",convert=TRUE)
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+## Old code to plot
 
@@
 #
-mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_NoSelection_pf06_hpf3-59_dd1-3.csv")
-sloss_sim <- mdl
-
 names(mdl)
@@
 
-require(plotly)
-plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))
+# require(plotly)
+# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))
 
@@
 
-```{r SLOSS_lambda4_simulations}
+```{r SLOSS_lambda4_simulations_lambda2,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 require(tidyverse)
@@
 
-* 100x100 world
-
-```{r SLOSS_notfrag_simulations}
+```{r SLOSS_notfrag_simulations, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 require(tidyverse)
@@
 
-```{r SLOSS_combined_plots}
+```{r SLOSS_combined_plots, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 
@@
 
-```{r SLOSS_combined_plots}
+```{r SLOSS_combined_plots1, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 require(tidyverse)

### 2023-05-09 RunSims.Rmd
Add manuscript.md

@@
 modelpath <- file.path(simfolder, "MultipleContactNeutralMigrat.nlogo")
-outpath <- file.path(simfolder,"Simulations")
+outpath <- file.path("Simulations")
 
@@
 
-## Lambda = 4
+## Lambda = 1-4
 
 * Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 ) 
+* Check simulations with Migration0-at-deforestation true or false to take into account that immigration happens at the borders and different kinds of deforestation patterns have different border size e.g.
+regular > randon block
 
-```{r SLOSS_lambda4_simulations_lambda4,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=TRUE}
+```{r SLOSS_lambda14_simulations,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=TRUE}
 
 # pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
-#
-nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3",
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3_brb1-4",
                             outpath=outpath,
@@
                             variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
-                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61))
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
                                              ),
@@
                                              "max-birds-species" = 100,
-                                             "birth-rate-birds" = 4,
                                              "death-rate-birds" = 1,
-                                             "migration-rate-birds" = 0.001,
+                                             "migration-rate-birds" = 0.0001,
                                              "Video" = "false",
@@
 #
-# Run 10 times this do not set the run variable for output 
+# Run nseeds=10 times 
 #
-# nseeds * number_of_parameters / split = 10 * 300 / 4 - total number of jobs = nseeds * split = 10*4
+# Total number of jobs = nseeds * split = 10*4 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  300 / 4 
 #
@@
 
+## Read Simulations
+
 ```{r readSimulations, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
@@
 sloss_sim <- bind_rows(sloss_sim,mdl)
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
 
-## Old code to plot
+```
+
+## Exploratory Plots
+
+```{r ExploratoryPlots, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 

### 2023-08-05 RunSims.Rmd
Updates simulattions and manuscrito

@@
 
-nl <- nl(nlversion = "6.2",
+nl <- nl(nlversion = "6.3",
          nlpath = netlogopath,
@@
 
-## Lambda = 1-4
+## Lambda = 1-4 NoSelection = Neutral  Regular
 
@@
 
-```{r SLOSS_lambda14_simulations,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=TRUE}
+```{r SLOSS_lambda14_sims_regular,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 
@@
 # brb = bird rate birds 
-nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3_brb1-4",
+nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3_brb1-4_regular",
                             outpath=outpath,
@@
                             runtime=0,
-                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path"),
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
                             variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
@@
 #
-# Total number of jobs = nseeds * split = 10*4 
-# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  300 / 4 
+# Total number of jobs = nseeds * split = 20*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
 #
@@
 nl@simdesign <- simdesign_ff(nl=nl,
-                               nseeds=10)
+                               nseeds=30)
 
@@
 tic()
-results <- run_nl_all(nl,split = 4)
+results <- run_nl_all(nl,split = 2)
 #results <- run_nl_all(nl,split = 4)
@@
 # Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+# Calculate cluster stats
 #
-mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_ff.csv"),skip=0) %>% 
-  mutate(mean_free_path=str_sub(mean_free_path,2,-2)) %>% 
-  separate(mean_free_path, c("non_degraded_mfp","degraded_mfp"), sep=" ",convert=TRUE)
+# Old 10 - 20 simulations 
+# mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_ff.csv"),skip=0) %>% 
+# mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_20_ff.csv"),skip=0) %>%
+
+mdl <- read_sloss_simul(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")
 
 sloss_sim <- bind_rows(sloss_sim,mdl)
+
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
+
+netlogo_evaluate_patch_distr(mdl$find_cluster_sizes[1])
+
+```
+
+## Plots - No selection - regular pattern
+
+```{r PlotsNoSelectionRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+
+# Neutral communities 
+#
+mdl <- sloss_sim %>% filter(deforestation_type == "regular",birds_behavior=="NoSelection")
+names(mdl)
+#
+# Plot for neutral communities
+# Calculate the power exponent 
+# pexp = (1 - 2 * birds_dispersal_distance ) / (1 - birds_dispersal_distance ), median_dispersal_distance = 2^(1/pexp-1) 
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
+
+names(mdl1)
+
+mdl1 <- mdl1 %>%  mutate(across(max_patch_mean:max_patch_qh,sqrt))
+
+#
+# Block patch size vs max patch size
+#
+mdl1 %>%  
+  ggplot(aes(x = max_patch_mean, y = habitat_patch_size, color = factor(habitat_patch_size))) + 
+  geom_point(  size = 1.2) + guides(color = "none") #+
+
+# require(plotly)
+# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))
+
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/Species_Habitat_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+
+#
+# Number of species by max_patch size 
+# 
+mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
+  ggplot(aes(x = max_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Max patch size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/Species_MaxPatch_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+
+#
+# shannon_diversity by max_patch size 
+# 
+mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
+  ggplot(aes(x = max_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Max patch size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = c(0.9,0.2)
+    )
+
+ggsave("Figures/Shannon_MaxPatch_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+
+```
+
+## Lambda = 1-4 / Random Block patterns
+
+* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )
+
+```{r SLOSS_lambda14_simulations_block,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3_brb1-4_block",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path", "find-cluster-sizes"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "prob-frag" =  0.6,
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"random block\"",
+                                             "birds-behavior" = "\"NoSelection\"",
+                                             "replacement-rate" = 0.0
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Read Simulations No Selection Random Block Patterns
+
+```{r readSimulationsBlock, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_sloss_simul(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_block_ff.csv")
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
 saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
@@
 
-## Exploratory Plots
+## Plots Neutral No Selection Random Block Patterns Plots
 
-```{r ExploratoryPlots, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+```{r NeutralRandomBlockPlots, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 
@@
 #
+mdl <- sloss_sim %>% filter(deforestation_type == "random block", birds_behavior == "NoSelection")
 names(mdl)
@@
 #
-mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch) ,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/
+
+names(mdl1)
+mdl1 <- mdl1 %>%  mutate(across(max_patch_mean:max_patch_qh,sqrt))
+
+#
+# Block patch size vs max patch size
+#
+mdl1 %>%  
+  ggplot(aes(x = max_patch_mean, y = habitat_patch_size, color = factor(habitat_patch_size))) + 
+  geom_point(  size = 1.2) + guides(color = "none") #+
+
+#
+# Number of species by patch block size
+#
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/Species_Habitat_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+
+#
+# Number of species by max_patch size 
+# 
+mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Max patch size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.83)
+    )
+
+ggsave("Figures/Species_MaxPatch_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+
+#
+# shannon_diversity by max_patch size 
+# 
+mdl1  %>% # filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
+  ggplot(aes(x = max_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Max patch size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = c(0.9,0.2)
+    )
+
+ggsave("Figures/Shannon_MaxPatch_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+
+```
+
+## Simulations Lambda = 1-4 / Regular patterns - Hierarchical communities
+
+* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )
+
+```{r Hierarchical_SLOSS_lambda14_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="Hierarchical_pf06_hps3-61_dd1-3_brb1-4_regular",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "prob-frag" =  0.6,
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"Hierarchical\"",
+                                             "replacement-rate" = 0.3
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Read Simulations hierarchical regular Patterns
+
+```{r readSimulationsHierarchicalRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_sloss_simul(outpath,"Hierarchical_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
+
+```
+
+# Plot for hierarchical communities
+
+```{r ExploratoryPlotsHierarchical, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+
+mdl <- sloss_sim %>% filter(deforestation_type == "regular", birds_behavior=="Hierarchical")
+names(mdl)
+#
+# summarize data 
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(birds_at_deforestation:count_birds,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+names(mdl1)
 
@@
 
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
+mdl1 %>% 
+  ggplot(aes(x=habitat_patch_size, y=calc_number_of_species_mean, color=birds_dispersal_distance)) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds))) +
+  geom_ribbon(
+    aes(
+      ymin = calc_number_of_species_ql,
+      ymax = calc_number_of_species_qh,
+      fill = birds_dispersal_distance
+    ),
+    alpha = 0.3,
+    colour = NA,
+    show.legend = FALSE
 
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
+  ) +
+  scale_color_viridis_c(guide = FALSE) + 
+  scale_fill_viridis_c() +
+  facet_grid(birds_dispersal_distance ~ birth_rate_birds) + 
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species")
 
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
 
-mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()
+ggsave("Figures/Hierarchical_Species_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
 
-# Plot for hierarchical communities 
+# Shannon-Diversity  
 #
-mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_Hierarchical_pf06_hpf3-59_dd1-3.csv")
-sloss_sim <- bind_rows(sloss_sim,mdl)
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
 
-#
-#
-mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()
-
-# Plot for BirthSelection communities 
-#
-mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_BirthSelection_pf06_hpf3-59_dd1-3.csv")
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-#
-#
-mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()
-
-# check how many simulations 
-mdl1 <-sloss_sim %>% count(habitat_patch_size,birds_dispersal_distance,birds_behavior)
+ggsave("Figures/Hierarchical_Shannon_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 
-## Lambda = 2
+## Simulations Lambda = 1-4 / Regular patterns - Birth Selection communities
 
-* For patch sizes 3 9 without connectivity between patches (dispersal 1.1 - 1.2 ) the patch populations get extinct 
-* Here I added another field "Migration0_at_deforestation"
+* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )
 
-```{r SLOSS_lambda4_simulations_lambda2,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
-require(tidyverse)
+```{r Selection_SLOSS_lambda14_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=TRUE}
 
-source("R/functions.r")
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="Selection_pf06_hps3-61_dd1-3_brb1-4_regular",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "prob-frag" =  0.6,
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"BirthSelection\"",
+                                             "replacement-rate" = 0.3
+                                             ))
 
-# Neutral communities 
 #
-# No migration after deforestation
-mdl <- read_netlogo_simul("Simulations/NoSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Read Simulations Birth Selection  regular Patterns
+
+```{r readSimulationsSelectionRegular, echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_sloss_simul(outpath,"Selection_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")
+
 sloss_sim <- bind_rows(sloss_sim,mdl)
 
-## With migration
-#
-mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_NoSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-names(mdl)
-#
-# Plot for neutral communities 
-#
-mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()
-
-# Plot for hierarchical communities 
-#
-mdl <- read_netlogo_simul("Simulations/Hierarchical_pf06_hpf3-59_dd1-3_lambda2.csv")
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_Hierarchical_pf06_hpf3-59_dd1-3_lambda2.csv")
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-#
-#
-mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()
-
-# Plot for BirthSelection communities 
-#
-
-mdl <- read_netlogo_simul("Simulations/BirthSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-## with migration
-
-mdl <- read_netlogo_simul("Simulations/MultipleContactNeutralMigrat_BirthSelection_pf06_hpf3-59_dd1-3_lambda2.csv")
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-# Add Migration0_at_deforestation=FALSE to previous simulations 
-#
-unique(sloss_sim$Migration0_at_deforestation)
-sloss_sim <- sloss_sim %>% mutate( Migration0_at_deforestation = if_else(is.na(Migration0_at_deforestation), FALSE, Migration0_at_deforestation ))
-
-#
-#
-mdl1 <- mdl %>% group_by(birds_dispersal_distance, habitat_patch_size) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=deltaPop,color=habitat_patch_size)) + geom_point() + scale_color_viridis_c(guide = FALSE) + facet_wrap( ~habitat_patch_size ) + theme_bw()
-
-mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>%
-  mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
 

### 2023-08-05 ms_SLOSS.Rmd
Updates simulattions and manuscrito

---
title: "Why there are more species in several small patches vs few large patches:
  a multi-species modeling approach"
bibliography: sloss_bib.bib
csl: "ecology-letters.csl"
output: pdf_document
---

Leonardo A. Saravia ^1^, Ulises Balza ^1^, Fernando Momo^2^

1. Centro Austral de Investigaciones Científicas (CADIC-CONICET). Houssay 200. Ushuaia, Argentina

2. Instituto de Ciencias. Universidad Nacional de General Sarmiento.	J.M. Gutierrez 1159 (1613), Los Polvorines.	Buenos Aires, Argentina.

## Abstract

As global biodiversity is rapidly declining due to habitat loss, it is important to determine how to protect it. Still, the policies of many conservation agencies are prioritizing the preservation of a single large habitat area (SL) vs several small areas of the same size (SS), besides empirical support favoring SS. However, to date, most of the theories regarding this issue were based on single-species models. The aim of this work is to explore spatial multi-species models to verify the SS > SL pattern and determine the causes of it. We used three types of models: a classical neutral, neutral with habitat selection, and a competitive hierarchical model.  All models have three parameters, extinction, reproduction, and dispersal distance. In the classical neutral model, species are equivalent and have the same parameters, when there is a reproduction event they send the propagule to a surrounding patch blindly, if the patch is already occupied the propagule dies. In the neutral-with-habitat selection model, the propagules select an empty patch so their survival is ensured. In the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant with a replacement rate parameter. We use 100 species and start simulating the colonization of an empty area then we destroy a fixed amount of habitat using different regular patch sizes. The results for all models are almost always that SS retains more species than SL. The extinction of species in patches depends on the quotient between reproduction and extinction rates in relation to patch size. The only case when SL > SS is when dispersal is not enough to connect patches, and patch size is smaller than the minimum to prevent extinction. If patch size is above the critical size to maintain local populations the SS > SL pattern is due to the sampling effect of aggregated species distribution combined with the reduction of dispersal ability of species. High dispersal rates produce homogenization of species distribution and a reduction of species richness and this is why habitat fragmentation by habitat loss retains more species when patches are smaller. A prediction of the model is that if a set of species have more dispersal capabilities the SS effect would be lower than for species with less dispersal.

## Keywords

## Introduction

Global biodiversity is rapidly declining due to habitat loss (quotes). Moreover, habitat fragmentation occurs at the same time with habitat loss, and the effects of both processes are difficult to disentangle. Landscape connectivity and organism mobility as a critical feature for population persistence [@Taylor1993]. Starting from a landmark paper [@Diamond1975], conservation biologists engaged in the challenging task of recommending protection measures given the environmental crisis coupled with lack of knowledge of most of the world’s biodiversity. Early recommendations claimed to follow ecological theory, in particular island biogeography theory to derive simple thumb rules decisions in order to better design protected areas. However, as noted rapidly, species/area relationships do not have a clear association with fragmentation per se but only with overall habitat loss [@Simberloff1982].
Aiming to design better management practices, this has led to a nearly 50-year Single Large vs Several Small (SLOSS) debate. While some authors argue that habitat loss is the most overwhelming effect on species loss, for others the complex factors arising from fragmentation further enhance species extinctions [@Fahrig2017;@Fletcher2018].

The SLOSS debate has been an important academic and practical discussion among conservation biology. In practice, a series of decision-maker targeted recommendations were proposed and remain rather uncritically in several conservation instances (quotes).
While the overall habitat loss is widely accepted as a process leading to long-term biodiversity loss (quotes), the propose that single large reserves retain more species that several small reserves of the same area (SL > SS) remain at a similar hierarchical state, even after several criticisms (reviewed in  @Fahrig2017). Moreover, empirical observational and experimental data shows the predominance of the opposite pattern (SS > SL) ever since the 80s and this fact has been reinforced to this day [@Simberloff1982; @Fahrig2020; @Hammill2020], add more. The actual processes driving to it remains elusive and it is probably multi-causal (quotes). In that scenario, model simulations could produce useful predictions targeting the relevant variables in order to better understand the processes leading to non-intuitive patterns.

The SLOSS debate has been approached from several theoretical frameworks, but most of them were based on single-species models. Moreover, spatial dimension in several works has been neglected as mere noise around long-term equilibrium dynamics (quotes, but see @Rybicki2013). In that context, spatial-explicit multi-species modeling could be meaningful as several organisms retain dispersal distance and other dispersal-related life traits far consistently among species (quotes).

The aim of this work is to explore spatial multi-species models to verify the SS > SL pattern and determine the causes of it. In particular, we simulate species retention after different fragmentation scenarios of the same overall habitat loss, therefore accounting for the interacting effects of habitat loss and fragmentation.

## Methods

### Model

The model represents a landscape occupied by multiple bird species that undergoes habitat degradation over time. The purpose is to explore the impact of fragmentation patterns and species interaction mechanisms on diversity and abundance distributions.

The environment comprises a 2D grid of patches that can be either suitable habitat or degraded. Bird agents have state variables for species identity, size, and color. Globals contain key parameters.

At initialization, an intact landscape is seeded with bird agents having randomized species identities. Habitat loss occurs probabilistically at a set time step based on three spatial patterns: random, regular fragmentation, or block removal.

During each time step, the following sequence occurs:

Habitat loss event if scheduled

Calculation of probabilities for reproduction, dispersal, migration

Birds undergo probabilistic dispersal and reproduction based on:

Neutral model: Offspring disperse locally based on power law dispersal kernel and settle in any vacant habitat
Hierarchical model: Species can displace lower ranked species during dispersal
Immigration allows spontaneous appearance of birds at habitat edges

Data collection on species diversity, abundance, and spatial distribution

The model implements stochasticity in multiple processes including dispersal distance, loss patterns, migration, and species selection. Emergent macro-level diversity and abundance patterns result from individual-level bird behaviors.

Key outputs analyzed are species occupancy, diversity indices, abundance distributions, and spatial patterns. Experiments examine the dynamics under varying habitat loss, dispersal, and competition scenarios.

* **MultipleContactNeutralMigrat.nlogo** A multiple species neutral/hierarchical model described in with 
	degraded sites that cannot be colonized

	<https://github.com/lsaravia/DynamicForestExtinction>

Regular habitat patterns: squares of a range of sizes are distributed regularly over the simulation grid. The side of the squares where 4,20,30,63 sites.

Random block habitat patterns: habitat blocks of the same size are randomly distributed over the simulation grid. The sizes are identical to the used for regular habitat simulations. This habitat simulation results in patches that have a minimum size equal to the block size but a maximum size that is bigger, so I characterize this with the maximum patch size.

## Results

![Relationship between number of species and habitat patch size for varying dispersal distances. Simulations were conducted using a model with 60% habitat destruction and regular habitat patterns, where all habitat patches are of equal size.](Figures/Species_Habitat_Dispersal_Lambda_regular.png)

![Relationship between Shannon Diversity and habitat patch size for varying dispersal distances. Simulations were conducted using a model with 60% habitat destruction and regular habitat patterns, where all habitat patches are of equal size.](Figures/Shannon_Habitat_Dispersal_Lambda_regular.png)

![Relationship between number of species and maximum patch size for varying dispersal distances. Simulations were conducted using a model with 60% habitat destruction and random block habitat patterns, where habitat blocks of the same size are randomly distributed over the simulation grid (i.e., the maximum patch size).](Figures/Species_MaxPatch_Dispersal_Lambda_block.png)

![Relationship between Shannon Diversity and maximum patch size for varying dispersal distances. Simulations were conducted using a model with 60% habitat destruction and random block habitat patterns, where habitat blocks of the same size are randomly distributed over the simulation grid (i.e., the maximum patch size).](Figures/Shannon_MaxPatch_Dispersal_Lambda_block.png)

## Discussion

Overall, our results confirm the previous found pattern and nearly all models retained more species under the SS rather than the SL scenario. In both cases, the loss of species is relevant when habitat loss takes place, and no significant extinction biased towards SS occurs as a long-term dynamic. The extinction of species in patches depends on the quotient between reproduction and extinction rates in relation to patch size, i.e., the ability of a certain patch to be a source of propagules. The only case when SL > SS is when dispersal is not enough to connect patches, and size is smaller than the minimum to prevent patch extinction, as predicted elsewhere [@Rybicki2013]. Letting aside those cases (i.e., not reaching minimum dispersal distance and intra-patch sink dynamics), our model was able to produce the patterns observed so far in field studies [@Fahrig2020]. In that sense, we can hypothesize that our model variables (i.e., dispersal ability, patch size and lambda) may be biologically meaningful and enough to explain why SS retains more species than SL. Importantly, the spatial arrangement of available patches was not necessary to produce the spatial autocorrelation of occupied patches known to be responsible at least in part of the patterns observed in nature [@Fahrig2021]. In contrast, high dispersal rates produce homogenization of species distribution and a reduction of species richness, especially if high dispersal species are also good competitors. In that scenario, SS and SL are approximately equivalent for species diversity retention. A prediction of the model is that if a set of species have more dispersal capabilities the SS effect would be smaller than for species with less dispersal abilities.

Our model further illustrate that available patches are not always re-occupied following fragmentation even if the minimum size is met, therefore reinforcing the fact that dispersal is the main driver of the pattern, and that chance is involved in the final species pool when fragmentation takes place. Empirical data in large carnivores, for instance, support that several available patches are not occupied after isolation [@Paviolo2016].

@Rybicki2013 suggested that both short-term and long-term extinctions are produced after fragmentation. Here, we observed that, in most cases, long-term loss of species in SS scenarios were less important than short-term extinctions in SL scenario.
Overall, our results support the main conclusion by @Fahrig2021, in that the conditions for SL > SS are rather restricted and would not be observable in nature frequently (see @Fahrig2021, Fig 3). In contrast, if few conditions are met (minimum patch size and dispersal), SS would retain more species than SL for nearly all situations. Evidence suggests that small patches, even single structures such as isolated trees, can be important for biodiversity and landscape connectivity [@Shafer1995; @Manning2006]. While connectivity management is case-specific, our results further support past claims [@Fahrig2020;@Fahrig2021], that assuming that SL is always better do not have empirical support. In our work, spreading of risk was not modelled (__maybe as time lag in at-patch competition?__), therefore reinforcing that the sampling effect is enough to produce a SS > SL pattern if dispersal is small enough to produce spatial autocorrelation of occupied patches [@Simberloff1976].

One large patch can enhance competitive exclusion, and that is more important the more similar species are. Hummingbirds?

### Limitations

We acknowledge the limitations of our study, and consider it a strength that we have identified them in this initial model. This allows us to continue working on them in future research.

Matrix is assumed as non-habitat

Genetic structured not considered

We are not modeling total diversity, rather diversity of a certain group with a set of similar species

### Conclusions

## Acknowledgments

## References

<!--

[a]: [UB1]Me parece que está bien el foco en esto, así no nos critican con boludeces como “ahh pero no tiene en cuenta que el pato en peligro puede en realidad necesitar…” tenemos que tratar de explicar por qué Fahrig encuentra lo que encuentra

ej Fletcher (2018): “Approximately one quarter (24.4%) of Fahrig's results focused on species richness, with little consideration of species identity. Species richness is useful for summarizing ecological patterns but can mask compositional changes that are highly relevant to conservation”

“the review had a narrow focus on a habitat vs non-habitat dichotomy, assuming the functionally-relevant habitat was appro[1]priately quantified and of comparatively similar quality across the landscape for the responses considered. For many species, the non-ha[1]bitat matrix may also provide resources and generic habitat delineation can obscure variation in habitat fragmentation effects”

[b]: Si explicitamos esas limitaciones, estamos cubiertos me parece quiero decir que somos conscientes de las limitaciones, y es una fortaleza que las hayamos identificado en este primer modelo para seguir trabajando luego en el futuro

-->

### 2023-08-05 R/functions.r

@@ -35,6 +35,14 @@ read_netlogo_simul <- function(fname,skip=6){
   return(mdl)
 }
 
+
+read_sloss_simul <- function(outpath,fname){
+  mdl <- read_netlogo_simul(file.path(outpath,fname),skip=0) %>% 
+  mutate(mean_free_path=str_sub(mean_free_path,2,-2)) %>% 
+  separate(mean_free_path, c("non_degraded_mfp","degraded_mfp"), sep=" ",convert=TRUE) %>% rowwise() %>%
+  mutate(netlogo_evaluate_patch_distr(find_cluster_sizes))
+}
+
 #' Estimate the critical point of species living inside and habitat (birds inside forest) 
 #'
 #' @param mdl 
@@ -72,3 +80,21 @@ estim_crit_point_size <- function(mdl,size){
   
   return(mdlp2)  
 }
+
+
+#' Evaluate patch distribution from netlogo list of patch sizes output "[]"
+#'
+#' @param br string with a netlogo list with [] and separated by blanks
+#'
+#' @return a data frame with results
+#' @export
+#'
+#' @examples
+netlogo_evaluate_patch_distr <- function(br) {
+  
+  pp <-  str_remove_all(br, "[\\[\\]]") %>% str_split(" ") %>% unlist() %>% as.numeric() 
+  # message(br)
+  tibble(max_patch=max(pp),tot_patch=sum(pp),num_patch=length(pp),mean_patch=mean(pp),median_patch=median(pp))
+
+}
+

### 2023-08-22 BESpresentation.Rmd
Update manuscript

---
title: "Why there are more species in several small patches vs few large patches: a multi-species modeling approach"
subtitle: ""
author:
- Leonardo A. Saravia, [Universidad Nacional De General Sarmiento]
date: "2021/07/06"
output:
  xaringan::moon_reader:
    css:
      - default-fonts
      - css/leo.css
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: true
      countdown: 60000
---

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble()
#xaringanExtra::use_webcam()
```
class:  center, middle

# SLOSS

## is it better to have Several Small or Single Large habitat for conservation?

---

background-image: url(Figures/Fletcher2019.png)
background-size: contain

## On one side

---

background-image: url(Figures/Fahrig2020.png)
background-size: contain

## On the other

---
background-image: url(Figures/Saravia2018.png)
background-size: contain

## A Neutral - hierarchical model

---

background-image: url(Figures/NeutralHier00.jpeg)
background-size: contain

## The Neutral side

---

background-image: url(Figures/NeutralHier01.jpeg)
background-size: contain

## The Neutral side

---
background-image: url(Figures/NeutralHier02.jpeg)
background-size: contain

## The Neutral side

---
background-image: url(Figures/NeutralHier03.jpeg)
background-size: contain

## The Hierarchical side

---
background-image: url(Figures/NeutralHier04.jpeg)
background-size: contain

## An additional Neutral flavor

---

## What happens without habitat loss

.center[
<img src="Figures/Species_NoFrag_Dispersal_Lambda.png" width="90%">
]

???

The general pattern is that with more dispersal there are less species 
for smaller lambda (1.5 is near the extinction threshold) the increase in
dispersal produces and increase in survival

So with more dispersal there is a bigger probabilty for one species to dominate

---

class:  center, middle

## A Fragmentation experiment

<video width="500" height="500">
<source src="NeutralNoSel_200x200_lambda2_disp3_patch30_noMigrat.mp4" type="video/mp4">
</video>

---
## Fragmentation - Number of species

.center[
<img src="Figures/Species_PatchSize_Dispersal_Migration.png" width="90%">
]

---
class:  center, middle

## The minimum patch size

<video width="500" height="500">
<source src="NeutralNoSel_200x200_lambda2_disp3_patch3_noMigrat.mp4" type="video/mp4">
</video>

---

## Fragmentation - Shannon Diversity

.center[
<img src="Figures/Shannon_PatchSize_Dispersal_Migration.png" width="90%">
]

---

## Model results agree with Fahrig

--
### There is a minimum patch size that depends on lambda

--
### After that threshold fragmentation increase diversity

--
# why?

--

## Because fragmentation acts as a barrier to dispersal

---

## Model predictions

--
### For communities with a smaller average dispersal distance framgmentation will have strong positive effect

--
### For communities with a bigger average dispersal distance fragmentation will have no effect

--
### To my knowledge there is no experiment relating fragmentation to dispersal distance.

???

Prediction can be compared with experiments

---
## How to translate this to complex patch configurations

### The mean free habitat path (HP)

The mean distance from a random place inside the habitat to reach the degraded area

### The mean free degraded path (DP)

The mean distance from a random place inside the degraded area to reach the habitat

### HP and DP in relation to mean dispersal distance (DD)

???

These fragmentation patterns are delivertely simple so in real cases we need to have more operational 
parameters

---
background-image: url(Figures/FreePathDispersal01.jpeg)
background-size: contain

## A connected landscape - lower no. species

---
background-image: url(Figures/FreePathDispersal02.jpeg)
background-size: contain

## A less connected landscape - Higher no. species

---
background-image: url(Figures/FreePathDispersal03.jpeg)
background-size: contain

## The ideal fragmented landscape

---

class: center, middle

# THE END

### 2023-08-22 RunSims.Rmd
Update manuscript

@@
 #
-mdl <- sloss_sim %>% filter(deforestation_type == "regular",birds_behavior=="NoSelection")
+mdl <- sloss_sim %>% filter(deforestation_type == "regular",birds_behavior=="NoSelection", habitat_patch_size != 199)
 names(mdl)
@@
 
-```{r Selection_SLOSS_lambda14_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=TRUE}
+```{r Selection_SLOSS_lambda14_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
 
@@
 
-## Not fragmented communities Lambda = 1.7 - 4
+# Plot for birth selection communities regular patterns
+
+```{r ExploratoryPlotsSelectionRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+
+mdl <- sloss_sim %>% filter(deforestation_type == "regular", birds_behavior=="BirthSelection")
+names(mdl)
+#
+# summarize data 
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(birds_at_deforestation:count_birds,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+names(mdl1)
+
+# require(plotly)
+# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))
+
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/Selection_Species_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/Selection_Shannon_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+
+```
+
+## Plots across neutral hierarchical selection modelos regular fragmentation
+
+```{r PlotsAcrossBehaviorRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+
+mdl <- sloss_sim %>% filter(deforestation_type == "regular")
+names(mdl)
+#
+# summarize data 
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,birds_behavior) %>% summarise(across(birds_at_deforestation:count_birds,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+names(mdl1)
+
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birds_behavior)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birds_behavior), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birds_behavior, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/AcrossBehav_Species_Dispersal_Lambda2_regular.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birds_behavior)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birds_behavior), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() +
+  facet_wrap(~ birds_behavior, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/AcrossBehav_Shannon_Dispersal_Lambda2_regular.png",width=8,height=6,units="in",dpi=600)
+
+```
+
+## Not fragmented communities Lambda = 1.7 - 4 - No Selection communities
+
+* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )
+
+```{r NoSelection_SLOSS_lambda14_notfrag,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="NoSelection_pf00_nofrag_dd1-3_brb1-4_regular",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "prob-frag" =  0.6,
+                                             "Deforestation-at-time" = 0,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"NoSelection\"",
+                                             "replacement-rate" = 0.3
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Read Simulations NoFragmentation No Selection  regular Patterns
+
+```{r readSimulationsNoFragNSelectionRegular, echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_sloss_simul(outpath,"NoSelection_pf00_nofrag_dd1-3_brb1-4_regular_ff.csv") %>% 
+  mutate(habitat_patch_size = world_width, max_patch = (world_width+1) * (world_height+1) )
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
+
+```
+
+## Not fragmented communities Lambda = 1.7 - 4 - Hierarchical communities
+
+```{r Hierarchical_SLOSS_lambda14_notfrag,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="Hierarchical_pf00_nofrag_dd1-3_brb1-4_regular",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "prob-frag" =  0.6,
+                                             "Deforestation-at-time" = 0,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"Hierarchical\"",
+                                             "replacement-rate" = 0.3
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Read Simulations Not Fragmented  Hierarchical = habitat_patch_size == 200
+
+```{r readSimulationsNoFragHierarchical, echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_sloss_simul(outpath,"Hierarchical_pf00_nofrag_dd1-3_brb1-4_regular_ff.csv") %>% 
+  mutate(habitat_patch_size = world_width, max_patch = (world_width+1) * (world_height+1) )
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
+
+#mdl <- sloss_sim %>% filter( Deforestation_at_time==0)
+
+```
+
+## Not fragmented communities Lambda = 1.7 - 4 - Selection communities
+
+```{r BirthSelection_SLOSS_lambda14_notfrag,echo=FALSE,message=FALSE,warning=FALSE,eval=TRUE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="BirthSelection_pf00_nofrag_dd1-3_brb1-4_regular",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "prob-frag" =  0.6,
+                                             "Deforestation-at-time" = 0,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"BirthSelection\"",
+                                             "replacement-rate" = 0.3
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Read Simulations Not Fragmented  Hierarchical = habitat_patch_size == 200
+
+```{r readSimulationsNoFragBirthSelection, echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_sloss_simul(outpath,"BirthSelection_pf00_nofrag_dd1-3_brb1-4_regular_ff.csv") %>% 
+  mutate(habitat_patch_size = world_width, max_patch = (world_width+1) * (world_height+1) )
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
+
+#mdl <- sloss_sim %>% filter( Deforestation_at_time==0)
+
+```
 

### 2023-08-22 ms_SLOSS.Rmd
Update manuscript

@@
 As global biodiversity is rapidly declining due to habitat loss, it is important to determine how to protect it.
-Still, the policies of many conservation agencies are prioritizing the preservation of a single large habitat area (SL) vs several small areas of the same size (SS), besides empirical support favoring SS.
+Still, the policies of many conservation agencies are prioritizing the preservation of a single large habitat area (SL) vs several small areas of the same size (SS), besides empirical support favouring SS.
 However, to date, most of the theories regarding this issue were based on single-species models.
@@
 Global biodiversity is rapidly declining due to habitat loss (quotes).
-Moreover, habitat fragmentation occurs at the same time with habitat loss, and the effects of both processes are difficult to disentangle.
+Moreover, habitat fragmentation occurs at the same time as habitat loss, and the effects of both processes are difficult to disentangle.
 Landscape connectivity and organism mobility as a critical feature for population persistence [@Taylor1993].
-Starting from a landmark paper [@Diamond1975], conservation biologists engaged in the challenging task of recommending protection measures given the environmental crisis coupled with lack of knowledge of most of the world’s biodiversity.
+Starting from a landmark paper [@Diamond1975], conservation biologists engaged in the challenging task of recommending protection measures given the environmental crisis coupled with a lack of knowledge of most of the world’s biodiversity.
 Early recommendations claimed to follow ecological theory, in particular island biogeography theory to derive simple thumb rules decisions in order to better design protected areas.
@@
 
-The SLOSS debate has been an important academic and practical discussion among conservation biology.
+The SLOSS debate has been an important academic and practical discussion in conservation biology.
 In practice, a series of decision-maker targeted recommendations were proposed and remain rather uncritically in several conservation instances (quotes).
-While the overall habitat loss is widely accepted as a process leading to long-term biodiversity loss (quotes), the propose that single large reserves retain more species that several small reserves of the same area (SL > SS) remain at a similar hierarchical state, even after several criticisms (reviewed in  @Fahrig2017).
+While the overall habitat loss is widely accepted as a process leading to long-term biodiversity loss (quotes), the proposition that single large reserves retain more species than several small reserves of the same area (SL > SS) remain at a similar hierarchical state, even after several criticisms (reviewed in  @Fahrig2017).
 Moreover, empirical observational and experimental data shows the predominance of the opposite pattern (SS > SL) ever since the 80s and this fact has been reinforced to this day [@Simberloff1982; @Fahrig2020; @Hammill2020], add more.
-The actual processes driving to it remains elusive and it is probably multi-causal (quotes).
+The actual processes driving to it remain elusive and it is probably multi-causal (quotes).
 In that scenario, model simulations could produce useful predictions targeting the relevant variables in order to better understand the processes leading to non-intuitive patterns.
@@
 Moreover, spatial dimension in several works has been neglected as mere noise around long-term equilibrium dynamics (quotes, but see @Rybicki2013).
-In that context, spatial-explicit multi-species modeling could be meaningful as several organisms retain dispersal distance and other dispersal-related life traits far consistently among species (quotes).
+In that context, spatial-explicit multi-species modelling could be meaningful as several organisms retain dispersal distance and other dispersal-related life traits far consistently among species (quotes).
 
@@
 
-The environment comprises a 2D grid of patches that can be either suitable habitat or degraded. Bird agents have state variables for species identity, size, and color. Globals contain key parameters.
+The environment comprises a 2D grid of patches that can be either suitable habitat or degraded. Bird agents have state variables for species identity, size, and colour. Globals contain key parameters.
 
@@
 The model implements stochasticity in multiple processes including dispersal distance, loss patterns, migration, and species selection.
-Emergent macro-level diversity and abundance patterns result from individual-level bird behaviors.
+Emergent macro-level diversity and abundance patterns result from individual-level bird behaviours.
 
@@
 * **MultipleContactNeutralMigrat.nlogo** A multiple species neutral/hierarchical model described in with 
-	degraded sites that cannot be colonized
+        degraded sites that cannot be colonised
 
@@
 
-Regular habitat patterns: squares of a range of sizes are distributed regularly over the simulation grid. The side of the squares where 4,20,30,63 sites.
+Regular habitat patterns: squares of a range of sizes are distributed regularly over the simulation grid. The side of the squares were 4,20,30,63 sites.
 
 Random block habitat patterns: habitat blocks of the same size are randomly distributed over the simulation grid.
-The sizes are identical to the used for regular habitat simulations.
+The sizes are identical to those used for regular habitat simulations.
 This habitat simulation results in patches that have a minimum size equal to the block size but a maximum size that is bigger, so I characterize this with the maximum patch size.
@@
 
-![Relationship between number of species and habitat patch size for varying dispersal distances.
-Simulations were conducted using a model with 60% habitat destruction and regular habitat patterns, where all habitat patches are of equal size.](Figures/Species_Habitat_Dispersal_Lambda_regular.png)
+![Relationship between number of species and habitat patch size for different neutral models and varying dispersal distances.
+Simulations were performed using a neutral model with a constant migration rate (0.0001).
+Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
+The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.
+Within the neutral models, three distinct approaches were employed.
+In the "No selection" model, juveniles dispersed to a random site within a distance determined by the dispersal kernel.
+If the site was occupied or non-habitat, they perished.
+In the hierarchical model, the key difference was that if a site was occupied, a superior competitor could replace an inferior competitor with a probability of replacement (0.3).
+Lastly, the "birth selection" model is identical the "No selection" model, except that juveniles selected an unoccupied habitable site within a maximum distance dictated by the dispersal kernel.](Figures/AcrossBehav_Species_Dispersal_Lambda2_regular.png)
 
-![Relationship between Shannon Diversity and habitat patch size for varying dispersal distances.
-Simulations were conducted using a model with 60% habitat destruction and regular habitat patterns, where all habitat patches are of equal size.](Figures/Shannon_Habitat_Dispersal_Lambda_regular.png)
-
-![Relationship between number of species and maximum patch size for varying dispersal distances.
-Simulations were conducted using a model with 60% habitat destruction and random block habitat patterns, where habitat blocks of the same size are randomly distributed over the simulation grid (i.e., the maximum patch size).](Figures/Species_MaxPatch_Dispersal_Lambda_block.png)
-
-![Relationship between Shannon Diversity and maximum patch size for varying dispersal distances.
-Simulations were conducted using a model with 60% habitat destruction and random block habitat patterns, where habitat blocks of the same size are randomly distributed over the simulation grid (i.e., the maximum patch size).](Figures/Shannon_MaxPatch_Dispersal_Lambda_block.png)
+![Relationship between Shannon Diversity and habitat patch size for different neutral models and varying dispersal distances.
+Simulations were performed using a neutral model with a constant migration rate (0.0001) and relationship of growth rate over death rate of 2.
+Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
+The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.
+Within the neutral models, three distinct approaches were employed.
+In the "No selection" model, juveniles dispersed to a random site within a distance determined by the dispersal kernel.
+If the site was occupied or non-habitat, they perished.
+In the hierarchical model, the key difference was that if a site was occupied, a superior competitor could replace an inferior competitor with a probability of replacement (0.3).
+Lastly, the "birth selection" model is identical the "No selection" model, except that juveniles selected an unoccupied habitable site within a maximum distance dictated by the dispersal kernel.](Figures/AcrossBehav_Shannon_Dispersal_Lambda2_regular.png)
 
@@
 
-Overall, our results confirm the previous found pattern and nearly all models retained more species under the SS rather than the SL scenario.
+Overall, our results confirm the previously found pattern and nearly all models retained more species under the SS rather than the SL scenario.
 In both cases, the loss of species is relevant when habitat loss takes place, and no significant extinction biased towards SS occurs as a long-term dynamic.
@@
 
-Our model further illustrate that available patches are not always re-occupied following fragmentation even if the minimum size is met, therefore reinforcing the fact that dispersal is the main driver of the pattern, and that chance is involved in the final species pool when fragmentation takes place.
+Our model further illustrates that available patches are not always re-occupied following fragmentation even if the minimum size is met, therefore reinforcing the fact that dispersal is the main driver of the pattern, and that chance is involved in the final species pool when fragmentation takes place.
 Empirical data in large carnivores, for instance, support that several available patches are not occupied after isolation [@Paviolo2016].
@@
 
-We acknowledge the limitations of our study, and consider it a strength that we have identified them in this initial model. This allows us to continue working on them in future research.
+We acknowledge the limitations of our study and consider it a strength that we have identified them in this initial model. This allows us to continue working on them in future research.
 

### 2023-08-22 R/functions.r

@@ -37,10 +37,14 @@ read_netlogo_simul <- function(fname,skip=6){
 
 
 read_sloss_simul <- function(outpath,fname){
-  mdl <- read_netlogo_simul(file.path(outpath,fname),skip=0) %>% 
-  mutate(mean_free_path=str_sub(mean_free_path,2,-2)) %>% 
-  separate(mean_free_path, c("non_degraded_mfp","degraded_mfp"), sep=" ",convert=TRUE) %>% rowwise() %>%
-  mutate(netlogo_evaluate_patch_distr(find_cluster_sizes))
+  mdl <- read_netlogo_simul(file.path(outpath,fname),skip=0)
+  if( "mean_free_path" %in% names(mdl)) {
+    mdl <- mdl %>% 
+    mutate(mean_free_path=str_sub(mean_free_path,2,-2)) %>% 
+    separate(mean_free_path, c("non_degraded_mfp","degraded_mfp"), sep=" ",convert=TRUE) %>% rowwise() %>%
+    mutate(netlogo_evaluate_patch_distr(find_cluster_sizes))
+  } 
+  return(mdl)
 }
 
 #' Estimate the critical point of species living inside and habitat (birds inside forest) 

### 2023-08-22 R/test_nlrx_mcnm.r

@@ -1,0 +1,46 @@
+#
+nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path"),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "birth-rate-birds" = 2,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.001,
+                                             "birds-dispersal-distance" = 1.1,
+                                             "Video" = "false",
+                                             "prob-frag" =  0.6,
+                                             "habitat-patch-size" = 3,
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"NoSelection\"",
+                                             "replacement-rate" = 0.0
+                            ))
+
+
+#
+# Run 10 times this do not set the run variable for output 
+#
+# nseeds * number_of_parameters / split = 10 * 300 / 4 - total number of jobs = nseeds * split = 10*4
+#
+
+nl@simdesign <- simdesign_simple(nl=nl, nseeds=1)
+
+results <- run_nl_all(nl)
+
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+
+#mdl <- tibble()
+mdl1 <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_simple.csv"),skip=0)
+mdl <- bind_rows(mdl,mdl1)
+

### 2023-10-02 ms_SLOSS.Rmd
Updated version minor corrections

@@
 As global biodiversity is rapidly declining due to habitat loss, it is important to determine how to protect it.
-Still, the policies of many conservation agencies are prioritizing the preservation of a single large habitat area (SL) vs several small areas of the same size (SS), besides empirical support favouring SS.
+Still, the policies of many conservation agencies are prioritizing the preservation of a single large habitat area (SL) vs.
+several small areas of the same size (SS), besides empirical support favouring SS.
 However, to date, most of the theories regarding this issue were based on single-species models.
-The aim of this work is to explore spatial multi-species models to verify the SS > SL pattern and determine the causes of it.
+This work aims to explore spatial multi-species models to verify the SS > SL pattern and determine the causes of it.
 We used three types of models: a classical neutral, neutral with habitat selection, and a competitive hierarchical model.
@@
 High dispersal rates produce homogenization of species distribution and a reduction of species richness and this is why habitat fragmentation by habitat loss retains more species when patches are smaller.
-A prediction of the model is that if a set of species have more dispersal capabilities the SS effect would be lower than for species with less dispersal.
+The model predicts that if a set of species have more dispersal capabilities the SS effect would be lower than for species with less dispersal.
 
@@
 Starting from a landmark paper [@Diamond1975], conservation biologists engaged in the challenging task of recommending protection measures given the environmental crisis coupled with a lack of knowledge of most of the world’s biodiversity.
-Early recommendations claimed to follow ecological theory, in particular island biogeography theory to derive simple thumb rules decisions in order to better design protected areas.
+Early recommendations claimed to follow ecological theory, in particular island biogeography theory to derive simple thumb rules decisions to better design protected areas.
 However, as noted rapidly, species/area relationships do not have a clear association with fragmentation per se but only with overall habitat loss [@Simberloff1982].
@@
 
-![Relationship between number of species and habitat patch size for different neutral models and varying dispersal distances.
-Simulations were performed using a neutral model with a constant migration rate (0.0001).
-Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
-The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.
-Within the neutral models, three distinct approaches were employed.
-In the "No selection" model, juveniles dispersed to a random site within a distance determined by the dispersal kernel.
-If the site was occupied or non-habitat, they perished.
-In the hierarchical model, the key difference was that if a site was occupied, a superior competitor could replace an inferior competitor with a probability of replacement (0.3).
-Lastly, the "birth selection" model is identical the "No selection" model, except that juveniles selected an unoccupied habitable site within a maximum distance dictated by the dispersal kernel.](Figures/AcrossBehav_Species_Dispersal_Lambda2_regular.png)
+![Relationship between the number of species and habitat patch size for varying dispersal distances.
+Simulations were conducted using a model with 60% habitat destruction and regular habitat patterns, where all habitat patches are of equal size.](Figures/Species_Habitat_Dispersal_Lambda_regular.png)
 
-![Relationship between Shannon Diversity and habitat patch size for different neutral models and varying dispersal distances.
-Simulations were performed using a neutral model with a constant migration rate (0.0001) and relationship of growth rate over death rate of 2.
-Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
-The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.
-Within the neutral models, three distinct approaches were employed.
-In the "No selection" model, juveniles dispersed to a random site within a distance determined by the dispersal kernel.
-If the site was occupied or non-habitat, they perished.
-In the hierarchical model, the key difference was that if a site was occupied, a superior competitor could replace an inferior competitor with a probability of replacement (0.3).
-Lastly, the "birth selection" model is identical the "No selection" model, except that juveniles selected an unoccupied habitable site within a maximum distance dictated by the dispersal kernel.](Figures/AcrossBehav_Shannon_Dispersal_Lambda2_regular.png)
+![Relationship between Shannon Diversity and habitat patch size for varying dispersal distances.
+Simulations were conducted using a model with 60% habitat destruction and regular habitat patterns, where all habitat patches are of equal size.](Figures/Shannon_Habitat_Dispersal_Lambda_regular.png)
+
+![Relationship between the number of species and maximum patch size for varying dispersal distances.
+Simulations were conducted using a model with 60% habitat destruction and random block habitat patterns, where habitat blocks of the same size are randomly distributed over the simulation grid (i.e., the maximum patch size).](Figures/Species_MaxPatch_Dispersal_Lambda_block.png)
+
+![Relationship between Shannon Diversity and maximum patch size for varying dispersal distances.
+Simulations were conducted using a model with 60% habitat destruction and random block habitat patterns, where habitat blocks of the same size are randomly distributed over the simulation grid (i.e., the maximum patch size).](Figures/Shannon_MaxPatch_Dispersal_Lambda_block.png)
 
@@
 Overall, our results confirm the previously found pattern and nearly all models retained more species under the SS rather than the SL scenario.
-In both cases, the loss of species is relevant when habitat loss takes place, and no significant extinction biased towards SS occurs as a long-term dynamic.
-The extinction of species in patches depends on the quotient between reproduction and extinction rates in relation to patch size, i.e., the ability of a certain patch to be a source of propagules.
+In both cases, the loss of species is relevant when habitat loss takes place, and no significant extinction bias towards SS occurs as a long-term dynamic.
+The extinction of species in patches depends on the quotient between reproduction and extinction rates to patch size, i.e., the ability of a certain patch to be a source of propagules.
 The only case when SL > SS is when dispersal is not enough to connect patches, and size is smaller than the minimum to prevent patch extinction, as predicted elsewhere [@Rybicki2013].
@@
 In that scenario, SS and SL are approximately equivalent for species diversity retention.
-A prediction of the model is that if a set of species have more dispersal capabilities the SS effect would be smaller than for species with less dispersal abilities.
+The model predicts that if a set of species have more dispersal capabilities the SS effect would be smaller than for species with less dispersal abilities.
 
@@
 Evidence suggests that small patches, even single structures such as isolated trees, can be important for biodiversity and landscape connectivity [@Shafer1995; @Manning2006].
-While connectivity management is case-specific, our results further support past claims [@Fahrig2020;@Fahrig2021], that assuming that SL is always better do not have empirical support.
-In our work, spreading of risk was not modelled (__maybe as time lag in at-patch competition?__), therefore reinforcing that the sampling effect is enough to produce a SS > SL pattern if dispersal is small enough to produce spatial autocorrelation of occupied patches [@Simberloff1976].
+While connectivity management is case-specific, our results further support past claims [@Fahrig2020;@Fahrig2021], that assuming that SL is always better does not have empirical support.
+In our work, spreading of risk was not modelled (__maybe as time lag in at-patch competition?__), therefore reinforcing that the sampling effect is enough to produce an SS > SL pattern if dispersal is small enough to produce spatial autocorrelation of occupied patches [@Simberloff1976].
 
@@
 
-We acknowledge the limitations of our study and consider it a strength that we have identified them in this initial model. This allows us to continue working on them in future research.
+We acknowledge the limitations of our study and consider it a strength that we have identified in this initial model. This allows us to continue working on them in future research.
 

### 2023-10-10 RunSims.Rmd
Many changes

--- 2023-08-05 RunSims.Rmd
@@
 
+} else if(Sys.info()['nodename'] =="MacLeonardo.local") {                     # Mac
+  simfolder <- "~/Dropbox/Projects/DynamicForestExtinction"
 }
@@
 mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
-                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
 
@@
 
+mdl1 %>%  
+  ggplot(aes(x = max_patch_mean, y = degraded_mfp_mean, color = factor(habitat_patch_size))) + 
+  geom_point(  size = 1.2) + guides(color = "none") #+
+
+mdl1 %>%  
+  ggplot(aes(x = max_patch_mean, y = non_degraded_mfp_mean, color = factor(habitat_patch_size))) + 
+  geom_point(  size = 1.2) + guides(color = "none") #+
+
+mdl1 %>%  
+  ggplot(aes(x = degraded_mfp_mean, y = non_degraded_mfp_mean, color = factor(habitat_patch_size))) + 
+  geom_point(  size = 1.2) + guides(color = "none") #+
+
 # require(plotly)
@@
   ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
   facet_wrap(~ birth_rate_birds, ncol = 4) + 
@@
   ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
   facet_wrap(~ birth_rate_birds, ncol = 4) + 
@@
   ggplot(aes(x = max_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
@@
 mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
-                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch) ,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,non_degraded_mfp,degraded_mfp,max_patch) ,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/
 
@@
 
+mdl1 %>%  
+  ggplot(aes(x = max_patch_mean, y = degraded_mfp_mean, color = factor(habitat_patch_size))) + 
+  geom_point(  size = 1.2) + guides(color = "none") #+
+
+mdl1 %>%  
+  ggplot(aes(x = max_patch_mean, y = non_degraded_mfp_mean, color = factor(habitat_patch_size))) + 
+  geom_point(  size = 1.2) + guides(color = "none") #+
+
+mdl1 %>%  
+  ggplot(aes(x = degraded_mfp_mean, y = non_degraded_mfp_mean, color = factor(habitat_patch_size))) + 
+  geom_point(  size = 1.2) + guides(color = "none") #+
+
 #
@@
   ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
@@
 mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
-  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
+  ggplot(aes(x = max_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
   facet_wrap(~ birth_rate_birds, ncol = 4) + 
@@
   ggplot(aes(x = max_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
   facet_wrap(~ birth_rate_birds, ncol = 4) + 
@@
 
-mdl <- sloss_sim %>% filter(deforestation_type == "regular", birds_behavior=="Hierarchical")
+mdl <- sloss_sim %>% filter(deforestation_type == "regular", birds_behavior=="Hierarchical", habitat_patch_size != 199)
 names(mdl)
@@
   ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
   facet_wrap(~ birth_rate_birds, ncol = 4) + 
@@
   ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
   facet_wrap(~ birth_rate_birds, ncol = 4) + 
@@
 
-## Not fragmented communities Lambda = 1.7 - 4
+# Plot for birth selection communities regular patterns
 
-```{r SLOSS_notfrag_simulations, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
-require(tidyverse)
+```{r ExploratoryPlotsSelectionRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 
-source("R/functions.r")
-
-# Not fragmented communities = habitat_patch_size = 100
-#
-mdl <- read_netlogo_simul("Simulations/NoHiBi_pf00_hpf3_dd1-3_lambda1.7-4.csv") %>% mutate(habitat_patch_size=100)
-
-# New Simulations 
-#
-mdl <- read_netlogo_simul("Simulations/NoHiBi_pf00_hpf3_dd1-3_lambda1.1-4.csv") %>% mutate(habitat_patch_size=100)
-
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-# Add Migration0_at_deforestation=FALSE to previous simulations 
-#
-unique(sloss_sim$Migration0_at_deforestation)
-sloss_sim <- sloss_sim %>% mutate( Migration0_at_deforestation = if_else(is.na(Migration0_at_deforestation), FALSE, Migration0_at_deforestation )) 
-unique(sloss_sim$habitat_patch_size)
-#sloss_sim <- sloss_sim %>% mutate( habitat_patch_size = if_else(habitat_patch_size==0, 100, habitat_patch_size ))
-
-saveRDS(sloss_sim, "Simulations/sloss_sim.rds")
-
+mdl <- sloss_sim %>% filter(deforestation_type == "regular", birds_behavior=="BirthSelection", habitat_patch_size != 199)
 names(mdl)
 #
-# Plot for neutral communities 
-# I filter birth_rate_birds 1.1 because these communities survive only by migration
+# summarize data 
 #
-mdl1 <- mdl %>% filter(birth_rate_birds!=1.1) %>% group_by(birds_dispersal_distance, birth_rate_birds,birds_behavior) %>% summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(birds_at_deforestation:count_birds,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+names(mdl1)
 
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_number_of_species,color=birth_rate_birds)) + geom_point() + geom_line() + scale_color_viridis_d(name="lambda") + facet_wrap( ~birds_behavior )  + theme_bw() + ylab("Number of species") + xlab("Dispersal Distance")
-ggsave("Figures/Species_NoFrag_Dispersal_Lambda.png",width=8,height=6,units="in",dpi=600)
+# require(plotly)
+# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))
 
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=calc_shannon_diversity,color=birth_rate_birds)) + geom_point() + scale_color_viridis_d() + facet_wrap( ~birds_behavior ) + theme_bw()
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
 
-mdl1 %>% ggplot(aes(x=birds_dispersal_distance,y=count_birds,color=birth_rate_birds)) + geom_point() + geom_line() + scale_color_viridis_d(name="lambda") + facet_wrap( ~birds_behavior )  + theme_bw() + ylab("Number of indiv.") + xlab("Dispersal Distance")
-ggsave("Figures/Indiv_NoFrag_Dispersal_Lambda.png",width=8,height=6,units="in",dpi=600)
+ggsave("Figures/Selection_Species_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
 
-mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3) %>%  mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=birth_rate_birds,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + scale_color_viridis_d() + theme_bw()+ facet_wrap( ~birds_behavior )
+# Shannon-Diversity  
+#
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/Selection_Shannon_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 
-## Add more simulations lambda = 2 with Migration
+## Plots across neutral hierarchical selection modelos regular fragmentation
 
-```{r SLOSS_combined_plots, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+```{r PlotsAcrossBehaviorRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 
-mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda2_Migrat.csv")
+mdl <- sloss_sim %>% filter(deforestation_type == "regular", habitat_patch_size != 199)
+names(mdl)
+#
+# summarize data 
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,birds_behavior) %>% summarise(across(birds_at_deforestation:count_birds,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+names(mdl1)
 
-sloss_sim <- bind_rows(sloss_sim,mdl)
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birds_behavior), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_wrap(~ birds_behavior, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
 
-mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda2_NoMigrat.csv") 
-sloss_sim <- bind_rows(sloss_sim,mdl)
+ggsave("Figures/AcrossBehav_Species_Dispersal_Lambda2_regular.png",width=8,height=6,units="in",dpi=600)
 
-mdl <- read_netlogo_simul("Simulations/NoHiBi_pf06_hpf3-61_dd1-3_lambda4_Migrat.csv") 
-sloss_sim <- bind_rows(sloss_sim,mdl)
+# Shannon-Diversity  
+#
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birds_behavior), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_wrap(~ birds_behavior, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
 
-save.image()
-
-#
-# 200 x 200 - Started simulation 10:47
-#
-mdl <- read_netlogo_simul("Simulations/NoHiBi_200_pf06_hpf3-61_dd1-3_lambda2_NoMigrat.csv")
-sloss_sim <- bind_rows(sloss_sim,mdl)
-sloss_sim <- sloss_sim %>% mutate( world_width=if_else(is.na(world_width),100,world_width))
-sloss_sim <- sloss_sim %>% mutate( world_width=if_else(world_width==99,100,world_width))
-
-saveRDS(sloss_sim, "Simulations/sloss_sim.rds")
+ggsave("Figures/AcrossBehav_Shannon_Dispersal_Lambda2_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 
-## Plots of combined behaviors and patch 100
+## Plots across neutral hierarchical selection modelos regular fragmentation
 
-```{r SLOSS_combined_plots1, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
-require(tidyverse)
+```{r PlotsAcrossBehaviorRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 
-source("R/functions.r")
+mdl <- sloss_sim %>% filter(birds_behavior=="NoSelection", habitat_patch_size != 199)
+names(mdl)
+#
+# summarize data 
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,deforestation_type) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+names(mdl1)
 
-#mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation | (birth_rate_birds==2 & habitat_patch_size==100)) %>%  
-sloss_sim %>% distinct(Migration0_at_deforestation, world_width)
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
 
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, deforestation_type), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_wrap(~ deforestation_type, ncol = 2) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/AcrossLoss_Species_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
 #
-# Count number of simulations
-#
-mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation, world_width==100 ) %>%  
-  count(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior)
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, deforestation_type), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_wrap(~ deforestation_type, ncol = 2) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Mean Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
 
-mdl1 <- sloss_sim %>% filter(Migration0_at_deforestation, world_width==100 ) %>%  
-  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior) %>%
-  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )
+ggsave("Figures/AcrossLoss_Species_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
 
-#mdl1 %>% filter(birds_dispersal_distance == 1.1 || birds_dispersal_distance == 3 || birds_dispersal_distance == 1.5) %>%  
-mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, deforestation_type), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_wrap(~ deforestation_type, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Mean Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
 
-ggsave("Figures/Species_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)
-
-mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_shannon_diversity,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Shannon diversity") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% NoMigration")
-ggsave("Figures/Shannon_PatchSize_Dispersal_NoMigration.png",width=8,height=6,units="in",dpi=600)
-
-#
-# With migration all the time
-#
-
-mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation ) %>%  
-  count(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior)
-
-#
-# Plot without SD
-#
-mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birth_rate_birds == 2, world_width==100 ) %>%  
-  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds,birds_behavior) %>%
-  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )
-
-mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% Migration")
-
-ggsave("Figures/Species_PatchSize_Dispersal_Migration.png",width=8,height=6,units="in",dpi=600)
-
-mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_shannon_diversity,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birds_behavior ) + ylab("Shannon diversity") + xlab("Patch Size") + ggtitle("Lambda=2 Degraded 60% Migration")
-ggsave("Figures/Shannon_PatchSize_Dispersal_Migration.png",width=8,height=6,units="in",dpi=600)
-
-#
-# From Lamda 2 to lambda 4 Neutral
-#
-
-mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birds_behavior == "NoSelection" ) %>%  
-  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds) %>%
-  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )
-
-mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birth_rate_birds ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Neutral - Degraded 60% - Migration 0.001")
-ggsave("Figures/Species_lambda2-4_PatchSize_Dispersal_NoSelection.png",width=8,height=6,units="in",dpi=600)
-
-#
-# From Lamda 2 to lambda 4 Hierarchical
-#
-
-mdl1 <- sloss_sim %>% filter(!Migration0_at_deforestation & habitat_patch_size!=100 & birds_behavior == "Hierarchical" ) %>%  
-  group_by(habitat_patch_size,birds_dispersal_distance, birth_rate_birds) %>%
-  summarise(across(birds_at_deforestation:count_birds,mean)) %>% mutate(birth_rate_birds = factor(birth_rate_birds)  )
-
-mdl1 %>% mutate(birds_dispersal_distance=factor(birds_dispersal_distance)) %>% ggplot(aes(x=habitat_patch_size,y=calc_number_of_species,color=birds_dispersal_distance)) + geom_point() + geom_line() + scale_color_viridis_d(name="Dispersal") + theme_bw()+ facet_wrap( ~birth_rate_birds ) + ylab("Number of species") + xlab("Patch Size") + ggtitle("Hierarchical - Degraded 60% - Migration 0.001")
-ggsave("Figures/Species_lambda2-4_PatchSize_Dispersal_Hierarchical.png",width=8,height=6,units="in",dpi=600)
+ggsave("Figures/AcrossLoss_Shannon_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
 

### 2023-10-10 ms_SLOSS.Rmd
Many changes

@@
 
-Global biodiversity is rapidly declining due to habitat loss (quotes).
-Moreover, habitat fragmentation occurs at the same time as habitat loss, and the effects of both processes are difficult to disentangle.
-Landscape connectivity and organism mobility as a critical feature for population persistence [@Taylor1993].
+Global biodiversity is rapidly declining due to habitat loss [@Young2016;@Chase2020;@Newbold2015].
+Moreover, habitat fragmentation occurs at the same time as habitat loss, and the effects of both processes are difficult to disentangle [@Haddad2015].
+Landscape connectivity and organism mobility are a critical feature for population persistence [@Taylor1993].
 Starting from a landmark paper [@Diamond1975], conservation biologists engaged in the challenging task of recommending protection measures given the environmental crisis coupled with a lack of knowledge of most of the world’s biodiversity.
@@
 
-The model represents a landscape occupied by multiple bird species that undergoes habitat degradation over time.
+The model represents a landscape occupied by multiple species that undergoes habitat degradation over time.
 The purpose is to explore the impact of fragmentation patterns and species interaction mechanisms on diversity and abundance distributions.
 
-The environment comprises a 2D grid of patches that can be either suitable habitat or degraded. Bird agents have state variables for species identity, size, and colour. Globals contain key parameters.
+The environment comprises a 2D grid of patches that can be either suitable habitat or degraded.
+Individuals of species have state variables for species identity, size, and colour.
+Globals contain key parameters.
 
-At initialization, an intact landscape is seeded with bird agents having randomized species identities.
-Habitat loss occurs probabilistically at a set time step based on three spatial patterns: random, regular fragmentation, or block removal.
+At initialization, an intact landscape is seeded with species having randomized species identities.
+Habitat loss occurs probabilistically at a given time based on three spatial patterns: random, regular fragmentation, or block removal.
 
@@
 
-Neutral model: Offspring disperse locally based on power law dispersal kernel and settle in any vacant habitat
+Neutral model: Offspring disperse locally based on power law dispersal kernel and settle in any vacant habitat.
+Two flavours "No Selection" propagules try to move to the distance determined by the dispersal kernel, if propagules fall in an inhabitable or occupied site they die, they maximize dispersal; "Selection" propagules select the from the available sites within the distance determined by dispersal, so propagules maximize survival.
+
 Hierarchical model: Species can displace lower ranked species during dispersal
+
 Immigration allows spontaneous appearance of birds at habitat edges
@@
 ## Results
+
+* With the smallest patch size and gowth rates less than 3 species cannot survive, for the neutral and the hierarchical models.
 

### 2023-10-25 ms_SLOSS.Rmd
Cambios en metodos e intro

@@
 
-The SLOSS debate has been approached from several theoretical frameworks, but most of them were based on single-species models.
-Moreover, spatial dimension in several works has been neglected as mere noise around long-term equilibrium dynamics (quotes, but see @Rybicki2013).
-In that context, spatial-explicit multi-species modelling could be meaningful as several organisms retain dispersal distance and other dispersal-related life traits far consistently among species (quotes).
+The SLOSS debate has been approached from several theoretical frameworks, but most of them were based on single-species models [@Fahrig2021].
+Moreover, spatial dimension in several works has been neglected as mere noise around long-term equilibrium dynamics.
+Recent studies using spatially explicit simulations revealed that while high fragmentation leads to more local extinctions due to the reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].
+
+We considered only one amount of habitat destruction in combination with different community dynamics, taking into account empoverished communities with different dispersal related traits, and different patterns of habitat loss.
+
+<!-- In that context, spatial-explicit multi-species modelling could be meaningful as several organisms retain dispersal distance and other dispersal-related life traits far consistently among species (quotes).-->
 
@@
 
-### Model
+### A spatial stochastic model
 
-The model represents a landscape occupied by multiple species that undergoes habitat degradation over time.
-The purpose is to explore the impact of fragmentation patterns and species interaction mechanisms on diversity and abundance distributions.
+We developed a spatially explicit multi-species model to explore the relationship between habitat loss and species preservation.
+In this model space is discretized into a grid and only one individual of one species can occupy a particular position.
+Each position represents an area fixed by the investigator to mimic the real system.
+Time is continuous, so the update of the model is asynchronous.
+Sites are selected at random and to perform one complete time interval $J$ sites have to be updated, where $J$ is the size of the grid.
 
-The environment comprises a 2D grid of patches that can be either suitable habitat or degraded.
-Individuals of species have state variables for species identity, size, and colour.
-Globals contain key parameters.
+We use absorving boundary conditions, which means that if an individual try to disperse outside the grid it dies.
+The size of the community is given by *J = dimX* x *dimY*, where *dimX* and *dimY* are the dimensions of the grid.
+Thus *J* is the maximum number of individuals in the simulated area.
+As in a classical neutral model, there is a metacommunity, i.e.
+a regional species pool from were species migrate [@Hubbell2001].
+All individuals have the same parameters and each species is assigned an indicator number.
 
-At initialization, an intact landscape is seeded with species having randomized species identities.
-Habitat loss occurs probabilistically at a given time based on three spatial patterns: random, regular fragmentation, or block removal.
+The model consists of three variations: a classical neutral model, a neutral model with habitat selection, and a competitive hierarchical model.
 
-During each time step, the following sequence occurs:
+The classical neutral variation where propagules (e.g., seeds) disperse passively and cannot actively choose their establishment site.
+During reproduction events, individuals disperse their propagules blindly to surrounding patches.
+If a patch is already occupied, the propagule fails to establish, resulting in its death.
+This variation captures the random dispersal and colonization processes typical of many plant species.
 
-Habitat loss event if scheduled
+The neutral variation with habitat selection where individuals have the ability to actively select suitable habitat patches.
+In this variation, propagules have the capacity to choose unoccupied patches, ensuring their survival and successful establishment.
+This modification reflects the influence of habitat availability and selection on species colonization and persistence.
+By incorporating habitat selection, this variation better captures the ecological processes and behaviors of species that are capable of actively seeking out suitable habitats for reproduction and survival.
 
-Calculation of probabilities for reproduction, dispersal, migration
+The competitive hierarchical variation extends the neutral variation with an additional layer of complexity by incorporating competitive interactions among species.
+Each species have an indicator number that determines its competitive dominance level, and more dominant species have the ability to replace less dominant ones at a specified replacement rate.
+This variation captures the influence of species interactions and competitive exclusion on species composition and abundance.
+It is particularly relevant for understanding communities where competitive interactions play a significant role in shaping species distributions and community structure.
 
-Birds undergo probabilistic dispersal and reproduction based on:
+There following process are included in the model: habitat loss, migration, reproduction, local dispersal, competition and death. After setting initial conditions the following events can happen:
 
-Neutral model: Offspring disperse locally based on power law dispersal kernel and settle in any vacant habitat.
-Two flavours "No Selection" propagules try to move to the distance determined by the dispersal kernel, if propagules fall in an inhabitable or occupied site they die, they maximize dispersal; "Selection" propagules select the from the available sites within the distance determined by dispersal, so propagules maximize survival.
+(1) Habitat loss: the sites are marked as non-habitable, making them unsuitable for occupancy, up to a fixed percentage of degraded habitat. Different degradation scenarios are modeled:
 
-Hierarchical model: Species can displace lower ranked species during dispersal
+  Regular degradation: Habitat is removed so as habitable patches form a regular pattern of squares with side given by an habitat-size parameter.
+As the percentage of degraded habitat is constant varying the habitat-size parameter results in either few large habitat patches (SL) or several small (SS).
 
-Immigration allows spontaneous appearance of birds at habitat edges
+  Random block degradation: Habitat square patches of a fixed size (habitat-size parameter) are located at random throughout the landscape.
+The sizes are identical to those used for regular degradation simulations, but it results in habitat patches that have a minimum size corresponding to the habitat-size  but a maximum size that is bigger, because habitat patches can superpose between them producing a larger patch.
+Then, I characterize this with the maximum habitat patch size.
 
-Data collection on species diversity, abundance, and spatial distribution
+(2) With $migration-rate$ an individual of a species $i$ can migrate from the metacommunity. A random site on the border of the grid is selected and if it is empty and habitable the site is occupied.
 
-The model implements stochasticity in multiple processes including dispersal distance, loss patterns, migration, and species selection.
-Emergent macro-level diversity and abundance patterns result from individual-level bird behaviours.
+(3) All individuals reproduce with $birth-rate$. When an individual reproduces the propagule disperses to the neighborhood with a dispersal kernel. Here we use an inverse power kernel [@Marco2011]:
 
-Key outputs analyzed are species occupancy, diversity indices, abundance distributions, and spatial patterns.
-Experiments examine the dynamics under varying habitat loss, dispersal, and competition scenarios.
+  $d(x) =  \frac{\alpha -1}{x_{min}} \left(\frac{x}{x_{min}} \right)^{-\alpha}$ with $mean =\frac{\alpha-1}{\alpha-2}x_{min}$ where $\alpha > 1$ and $x \ge x_{min}$.
 
-* **MultipleContactNeutralMigrat.nlogo** A multiple species neutral/hierarchical model described in with 
-        degraded sites that cannot be colonised
+  where $d(x)$ is the probability that an individual disperses a distance $x$ from the parent. In all cases we used $x_{min} = 1$. We use as parameter the mean $dispersal-distance$.
 
-        <https://github.com/lsaravia/DynamicForestExtinction>
+  What happens to the propagule when arrives to the selected site is determined by the variation of the model simulated:
 
-Regular habitat patterns: squares of a range of sizes are distributed regularly over the simulation grid. The side of the squares were 4,20,30,63 sites.
+  a. Classical Neutral: If the target patch is already occupied by a species, the propagule perishes.
 
-Random block habitat patterns: habitat blocks of the same size are randomly distributed over the simulation grid.
-The sizes are identical to those used for regular habitat simulations.
-This habitat simulation results in patches that have a minimum size equal to the block size but a maximum size that is bigger, so I characterize this with the maximum patch size.
+  b. Neutral with Habitat Selection:  Propagules in this version selectively target unoccupied patches for reproduction within the distance given by the dispersal kernel, ensuring their survival.
+
+  c. Competitive Hierarchical: species are organized based on competitive dominance based on their id number, if the site is occupied by a species with lower id-number is replaced by the propagule.
+
+(4) Individuals die with a $death-rate$ independent of other events.
+
+For all simulations qwe used a metacommunity of 100 species.
+The simulations commence with the colonization of an empty habitat area.
+After 600 time steps fixed amount of 60% habitat is subsequently subjected to destruction, employing regular or random block degradation to mimic habitat fragmentation due to habitat loss.
+
+The Netlogo model source code is available at Github <https://github.com/lsaravia/DynamicForestExtinction> and the R source code to perform simulations and analysis is located at <https://github.com/EcoComplex/multiSLOSS>.
 

### 2023-11-27 ms_SLOSS.Rmd
Updated manuscript

@@
 In practice, a series of decision-maker targeted recommendations were proposed and remain rather uncritically in several conservation instances (quotes).
-While the overall habitat loss is widely accepted as a process leading to long-term biodiversity loss (quotes), the proposition that single large reserves retain more species than several small reserves of the same area (SL > SS) remain at a similar hierarchical state, even after several criticisms (reviewed in  @Fahrig2017).
+While the overall habitat loss is widely accepted as a process leading to long-term biodiversity loss [@Keil2015;@Hodgson2011], the proposition that single large reserves retain more species than several small reserves of the same area (SL > SS) remain at a similar hierarchical state, even after several criticisms (reviewed in  @Fahrig2017).
 Moreover, empirical observational and experimental data shows the predominance of the opposite pattern (SS > SL) ever since the 80s and this fact has been reinforced to this day [@Simberloff1982; @Fahrig2020; @Hammill2020], add more.
@@
 
-We considered only one amount of habitat destruction in combination with different community dynamics, taking into account empoverished communities with different dispersal related traits, and different patterns of habitat loss.
-
-<!-- In that context, spatial-explicit multi-species modelling could be meaningful as several organisms retain dispersal distance and other dispersal-related life traits far consistently among species (quotes).-->
+We considered only one amount of habitat destruction in combination with different community dynamics, neutral and competitive, taking into account healthy and mpoverished communities with different dispersal related traits, and different patterns of habitat loss.
 
@@
 In particular, we simulate species retention after different fragmentation scenarios of the same overall habitat loss, therefore accounting for the interacting effects of habitat loss and fragmentation.
+
+<!-- In that context, spatial-explicit multi-species modelling could be meaningful as several organisms retain dispersal distance and other dispersal-related life traits far consistently among species (quotes).-->
 
@@
 
-  $d(x) =  \frac{\alpha -1}{x_{min}} \left(\frac{x}{x_{min}} \right)^{-\alpha}$ with $mean =\frac{\alpha-1}{\alpha-2}x_{min}$ where $\alpha > 1$ and $x \ge x_{min}$.
+    $d(x) =  \left(\frac{x}{x_{min}} \right)^{-\alpha}$ with $mean =\frac{\alpha-1}{\alpha-2}x_{min}$ where $\alpha > 1$ and $x \ge x_{min}$.
 
@@
 
-  c. Competitive Hierarchical: species are organized based on competitive dominance based on their id number, if the site is occupied by a species with lower id-number is replaced by the propagule.
+    c.
+Competitive Hierarchical: Species are ranked by a competitive dominance hierarchy based on their internal identification   (id) number.
+If a dispersing propagule arrives at an occupied site, it will only establish if its id number is higher than the resident species.
+Otherwise, the propagule perishes without establishing.
+This simulates a system where dominant competitor species displace rarer or weaker species through interspecific competition during colonization of new sites.
+The hierarchy prioritizes species with lower id numbers as stronger competitors.
 
@@
 
-For all simulations qwe used a metacommunity of 100 species.
+For all simulations we used a metacommunity of 100 species.
 The simulations commence with the colonization of an empty habitat area.
 After 600 time steps fixed amount of 60% habitat is subsequently subjected to destruction, employing regular or random block degradation to mimic habitat fragmentation due to habitat loss.
+<!-- why we don't use completely random degradation patterns -->
+
+The parameters encompass a range of $growth-rates$ while holding the $death-rate$ fixed at 1.
+In this model, the relationship between these parameters, $\lambda = growth-rate / death-rate$, is the main determinant of the fate of an individual of a species.
+In similar non-spatial models, this relationship must exceed 1 for a species to persist, but in spatially explicit models the value needs to be higher, around 1.6.
+Thus we used a minimum $\lambda$ of 1.7, we also fixed the $migration-rate$ at 0.0001.
+We simulate each model variant using the parameters shown in Table 1.
+For the neutral classical version only, we also simulate random block habitat patterns.
+
+| $\lambda$ | Dispersal Distance | Habitat Patch size |
+| --------: | ---------:         | ---------:         |
+| 1.7       | 1.1                | 3                  |
+| 2.0       | 1.2                | 9                  |
+| 3.0       | 1.5                | 29                 |
+| 4.0       | 2                  | 61                 |
+|           | 3                  |                    |
+
+Table: Parameter values used in simulations.
+Factorial combinations of these parameters were used to run simulations.
+$\lambda$ represents the ratio of growth rate to death rate.
+"Dispersal distance" refers to the mean of the power law dispersal kernel.
+"Habitat patch size" indicates the side length of habitat patches modeled as squares.
+
+The classical neutral model and the habitat selection model could represent endpoints on a continuum of models where the matrix is either completely adverse or not adverse but still unsuitable for reproduction.
+In the habitat selection model, individuals can search across both the matrix and suitable habitat until finding an empty place.
+In contrast, for the classical neutral model, individuals have limited or no capacity to search, and the matrix is necessarily detrimental because it increases the probability of death by landing on an unsitable site.
 
@@
 
-* With the smallest patch size and gowth rates less than 3 species cannot survive, for the neutral and the hierarchical models.
+In our neutral model simulations with a migration rate of \(0.0001\) and regular habitat patterns, species richness varied with habitat patch sizes, influenced by dispersal distance and the \(\lambda\) parameter (Figure 1).
+Species richness increased with \(\lambda\), with the highest richness at smaller patch sizes for higher \(\lambda\) values.
+Conversely, a \(\lambda\) of 3 or less showed reduced richness at the smallest patch sizes for lower dispersal distances.
+The interplay of dispersal distances, habitat sizes, and \(\lambda\) impacts community survival and species richness.
+This suggests a threshold patch size influenced by these 3 parameters.
+Extended dispersal distances enhance species survival in scenarios with low \(\lambda\) and smaller patches but reduce richness beyond a certain threshold.
+Similar patterns are evident for Shannon diversity (Figure 2).
 
-![Relationship between the number of species and habitat patch size for varying dispersal distances.
-Simulations were conducted using a model with 60% habitat destruction and regular habitat patterns, where all habitat patches are of equal size.](Figures/Species_Habitat_Dispersal_Lambda_regular.png)
+![Relationship between number of species and habitat patch size for varying dispersal distances and $\lambda$ parameter (across panels).
+Simulations were performed using a neutral model with a constant migration rate (0.0001).
+Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
+The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.](Figures/Species_Habitat_Dispersal_Lambda_regular.png)
 
-![Relationship between Shannon Diversity and habitat patch size for varying dispersal distances.
-Simulations were conducted using a model with 60% habitat destruction and regular habitat patterns, where all habitat patches are of equal size.](Figures/Shannon_Habitat_Dispersal_Lambda_regular.png)
+![Relationship between Shannon Diversity and habitat patch size for varying dispersal distances and $\lambda$ parameter (across panels).
+Simulations were performed using a neutral model with a constant migration rate (0.0001).
+Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
+The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.](Figures/Shannon_Habitat_Dispersal_Lambda_regular.png)
 
-![Relationship between the number of species and maximum patch size for varying dispersal distances.
-Simulations were conducted using a model with 60% habitat destruction and random block habitat patterns, where habitat blocks of the same size are randomly distributed over the simulation grid (i.e., the maximum patch size).](Figures/Species_MaxPatch_Dispersal_Lambda_block.png)
+For the simulations made with random block habitat patterns, the patches do not have the same sizes but there is a distribution with a minimum size determined by the patch-size parameter and a bigger maximum size.
+Again there is a simimlar pattern with a maximum richness after the threshold for $\lambda$ values less than 3 and for values greater or equal the maximum is at the lower habitat patch size (Figure S1 & S2).
 
-![Relationship between Shannon Diversity and maximum patch size for varying dispersal distances.
-Simulations were conducted using a model with 60% habitat destruction and random block habitat patterns, where habitat blocks of the same size are randomly distributed over the simulation grid (i.e., the maximum patch size).](Figures/Shannon_MaxPatch_Dispersal_Lambda_block.png)
+![Relationship between number of species and habitat patch size for different versions of the model (across panels) and varying dispersal distances.
+The neutral is the clasical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
+Simulations were performed using a fixed growth rate (2), migration and a replacement rate of 0.0001 and 0.3 respectively.
+Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
+The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.](Figures/AcrossBehav_Species_Dispersal_Lambda2_regular.png)
+
+![Relationship between Shannon Diversity and habitat patch size for different versions of the model (across panels) and varying dispersal distances.
+The neutral is the clasical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
+Simulations were performed using a fixed growth rate (2), migration and a replacement rate of 0.0001 and 0.3 respectively.
+Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
+The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.](Figures/AcrossBehav_Shannon_Dispersal_Lambda2_regular.png)
 
@@
 
-Overall, our results confirm the previously found pattern and nearly all models retained more species under the SS rather than the SL scenario.
-In both cases, the loss of species is relevant when habitat loss takes place, and no significant extinction bias towards SS occurs as a long-term dynamic.
-The extinction of species in patches depends on the quotient between reproduction and extinction rates to patch size, i.e., the ability of a certain patch to be a source of propagules.
-The only case when SL > SS is when dispersal is not enough to connect patches, and size is smaller than the minimum to prevent patch extinction, as predicted elsewhere [@Rybicki2013].
-Letting aside those cases (i.e., not reaching minimum dispersal distance and intra-patch sink dynamics), our model was able to produce the patterns observed so far in field studies [@Fahrig2020].
-In that sense, we can hypothesize that our model variables (i.e., dispersal ability, patch size and lambda) may be biologically meaningful and enough to explain why SS retains more species than SL.
-Importantly, the spatial arrangement of available patches was not necessary to produce the spatial autocorrelation of occupied patches known to be responsible at least in part of the patterns observed in nature [@Fahrig2021].
-In contrast, high dispersal rates produce homogenization of species distribution and a reduction of species richness, especially if high dispersal species are also good competitors.
-In that scenario, SS and SL are approximately equivalent for species diversity retention.
-The model predicts that if a set of species have more dispersal capabilities the SS effect would be smaller than for species with less dispersal abilities.
+Our research contributes to the ongoing Single Large vs Several Small (SLOSS) debate in conservation biology, offering new insights based on our model findings.
+Consistent with previous studies, the results from our simulations primarily support the SS > SL pattern for species retention, as observed across various model scenarios [@Fahrig2020].
+This trend underlines the greater efficacy of multiple smaller habitats in preserving species diversity compared to a single large habitat under similar conditions of habitat loss.
+A critical factor in this dynamic is the balance between reproduction and extinction rates ($\lambda$ parameter) relative to the size of the habitat patches, influencing their role as sources of propagules.
+Notably, the SL > SS scenario emerges only when dispersal limitations prevent patch connectivity, and patch sizes are below a certain threshold to prevent local extinctions, aligning with Rybicki's findings [@Rybicki2013].
 
-Our model further illustrates that available patches are not always re-occupied following fragmentation even if the minimum size is met, therefore reinforcing the fact that dispersal is the main driver of the pattern, and that chance is involved in the final species pool when fragmentation takes place.
-Empirical data in large carnivores, for instance, support that several available patches are not occupied after isolation [@Paviolo2016].
+In neutral models one species by chance tends to dominate a patch [@Hubbell2001], so in fragmented landscapes and once the species increments its density is very unprobable that another species replaces it so with more patches the probability of having more species is augmented.
+This effect is enhanced when individuals can select the habitat, thus fragmentation have a positive effect in the number of species.
+For competitive communities, there is a higher probability that different species dominate different smaller patches, compared with a patch.
+Thus, again fragmentation has a positive effect on number of species.
+This positive effect is counteracted by disperal, high dispersal rates tended to lead to a homogenization of species distributions, so fragmented landscapes tend to be more like not fragmented ones reducing overall species richness.
+This pattern was particularly pronounced for species combining high dispersal capabilities with competitive strengths, indicating a complex interplay between dispersal, competitive interactions, and species diversity [@Fahrig2020].
 
-@Rybicki2013 suggested that both short-term and long-term extinctions are produced after fragmentation.
-Here, we observed that, in most cases, long-term loss of species in SS scenarios were less important than short-term extinctions in SL scenario.
-Overall, our results support the main conclusion by @Fahrig2021, in that the conditions for SL > SS are rather restricted and would not be observable in nature frequently (see @Fahrig2021, Fig 3).
-In contrast, if few conditions are met (minimum patch size and dispersal), SS would retain more species than SL for nearly all situations.
-Evidence suggests that small patches, even single structures such as isolated trees, can be important for biodiversity and landscape connectivity [@Shafer1995; @Manning2006].
-While connectivity management is case-specific, our results further support past claims [@Fahrig2020;@Fahrig2021], that assuming that SL is always better does not have empirical support.
-In our work, spreading of risk was not modelled (__maybe as time lag in at-patch competition?__), therefore reinforcing that the sampling effect is enough to produce an SS > SL pattern if dispersal is small enough to produce spatial autocorrelation of occupied patches [@Simberloff1976].
+The spatial arrangement of habitat patches, contrary to some expectations, did not produce different species richness patterns regarding fragmentation with few exceptions.
+Instead, the interplay between dispersal, patch size and between patch distance emerged as the primary factor influencing biodiversity patterns following habitat fragmentation, corroborated by empirical studies in large carnivores [@Paviolo2016].
+The exceptions could be important, because stressed communities---represented in our models with lowest $\lambdas$--- with low dispersal may have a continuous increment of richnes with patch size, observed only in random block degradation patterns.
+This could be related to the optimal patch size for random block degradations tends to be higher than for regular degradation patterns, so we could hypothesize that the optimal was outside the simulated range for these stressed communities.
 
-One large patch can enhance competitive exclusion, and that is more important the more similar species are. Hummingbirds?
+Our findings support Fahrig's [-@Fahrig2021] conclusion that conditions favoring SL > SS are limited and infrequently observed in nature [@Fahrig2021].
+Under conditions of adequate patch size and dispersal capabilities, the SS scenario typically retains more species.
+This underscores the ecological significance of smaller habitat patches, such as isolated trees, in maintaining biodiversity and landscape connectivity [@Shafer1995; @Manning2006].
+Our results thus challenge the conventional preference for larger reserves in conservation planning and suggest a more nuanced, context-specific approach.
+
+We also observed that the temporal dynamics of species extinction post-fragmentation vary, with long-term species loss in SS scenarios generally less severe than short-term extinctions in SL scenarios.
+This highlights the importance of considering both short-term and long-term impacts of habitat fragmentation on biodiversity.
+
+In conclusion, our study adds valuable perspectives to the SLOSS debate, illustrating the complexities of habitat fragmentation and species conservation.
+While our model has inherent limitations, it provides a useful framework for understanding key ecological processes.
+Future research should aim to extend these findings to more diverse ecological settings and incorporate additional factors influencing biodiversity in fragmented landscapes.
+
+<!-- Ver si se puede decir algo sobre las predicciones de la TABLA 1 de @Fahrig2021 -->
+
+<!-- One large patch can enhance competitive exclusion, and that is more important the more similar species are. Hummingbirds? -->
 
@@
 We acknowledge the limitations of our study and consider it a strength that we have identified in this initial model. This allows us to continue working on them in future research.
+
+In our work, spreading of risk was not modelled (__maybe as time lag in at-patch competition?__), therefore reinforcing that the sampling effect is enough to produce an SS > SL pattern if dispersal is small enough to produce spatial autocorrelation of occupied patches [@Simberloff1976].
 

### 2023-12-04 RunSims.Rmd
Updated manuscript

@@
 ```{r PlotsNoSelectionRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+# read data
+#
+sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
 
@@
   theme(
-    legend.position = c(0.9,0.8)
+    legend.position = c(0.9,0.2)
     )
@@
 mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,deforestation_type) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+names(mdl1)
+
+mdl1 <- mdl1 %>%  mutate(across(max_patch_mean:max_patch_qh,sqrt))
+
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = max_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, deforestation_type), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_wrap(~ deforestation_type, ncol = 2) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Max Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/AcrossLoss_Species_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
+
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = max_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, deforestation_type), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_wrap(~ deforestation_type, ncol = 4) + 
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Max Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = c(0.9,0.8)
+    )
+
+ggsave("Figures/AcrossLoss_Shannon_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
+
+#
+# summarize data mean patch size
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
                          group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,deforestation_type) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
@@
   labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
-  theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/AcrossLoss_Species_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
-
-# Shannon-Diversity  
-#
-mdl1 %>%  filter(birth_rate_birds == 2) %>% 
-  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
-  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
-  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, deforestation_type), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ deforestation_type, ncol = 2) + 
-  labs(color = "Dispersal\n distance") +
   theme_bw() + xlab("Mean Habitat Patch Size") + ylab("Number of species") +
@@
 
-ggsave("Figures/AcrossLoss_Species_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
+ggsave("Figures/AcrossLoss_Species_Dispersal_Lambda2_mean.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
 
@@
 
-ggsave("Figures/AcrossLoss_Shannon_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
+ggsave("Figures/AcrossLoss_Shannon_Dispersal_Lambda2_mean.png",width=8,height=6,units="in",dpi=600)
 

### 2023-12-04 ms_SLOSS.Rmd
Updated manuscript

@@
 
-We considered only one amount of habitat destruction in combination with different community dynamics, neutral and competitive, taking into account healthy and mpoverished communities with different dispersal related traits, and different patterns of habitat loss.
+We considered only one amount of habitat destruction in combination with different community dynamics, neutral and competitive, taking into account healthy and empoverished communities with different dispersal related traits, and different patterns of habitat loss.
 
@@
 For the neutral classical version only, we also simulate random block habitat patterns.
+For each combination of parameters we run 30 simulations up to 1500 time steps and recorded the number of species, the Shannon Diversity index [@Magurran2013] and the total number of individuals.
+For some representative combiations of parameters we take snapshots of the model output showing the typical dynamics on time of the 3 recorded quantities and the rank abundance species distribution [@Newman2005; @May1975]
 
@@
 
-The Netlogo model source code is available at Github <https://github.com/lsaravia/DynamicForestExtinction> and the R source code to perform simulations and analysis is located at <https://github.com/EcoComplex/multiSLOSS>.
+The model was developed in Netlogo 6.3.0 and model source code is available at Github <https://github.com/lsaravia/DynamicForestExtinction>, the Overview, Design concepts, Details (ODD), protocol for describing individual-based models is include within the model source code [@Grimm2010].
+We performed the simulations and analisis usin R version 4.3.1 [@RCoreTeam2023], the R source code to perform simulations and analysis is located at <https://github.com/EcoComplex/multiSLOSS>.
 
@@
 The interplay of dispersal distances, habitat sizes, and \(\lambda\) impacts community survival and species richness.
-This suggests a threshold patch size influenced by these 3 parameters.
-Extended dispersal distances enhance species survival in scenarios with low \(\lambda\) and smaller patches but reduce richness beyond a certain threshold.
+For $\lambda < 4$ this suggests a threshold patch size influenced by these 3 parameters, for $\lambda = 4$ patch size decrease species richness.
+The differences between increasing $\lambda$'s almost dissapear when dispersal distance is higher.
+Extended dispersal distances could enhance species survival in scenarios with low \(\lambda\) and smaller patches but reduce richness beyond a certain threshold.
 Similar patterns are evident for Shannon diversity (Figure 2).
@@
 
-For the simulations made with random block habitat patterns, the patches do not have the same sizes but there is a distribution with a minimum size determined by the patch-size parameter and a bigger maximum size.
-Again there is a simimlar pattern with a maximum richness after the threshold for $\lambda$ values less than 3 and for values greater or equal the maximum is at the lower habitat patch size (Figure S1 & S2).
+For the simulations made with random block habitat patterns, the patches do not have the same sizes but there is a distribution with a minimum size determined by the patch-size parameter and a bigger maximum size (Figures S9 - S11).
+The pattern is similar to the regular case,  with a maximum richness for intermediate patch sizes and for values of $\lambda$ higher than 2 the maximum is at the lower habitat patch size (Figure S1 & S2).
+The sizes of patches for the random block simulations are bigger so species richness and diversity are lower, except for the minimum patch size (Figure S7 & S8).
+Examples of simulations with random block and regular habitat patterns are shown in the supplementary material (Figures S9 & S32).
+In these figures we can observe that for the minimum patch size only migrants survive for regular habitat patterns with neutral and hierarchical models.
 
 ![Relationship between number of species and habitat patch size for different versions of the model (across panels) and varying dispersal distances.
-The neutral is the clasical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
+The neutral is the clasical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat-selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
 Simulations were performed using a fixed growth rate (2), migration and a replacement rate of 0.0001 and 0.3 respectively.
@@
 ![Relationship between Shannon Diversity and habitat patch size for different versions of the model (across panels) and varying dispersal distances.
-The neutral is the clasical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
+The neutral is the clasical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat-selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
 Simulations were performed using a fixed growth rate (2), migration and a replacement rate of 0.0001 and 0.3 respectively.
@@
 The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.](Figures/AcrossBehav_Shannon_Dispersal_Lambda2_regular.png)
+
+Comparing the three versions of the model at $\lambda =2$, for higher dispersal the neutral model tend to have higher richness and Shannon diversity at higher patch sizes, followed by the hierarchical model and the neutral-with-habitat-selection model (Figure 3 & 4).
+At the lowest patch size, only in the neutral-with-habitat-selection have populations that can survive, because in the other model versions the individuals observed are migrants that cannot establish a local populations an become extinct (See Figures S5 - S15).
+Thus an intermediate patch size with a maximum species richness is not observed for this model.
+
+Examples of the final pattern of species and habitat distributions is shown in Figures S9- S32.
+Species spatial distributions closely match the spatial distribution previous to the habitat destruction for neutral and neutral-with-habitat-selection models.
+Is also clearly observed that for the hierarchical model there is a strong dominance of the more competitive species where Shannon Diversity is very low for higher dispersal rate (Figure 4 & S23).
 
@@
 This trend underlines the greater efficacy of multiple smaller habitats in preserving species diversity compared to a single large habitat under similar conditions of habitat loss.
-A critical factor in this dynamic is the balance between reproduction and extinction rates ($\lambda$ parameter) relative to the size of the habitat patches, influencing their role as sources of propagules.
-Notably, the SL > SS scenario emerges only when dispersal limitations prevent patch connectivity, and patch sizes are below a certain threshold to prevent local extinctions, aligning with Rybicki's findings [@Rybicki2013].
+A critical factor in this dynamic is the balance between reproduction and extinction rates ($\lambda$ parameter) and dispersal relative to the size of the habitat patches, influencing their role as sources of propagules.
+Notably, the SL > SS scenario emerges only when dispersal limitations prevent patch connectivity, and patch sizes are above a certain threshold to prevent local extinctions, aligning with Rybicki's [-@Rybicki2020] findings.
 
 In neutral models one species by chance tends to dominate a patch [@Hubbell2001], so in fragmented landscapes and once the species increments its density is very unprobable that another species replaces it so with more patches the probability of having more species is augmented.
-This effect is enhanced when individuals can select the habitat, thus fragmentation have a positive effect in the number of species.
-For competitive communities, there is a higher probability that different species dominate different smaller patches, compared with a patch.
-Thus, again fragmentation has a positive effect on number of species.
-This positive effect is counteracted by disperal, high dispersal rates tended to lead to a homogenization of species distributions, so fragmented landscapes tend to be more like not fragmented ones reducing overall species richness.
+This effect is enhanced when individuals can select the habitat, because of the increased survival of the propagules, thus fragmentation have a positive effect in the number of species.
+For competitive communities, there is a higher probability that different species dominate different smaller patches, compared with a bigger patch.
+Thus, again fragmentation has a positive effect on number of species, this effect was called stochastic divergence [@Riva2023].
+In smaller patches there is an increase of extinction risk because they support smaller populations [@Laurance2002;@Shafer1995], however as patches become smaller they also become more numerous, reducing the probability of extinction over the entire set of patches due to spreading the risk effect [@Crowley1977].
+The combination of these two effects, the stochastic divergence and the risk effect, explain the highest number of species with patch size over the threshold of species survival.
+
+This positive effect is counteracted by dispersal, high dispersal rates tended to a homogenization of species distributions, so fragmented landscapes tend to be more like not fragmented ones reducing overall species richness.
 This pattern was particularly pronounced for species combining high dispersal capabilities with competitive strengths, indicating a complex interplay between dispersal, competitive interactions, and species diversity [@Fahrig2020].
 
-The spatial arrangement of habitat patches, contrary to some expectations, did not produce different species richness patterns regarding fragmentation with few exceptions.
-Instead, the interplay between dispersal, patch size and between patch distance emerged as the primary factor influencing biodiversity patterns following habitat fragmentation, corroborated by empirical studies in large carnivores [@Paviolo2016].
-The exceptions could be important, because stressed communities---represented in our models with lowest $\lambdas$--- with low dispersal may have a continuous increment of richnes with patch size, observed only in random block degradation patterns.
-This could be related to the optimal patch size for random block degradations tends to be higher than for regular degradation patterns, so we could hypothesize that the optimal was outside the simulated range for these stressed communities.
+Edge effects are present in the neutral and hierarchical models, but not in the neutral-with-habitat-selection model.
+In the first two models, the edge effect is due to the fact that individuals in the border of the patch have less neighbors and the probability of extinction is higher.
+In the neutral-with-habitat-selection model, the edge effect is not present because the individuals search for empty patches, so the individuals in the border of the patch have the same probability of extinction that the individuals in the interior of the patch.
+Thus when patches are smaller the edge proportion is higher and the edge effect is more important, but this plays a more important role for the minimal patch size for survival of species than for the species richness.
 
-Our findings support Fahrig's [-@Fahrig2021] conclusion that conditions favoring SL > SS are limited and infrequently observed in nature [@Fahrig2021].
+The spatial arrangement of habitat patches, contrary to some expectations, did not produce different species richness patterns regarding habitat patch size.
+The exceptions are stressed communities---represented in our models with lowest $\lambda$--- with low dispersal may have a continuous increment of richness with patch size, observed only in random block degradation patterns.
+This could be related to the optimal patch size for random block habitat loss patterns tends to be higher than for regular patterns, so we could hypothesize that the optimal was outside the simulated range for these stressed communities.
+The aggregation of patches in the random block pattern (implies aggregation of the degraded habitat) could be positive if individual habitat patches are very small, because it allows some patches to surpass the minimum size required for survival.
+For bigger patch sizes it could be detrimental because the aggregation produces even bigger patch sizes that decrease species richness.
+A similar effect was observed on experimental landscapes [@Fletcher2023], aggregated patterns of habitat loss are detrimental for patch occupancy and eventually could produce the same effects observed in our models about richness.
+
+Naturally, there are limitations to our modeling work.
+Firstly, it is important to note that more research is needed to apply our to determine the effect of fragmentation in real-world communities.
+We assume that all species have the same parameters, thus hypothetically determining these parameters for all the species, and using the mean to determine the responses to fragmentation would render an approximate value for the effects and thresholds.
+Also an improved version of the model with different values of the parameters for each species could be developed and compared with the neutral model.
+The effects of fragmentation could be influenced by many other ecological/behavioral characteristics of the species not included in the model, environmental heterogeneity, also the size of the focal landscape compared to dispersal distances, that determines the scale of the dynamics could influence quantitative results [@Rybicki2020].
+Besides that, the approach of using mean values of the parameters of relatively simple multispecies models has been applied and tested to predict species biomass [@Fort2018].
+
+In our simulations we fixed the amount of habitat loss to 60%, first it in previous simulation studies it was shown that  fragmentation was important when the habitat amount is low, around 10% [@Hanski2015; @Rybicki2013].
+but it has been observed that the amount of habitat loss could influence the effect of fragmentation [@Loke2019; @With2016].
+It was suggested that with very low habitat amount (e.g 7%) the arrangement of patches does not matters, this could paired with our results of stressed communities
+
+Our findings support Fahrig's [-@Fahrig2021] conclusion that conditions favoring SL > SS are limited and infrequently observed in nature.
 Under conditions of adequate patch size and dispersal capabilities, the SS scenario typically retains more species.
-This underscores the ecological significance of smaller habitat patches, such as isolated trees, in maintaining biodiversity and landscape connectivity [@Shafer1995; @Manning2006].
-Our results thus challenge the conventional preference for larger reserves in conservation planning and suggest a more nuanced, context-specific approach.
-
-We also observed that the temporal dynamics of species extinction post-fragmentation vary, with long-term species loss in SS scenarios generally less severe than short-term extinctions in SL scenarios.
-This highlights the importance of considering both short-term and long-term impacts of habitat fragmentation on biodiversity.
+This underscores the ecological significance of smaller habitat patches, such as isolated trees, in maintaining biodiversity and landscape connectivity [@Yan2021; @Manning2006; @Shafer1995].
+Our results thus challenge the conventional preference for larger reserves in conservation planning and support the paradigm shift in conservation towards considering sets of many small patches generally having  higher biodiversity than sets of few large patches [@Riva2023].
 
@@
 Future research should aim to extend these findings to more diverse ecological settings and incorporate additional factors influencing biodiversity in fragmented landscapes.
+
+<!-- We also observed without a systematic exploration the temporal dynamics of species extinction, short-term extinctions are more severe than long-term species loss (extinction debt), but extinctions can continue several years.
+The importance of long term species loss is lower in our model than in similar neutral models [@Thompson2019] This highlights the importance of considering both short-term and long-term impacts of habitat fragmentation on biodiversity.
+-->
 

### 2023-12-04 ms_SLOSS.Rmd
Updated conclusions

@@
 
-In our simulations we fixed the amount of habitat loss to 60%, first it in previous simulation studies it was shown that  fragmentation was important when the habitat amount is low, around 10% [@Hanski2015; @Rybicki2013].
-but it has been observed that the amount of habitat loss could influence the effect of fragmentation [@Loke2019; @With2016].
-It was suggested that with very low habitat amount (e.g 7%) the arrangement of patches does not matters, this could paired with our results of stressed communities
+In previous simulation studies it was shown that  fragmentation was important when the habitat amount is low, around 10% [@Hanski2015; @Rybicki2013], but it has been observed that with levels habitat around 10-20% the effect of fragmentation is positive or have an intermediate maximum [@Loke2019; @With2016], and with lower habitat amount (e.g 7%) the arrangement of patches does not matters, we obtained similar patterns with 40% habitat amount, which is near the amount used in newer simulated studies (32%) that also obtain more species richness with more fragmentation [@Rybicki2020].
+
+The novel aspect of our model is that it consider communities that are assembled from a metacommunity in which individuals migrate to empty habitat, and then after the number of species is stabilized the habitat is fragmented.
+This could be a more realistic scenario than if all species are present with initially random patterns at an already fragmented habitat [@Rybicki2020].
+This could be more important the more heterogeneous was species spatial distribution pattern before fragmentation.
 

### 2023-12-04 ms_SLOSS.Rmd
More languaje updates to the ms

--- 2023-12-04 ms_SLOSS.Rmd
@@
 The extinction of species in patches depends on the quotient between reproduction and extinction rates in relation to patch size.
-The only case when SL > SS is when dispersal is not enough to connect patches, and patch size is smaller than the minimum to prevent extinction.
+The only case when SL > SS is when dispersal is insufficient to connect patches, and patch size is smaller than the minimum to prevent extinction.
 If patch size is above the critical size to maintain local populations the SS > SL pattern is due to the sampling effect of aggregated species distribution combined with the reduction of dispersal ability of species.
@@
 Moreover, habitat fragmentation occurs at the same time as habitat loss, and the effects of both processes are difficult to disentangle [@Haddad2015].
-Landscape connectivity and organism mobility are a critical feature for population persistence [@Taylor1993].
+Landscape connectivity and organism mobility are critical features for population persistence [@Taylor1993].
 Starting from a landmark paper [@Diamond1975], conservation biologists engaged in the challenging task of recommending protection measures given the environmental crisis coupled with a lack of knowledge of most of the world’s biodiversity.
@@
 The actual processes driving to it remain elusive and it is probably multi-causal (quotes).
-In that scenario, model simulations could produce useful predictions targeting the relevant variables in order to better understand the processes leading to non-intuitive patterns.
+In that scenario, model simulations could produce useful predictions targeting the relevant variables to better understand the processes leading to non-intuitive patterns.
 
@@
 
-We considered only one amount of habitat destruction in combination with different community dynamics, neutral and competitive, taking into account healthy and empoverished communities with different dispersal related traits, and different patterns of habitat loss.
+We considered only one amount of habitat destruction in combination with different community dynamics, neutral and competitive, taking into account healthy and impoverished communities with different dispersal related traits, and different patterns of habitat loss.
 
-The aim of this work is to explore spatial multi-species models to verify the SS > SL pattern and determine the causes of it.
+This work aims to explore spatial multi-species models to verify the SS > SL pattern and determine the causes of it.
 In particular, we simulate species retention after different fragmentation scenarios of the same overall habitat loss, therefore accounting for the interacting effects of habitat loss and fragmentation.
@@
 As in a classical neutral model, there is a metacommunity, i.e.
-a regional species pool from were species migrate [@Hubbell2001].
+a regional species pool from where species migrate [@Hubbell2001].
 All individuals have the same parameters and each species is assigned an indicator number.
@@
 
-The neutral variation with habitat selection where individuals have the ability to actively select suitable habitat patches.
-In this variation, propagules have the capacity to choose unoccupied patches, ensuring their survival and successful establishment.
+The neutral variation with habitat selection where individuals can actively select suitable habitat patches.
+In this variation, propagules can choose unoccupied patches, ensuring their survival and successful establishment.
 This modification reflects the influence of habitat availability and selection on species colonization and persistence.
@@
 The competitive hierarchical variation extends the neutral variation with an additional layer of complexity by incorporating competitive interactions among species.
-Each species have an indicator number that determines its competitive dominance level, and more dominant species have the ability to replace less dominant ones at a specified replacement rate.
+Each species has an indicator number that determines its competitive dominance level, and more dominant species can replace less dominant ones at a specified replacement rate.
 This variation captures the influence of species interactions and competitive exclusion on species composition and abundance.
@@
 
-There following process are included in the model: habitat loss, migration, reproduction, local dispersal, competition and death. After setting initial conditions the following events can happen:
+The following processes are included in the model: habitat loss, migration, reproduction, local dispersal, competition and death. After setting initial conditions the following events can happen:
 
@@
 
-    Regular degradation: Habitat is removed so as habitable patches form a regular pattern of squares with side given by an habitat-size parameter.
-As the percentage of degraded habitat is constant varying the habitat-size parameter results in either few large habitat patches (SL) or several small (SS).
+    Regular degradation: Habitat is removed so that habitable patches form a regular pattern of squares with side given by an habitat-size parameter.
+As the percentage of degraded habitat is constant varying the habitat-size parameter results in either a few large habitat patches (SL) or several small (SS).
 
     Random block degradation: Habitat square patches of a fixed size (habitat-size parameter) are located at random throughout the landscape.
-The sizes are identical to those used for regular degradation simulations, but it results in habitat patches that have a minimum size corresponding to the habitat-size  but a maximum size that is bigger, because habitat patches can superpose between them producing a larger patch.
+The sizes are identical to those used for regular degradation simulations, but it results in habitat patches that have a minimum size corresponding to the habitat-size  but a maximum size that is bigger because habitat patches can superpose between them producing a larger patch.
 Then, I characterize this with the maximum habitat patch size.
@@
 
-    What happens to the propagule when arrives to the selected site is determined by the variation of the model simulated:
+    What happens to the propagule when arrives at the selected site is determined by the variation of the model simulated:
 
@@
 Competitive Hierarchical: Species are ranked by a competitive dominance hierarchy based on their internal identification   (id) number.
-If a dispersing propagule arrives at an occupied site, it will only establish if its id number is higher than the resident species.
+If a dispersing propagule arrives at an occupied site, it will only be established if its id number is higher than the resident species.
 Otherwise, the propagule perishes without establishing.
@@
 For the neutral classical version only, we also simulate random block habitat patterns.
-For each combination of parameters we run 30 simulations up to 1500 time steps and recorded the number of species, the Shannon Diversity index [@Magurran2013] and the total number of individuals.
-For some representative combiations of parameters we take snapshots of the model output showing the typical dynamics on time of the 3 recorded quantities and the rank abundance species distribution [@Newman2005; @May1975]
+For each combination of parameters we ran 30 simulations up to 1500 time steps and recorded the number of species, the Shannon Diversity index [@Magurran2013] and the total number of individuals.
+For some representative combinations of parameters we take snapshots of the model output showing the typical dynamics on time of the 3 recorded quantities and the rank abundance species distribution [@Newman2005; @May1975]
 
@@
 The classical neutral model and the habitat selection model could represent endpoints on a continuum of models where the matrix is either completely adverse or not adverse but still unsuitable for reproduction.
-In the habitat selection model, individuals can search across both the matrix and suitable habitat until finding an empty place.
-In contrast, for the classical neutral model, individuals have limited or no capacity to search, and the matrix is necessarily detrimental because it increases the probability of death by landing on an unsitable site.
+In the habitat selection model, individuals can search across both the matrix and suitable habitat until they find an empty place.
+In contrast, for the classical neutral model, individuals have limited or no capacity to search, and the matrix is necessarily detrimental because it increases the probability of death by landing on an unsuitable site.
 
-The model was developed in Netlogo 6.3.0 and model source code is available at Github <https://github.com/lsaravia/DynamicForestExtinction>, the Overview, Design concepts, Details (ODD), protocol for describing individual-based models is include within the model source code [@Grimm2010].
-We performed the simulations and analisis usin R version 4.3.1 [@RCoreTeam2023], the R source code to perform simulations and analysis is located at <https://github.com/EcoComplex/multiSLOSS>.
+The model was developed in Netlogo 6.3.0 and model source code is available at Github <https://github.com/lsaravia/DynamicForestExtinction>, the Overview, Design concepts, Details (ODD), protocol for describing individual-based models is included within the model source code [@Grimm2010].
+We performed the simulations and analysis using R version 4.3.1 [@RCoreTeam2023], the R source code to perform simulations and analysis is located at <https://github.com/EcoComplex/multiSLOSS>.
 
@@
 The interplay of dispersal distances, habitat sizes, and \(\lambda\) impacts community survival and species richness.
-For $\lambda < 4$ this suggests a threshold patch size influenced by these 3 parameters, for $\lambda = 4$ patch size decrease species richness.
-The differences between increasing $\lambda$'s almost dissapear when dispersal distance is higher.
+For $\lambda < 4$ this suggests a threshold patch size influenced by these 3 parameters, for $\lambda = 4$ patch size decreases species richness.
+The differences between increasing $\lambda$'s almost disappear when dispersal distance is higher.
 Extended dispersal distances could enhance species survival in scenarios with low \(\lambda\) and smaller patches but reduce richness beyond a certain threshold.
@@
 
-![Relationship between number of species and habitat patch size for varying dispersal distances and $\lambda$ parameter (across panels).
+![Relationship between the number of species and habitat patch size for varying dispersal distances and $\lambda$ parameter (across panels).
 Simulations were performed using a neutral model with a constant migration rate (0.0001).
@@
 For the simulations made with random block habitat patterns, the patches do not have the same sizes but there is a distribution with a minimum size determined by the patch-size parameter and a bigger maximum size (Figures S9 - S11).
-The pattern is similar to the regular case,  with a maximum richness for intermediate patch sizes and for values of $\lambda$ higher than 2 the maximum is at the lower habitat patch size (Figure S1 & S2).
+The pattern is similar to the regular case,  with a maximum richness for intermediate patch sizes and values of $\lambda$ higher than 2 the maximum is at the lower habitat patch size (Figure S1 & S2).
 The sizes of patches for the random block simulations are bigger so species richness and diversity are lower, except for the minimum patch size (Figure S7 & S8).
@@
 
-![Relationship between number of species and habitat patch size for different versions of the model (across panels) and varying dispersal distances.
-The neutral is the clasical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat-selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
-Simulations were performed using a fixed growth rate (2), migration and a replacement rate of 0.0001 and 0.3 respectively.
+![Relationship between the number of species and habitat patch size for different versions of the model (across panels) and varying dispersal distances.
+The neutral is the classical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat-selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
+Simulations were performed using a fixed growth rate (2), migration and replacement rate of 0.0001 and 0.3 respectively.
 Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
@@
 ![Relationship between Shannon Diversity and habitat patch size for different versions of the model (across panels) and varying dispersal distances.
-The neutral is the clasical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat-selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
-Simulations were performed using a fixed growth rate (2), migration and a replacement rate of 0.0001 and 0.3 respectively.
+The neutral is the classical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat-selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
+Simulations were performed using a fixed growth rate (2), migration and replacement rate of 0.0001 and 0.3 respectively.
 Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
@@
 Comparing the three versions of the model at $\lambda =2$, for higher dispersal the neutral model tend to have higher richness and Shannon diversity at higher patch sizes, followed by the hierarchical model and the neutral-with-habitat-selection model (Figure 3 & 4).
-At the lowest patch size, only in the neutral-with-habitat-selection have populations that can survive, because in the other model versions the individuals observed are migrants that cannot establish a local populations an become extinct (See Figures S5 - S15).
+At the lowest patch size, only in the neutral-with-habitat-selection have populations that can survive, because in the other model versions the individuals observed are migrants that cannot establish local populations and become extinct (See Figures S5 - S15).
 Thus an intermediate patch size with a maximum species richness is not observed for this model.
@@
 
-In neutral models one species by chance tends to dominate a patch [@Hubbell2001], so in fragmented landscapes and once the species increments its density is very unprobable that another species replaces it so with more patches the probability of having more species is augmented.
-This effect is enhanced when individuals can select the habitat, because of the increased survival of the propagules, thus fragmentation have a positive effect in the number of species.
+In neutral models one species by chance tends to dominate a patch [@Hubbell2001], so in fragmented landscapes, once the species increments its density is very improbable that another species replaces it so with more patches the probability of having more species is augmented.
+This effect is enhanced when individuals can select the habitat, because of the increased survival of the propagules, thus fragmentation has a positive effect on the number of species.
 For competitive communities, there is a higher probability that different species dominate different smaller patches, compared with a bigger patch.
-Thus, again fragmentation has a positive effect on number of species, this effect was called stochastic divergence [@Riva2023].
-In smaller patches there is an increase of extinction risk because they support smaller populations [@Laurance2002;@Shafer1995], however as patches become smaller they also become more numerous, reducing the probability of extinction over the entire set of patches due to spreading the risk effect [@Crowley1977].
+Thus, again fragmentation has a positive effect on the number of species, this effect is called stochastic divergence [@Riva2023].
+In smaller patches there is an increase in extinction risk because they support smaller populations [@Laurance2002;@Shafer1995], however as patches become smaller they also become more numerous, reducing the probability of extinction over the entire set of patches due to spreading the risk effect [@Crowley1977].
 The combination of these two effects, the stochastic divergence and the risk effect, explain the highest number of species with patch size over the threshold of species survival.
 
-This positive effect is counteracted by dispersal, high dispersal rates tended to a homogenization of species distributions, so fragmented landscapes tend to be more like not fragmented ones reducing overall species richness.
+This positive effect is counteracted by dispersal, high dispersal rates tend to a homogenization of species distributions, so fragmented landscapes tend to be more like not fragmented ones reducing overall species richness.
 This pattern was particularly pronounced for species combining high dispersal capabilities with competitive strengths, indicating a complex interplay between dispersal, competitive interactions, and species diversity [@Fahrig2020].
@@
 Edge effects are present in the neutral and hierarchical models, but not in the neutral-with-habitat-selection model.
-In the first two models, the edge effect is due to the fact that individuals in the border of the patch have less neighbors and the probability of extinction is higher.
-In the neutral-with-habitat-selection model, the edge effect is not present because the individuals search for empty patches, so the individuals in the border of the patch have the same probability of extinction that the individuals in the interior of the patch.
-Thus when patches are smaller the edge proportion is higher and the edge effect is more important, but this plays a more important role for the minimal patch size for survival of species than for the species richness.
+In the first two models, the edge effect is because individuals in the patch's border have fewer neighbors and the probability of extinction is higher.
+In the neutral-with-habitat-selection model, the edge effect is not present because the individuals search for empty patches, so the individuals on the border of the patch have the same probability of extinction as the individuals in the interior of the patch.
+Thus when patches are smaller the edge proportion is higher and the edge effect is more important, but this plays a more important role in the minimal patch size for survival of species than for the species richness.
 
@@
 This could be related to the optimal patch size for random block habitat loss patterns tends to be higher than for regular patterns, so we could hypothesize that the optimal was outside the simulated range for these stressed communities.
-The aggregation of patches in the random block pattern (implies aggregation of the degraded habitat) could be positive if individual habitat patches are very small, because it allows some patches to surpass the minimum size required for survival.
+The aggregation of patches in the random block pattern (implies aggregation of the degraded habitat) could be positive if individual habitat patches are very small because it allows some patches to surpass the minimum size required for survival.
 For bigger patch sizes it could be detrimental because the aggregation produces even bigger patch sizes that decrease species richness.
-A similar effect was observed on experimental landscapes [@Fletcher2023], aggregated patterns of habitat loss are detrimental for patch occupancy and eventually could produce the same effects observed in our models about richness.
+A similar effect was observed on experimental landscapes [@Fletcher2023], aggregated patterns of habitat loss are detrimental to patch occupancy and eventually could produce the same effects observed in our models about richness.
 
@@
 We assume that all species have the same parameters, thus hypothetically determining these parameters for all the species, and using the mean to determine the responses to fragmentation would render an approximate value for the effects and thresholds.
-Also an improved version of the model with different values of the parameters for each species could be developed and compared with the neutral model.
-The effects of fragmentation could be influenced by many other ecological/behavioral characteristics of the species not included in the model, environmental heterogeneity, also the size of the focal landscape compared to dispersal distances, that determines the scale of the dynamics could influence quantitative results [@Rybicki2020].
+Also, an improved version of the model with different values of the parameters for each species could be developed and compared with the neutral model.
+The effects of fragmentation could be influenced by many other ecological/behavioural characteristics of the species not included in the model, environmental heterogeneity, also the size of the focal landscape compared to dispersal distances, which determines the scale of the dynamics could influence quantitative results [@Rybicki2020].
 Besides that, the approach of using mean values of the parameters of relatively simple multispecies models has been applied and tested to predict species biomass [@Fort2018].
 
-In our simulations we fixed the amount of habitat loss to 60%, first it in previous simulation studies it was shown that  fragmentation was important when the habitat amount is low, around 10% [@Hanski2015; @Rybicki2013].
-but it has been observed that the amount of habitat loss could influence the effect of fragmentation [@Loke2019; @With2016].
-It was suggested that with very low habitat amount (e.g 7%) the arrangement of patches does not matters, this could paired with our results of stressed communities
+In previous simulation studies, it was shown that fragmentation was important when the habitat amount is low, around 10% [@Hanski2015; @Rybicki2013], but it has been observed that with levels habitat around 10-20% the effect of fragmentation is positive or have an intermediate maximum [@Loke2019; @With2016], and with lower habitat amount (e.g 7%) the arrangement of patches does not matters, we obtained similar patterns with 40% habitat amount, which is near the amount used in newer simulated studies (32%) that also obtain more species richness with more fragmentation [@Rybicki2020].
+
+The novel aspect of our model is that it considers communities that are assembled from a metacommunity in which individuals migrate to an empty habitat, and then after the number of species is stabilized the habitat is fragmented.
+This could be a more realistic scenario than if all species are present with initially random patterns at an already fragmented habitat [@Rybicki2020].
+This could be more important the more heterogeneous was species spatial distribution pattern before fragmentation.
 
@@
 
-<!-- We also observed without a systematic exploration the temporal dynamics of species extinction, short-term extinctions are more severe than long-term species loss (extinction debt), but extinctions can continue several years.
-The importance of long term species loss is lower in our model than in similar neutral models [@Thompson2019] This highlights the importance of considering both short-term and long-term impacts of habitat fragmentation on biodiversity.
--->
-
-<!-- Ver si se puede decir algo sobre las predicciones de la TABLA 1 de @Fahrig2021 -->
-
-<!-- One large patch can enhance competitive exclusion, and that is more important the more similar species are. Hummingbirds? -->
-
-### Limitations
-
-We acknowledge the limitations of our study and consider it a strength that we have identified in this initial model. This allows us to continue working on them in future research.
-
-In our work, spreading of risk was not modelled (__maybe as time lag in at-patch competition?__), therefore reinforcing that the sampling effect is enough to produce an SS > SL pattern if dispersal is small enough to produce spatial autocorrelation of occupied patches [@Simberloff1976].
-
-Matrix is assumed as non-habitat
-
-Genetic structured not considered
-
-We are not modeling total diversity, rather diversity of a certain group with a set of similar species
-
-### Conclusions
-
 ## Acknowledgments
@@
 ## References
-
-<!--
-
-[a]: [UB1]Me parece que está bien el foco en esto, así no nos critican con boludeces como “ahh pero no tiene en cuenta que el pato en peligro puede en realidad necesitar…” tenemos que tratar de explicar por qué Fahrig encuentra lo que encuentra
-
-ej Fletcher (2018): “Approximately one quarter (24.4%) of Fahrig's results focused on species richness, with little consideration of species identity.
-Species richness is useful for summarizing ecological patterns but can mask compositional changes that are highly relevant to conservation”
-
-“the review had a narrow focus on a habitat vs non-habitat dichotomy, assuming the functionally-relevant habitat was appro[1]priately quantified and of comparatively similar quality across the landscape for the responses considered.
-For many species, the non-ha[1]bitat matrix may also provide resources and generic habitat delineation can obscure variation in habitat fragmentation effects”
-
-[b]: Si explicitamos esas limitaciones, estamos cubiertos me parece quiero decir que somos conscientes de las limitaciones, y es una fortaleza que las hayamos identificado en este primer modelo para seguir trabajando luego en el futuro
-
--->

### 2024-02-02 RunSims.Rmd
Update to figures and ms

@@
   labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
   theme(
@@
 names(mdl1)
+behavior_order <- c("NoSelection", "Hierarchical", "BirthSelection")
+behavior_labels <- c("Neutral", "Hierarchical", "Habitat Selection")
+
+mdl1 <- mdl1 %>% mutate(birds_behavior= factor(birds_behavior, levels = behavior_order, labels = behavior_labels))
 
@@
   theme(
-    legend.position = c(0.9,0.8)
+    legend.position = c(0.85,0.18)
     )

### 2024-02-02 ms_SLOSS.Rmd
More languaje updates to the ms
Update to figures and ms

@@
 As global biodiversity is rapidly declining due to habitat loss, it is important to determine how to protect it.
-Still, the policies of many conservation agencies are prioritizing the preservation of a single large habitat area (SL) vs.
-several small areas of the same size (SS), besides empirical support favouring SS.
-However, to date, most of the theories regarding this issue were based on single-species models.
+The policies of many conservation agencies are still prioritizing the preservation of a single large habitat area (SL) vs.
+several small areas of the same size (SS), despite empirical support favouring SS.
+However, to date, most theories regarding this issue were based on single-species models.
 This work aims to explore spatial multi-species models to verify the SS > SL pattern and determine the causes of it.
 We used three types of models: a classical neutral, neutral with habitat selection, and a competitive hierarchical model.
-All models have three parameters, extinction, reproduction, and dispersal distance.
-In the classical neutral model, species are equivalent and have the same parameters, when there is a reproduction event they send the propagule to a surrounding patch blindly, if the patch is already occupied the propagule dies.
+All models have three parameters, namely extinction, reproduction, and dispersal distance.
+In the classical neutral model, species are equivalent and have the same parameters; when there is a reproduction event they send the propagule to a surrounding patch blindly, and if the patch is already occupied the propagule dies.
 In the neutral-with-habitat selection model, the propagules select an empty patch so their survival is ensured.
@@
 
-Global biodiversity is rapidly declining due to habitat loss [@Young2016;@Chase2020;@Newbold2015].
-Moreover, habitat fragmentation occurs at the same time as habitat loss, and the effects of both processes are difficult to disentangle [@Haddad2015].
+Global biodiversity is rapidly declining due to, among other factors, habitat loss [@Young2016;@Chase2020;@Newbold2015].
+As habitat fragmentation occurs at the same time as habitat loss, the effects of both processes are difficult to disentangle [@Haddad2015].
 Landscape connectivity and organism mobility are critical features for population persistence [@Taylor1993].
-Starting from a landmark paper [@Diamond1975], conservation biologists engaged in the challenging task of recommending protection measures given the environmental crisis coupled with a lack of knowledge of most of the world’s biodiversity.
-Early recommendations claimed to follow ecological theory, in particular island biogeography theory to derive simple thumb rules decisions to better design protected areas.
-However, as noted rapidly, species/area relationships do not have a clear association with fragmentation per se but only with overall habitat loss [@Simberloff1982].
-Aiming to design better management practices, this has led to a nearly 50-year Single Large vs Several Small (SLOSS) debate.
+Starting from a landmark paper [@Diamond1975], conservation biologists engaged in the challenging task of recommending protection measures given the environmental crisis coupled with a lack of knowledge of most of the world’s biodiversity and limited funding for conservation.
+In that context, early recommendations claiming to follow ecological theory derived simple thumb rules decisions to better design protected areas.
+However, as noted rapidly, species/area relationships do not have a clear association with fragmentation per se but mainly with overall habitat loss [@Simberloff1982].
+Aiming to design better management practices, this led to a nearly 50-year Single Large vs Several Small (SLOSS) debate.
 While some authors argue that habitat loss is the most overwhelming effect on species loss, for others the complex factors arising from fragmentation further enhance species extinctions [@Fahrig2017;@Fletcher2018].
 
-The SLOSS debate has been an important academic and practical discussion in conservation biology.
-In practice, a series of decision-maker targeted recommendations were proposed and remain rather uncritically in several conservation instances (quotes).
+The SLOSS debate has been an important academic and intellectual discussion in conservation biology.
+However, in practical implementation, numerous conservation agencies persist in prioritizing the safeguarding of extensive, contiguous habitat areas, often neglecting the protection of smaller patches of natural habitat [@Armsworth2018].
+As an example, the current emphasis on 'rewilding' in Europe is geared towards the conservation and restoration of contiguous natural habitats, with a specified 'core area' of at least 10,000 hectares [@EuroparcFederation2013].
+Similarly, conservation efforts for large water bodies take precedence, leaving small ponds or pondscapes largely excluded from priority considerations [@Hill2018].
+
 While the overall habitat loss is widely accepted as a process leading to long-term biodiversity loss [@Keil2015;@Hodgson2011], the proposition that single large reserves retain more species than several small reserves of the same area (SL > SS) remain at a similar hierarchical state, even after several criticisms (reviewed in  @Fahrig2017).
-Moreover, empirical observational and experimental data shows the predominance of the opposite pattern (SS > SL) ever since the 80s and this fact has been reinforced to this day [@Simberloff1982; @Fahrig2020; @Hammill2020], add more.
-The actual processes driving to it remain elusive and it is probably multi-causal (quotes).
-In that scenario, model simulations could produce useful predictions targeting the relevant variables to better understand the processes leading to non-intuitive patterns.
+Empirical observational and experimental data shows the predominance of the opposite pattern (SS > SL) ever since the 80s and this fact has been reinforced to this day [@Simberloff1982; @Fahrig2020; @Hammill2020].
+The actual processes driving it remain elusive and it is probably multi-causal.
+In that scenario, model simulations could produce useful hypotheses to better understand the processes leading to non-intuitive patterns of biodiversity loss.
 
@@
 Moreover, spatial dimension in several works has been neglected as mere noise around long-term equilibrium dynamics.
-Recent studies using spatially explicit simulations revealed that while high fragmentation leads to more local extinctions due to the reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].
+[cita?] However, recent studies using spatially explicit simulations revealed that while high fragmentation leads to more local extinctions due to the reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].
 
-We considered only one amount of habitat destruction in combination with different community dynamics, neutral and competitive, taking into account healthy and impoverished communities with different dispersal related traits, and different patterns of habitat loss.
+By examining spatially explicit models with a fixed level of habitat destruction alongside neutral and competitive community dynamics, encompassing diverse dispersal-related traits and various geometries of habitat loss, we aim to align the model parameters with the recently proposed SLOSS cube hypothesis  [@Fahrig2021].
+This hypothesis serves as a comprehensive framework summarizing existing SLOSS predictions and empirical findings.
+According to the SLOSS cube hypothesis, the condition where SL > SS will prevail is contingent upon the simultaneous occurrence of the following conditions: low between-patch movement rate, minimal impact of spreading-of-risk on population dynamics, and low species spatial clumping [@Fahrig2021].
 
-This work aims to explore spatial multi-species models to verify the SS > SL pattern and determine the causes of it.
+This work aims to explore spatial multi-species models to verify the conditions when SS > SL pattern holds and determine the causes of it.
 In particular, we simulate species retention after different fragmentation scenarios of the same overall habitat loss, therefore accounting for the interacting effects of habitat loss and fragmentation.
@@
 
-We developed a spatially explicit multi-species model to explore the relationship between habitat loss and species preservation.
-In this model space is discretized into a grid and only one individual of one species can occupy a particular position.
+We developed a spatially explicit multi-species model to explore the relationship between habitat loss and species retention at the regional metapopulation level.
+In this model, space is discretized into a grid and only one individual of one species can occupy a particular position.
 Each position represents an area fixed by the investigator to mimic the real system.
@@
 
-We use absorving boundary conditions, which means that if an individual try to disperse outside the grid it dies.
+We use absorbing boundary conditions, which means that if an individual tries to disperse outside the grid it dies.
 The size of the community is given by *J = dimX* x *dimY*, where *dimX* and *dimY* are the dimensions of the grid.
@@
 
-The classical neutral variation where propagules (e.g., seeds) disperse passively and cannot actively choose their establishment site.
+The classical neutral variation is where propagules (e.g., seeds) disperse passively and cannot actively choose their establishment site.
 During reproduction events, individuals disperse their propagules blindly to surrounding patches.
@@
 
-The neutral variation with habitat selection where individuals can actively select suitable habitat patches.
+The neutral variation with habitat selection is where individuals can actively select suitable habitat patches.
 In this variation, propagules can choose unoccupied patches, ensuring their survival and successful establishment.
 This modification reflects the influence of habitat availability and selection on species colonization and persistence.
-By incorporating habitat selection, this variation better captures the ecological processes and behaviors of species that are capable of actively seeking out suitable habitats for reproduction and survival.
+By incorporating habitat selection, this variation better captures the ecological processes and behaviors of species that are capable of actively seeking out suitable habitats for reproduction and survival, like most animals
 
@@
 This variation captures the influence of species interactions and competitive exclusion on species composition and abundance.
-It is particularly relevant for understanding communities where competitive interactions play a significant role in shaping species distributions and community structure.
+It is particularly relevant for understanding communities where competitive interactions play a significant role in shaping species distributions and community structure (e.g., hummingbirds).
 
@@
 
-    Regular degradation: Habitat is removed so that habitable patches form a regular pattern of squares with side given by an habitat-size parameter.
+    Regular degradation: Habitat is removed so that habitable patches form a regular pattern of squares with sides given by a habitat-size parameter.
 As the percentage of degraded habitat is constant varying the habitat-size parameter results in either a few large habitat patches (SL) or several small (SS).
@@
     Random block degradation: Habitat square patches of a fixed size (habitat-size parameter) are located at random throughout the landscape.
-The sizes are identical to those used for regular degradation simulations, but it results in habitat patches that have a minimum size corresponding to the habitat-size  but a maximum size that is bigger because habitat patches can superpose between them producing a larger patch.
-Then, I characterize this with the maximum habitat patch size.
+The sizes are identical to those used for regular degradation simulations, but it results in habitat patches that have a minimum size corresponding to the habitat-size but a maximum size that is bigger because habitat patches can overlap between them producing a larger patch.
+Then, we characterize this with the maximum habitat patch size.
 
-(2) With $migration-rate$ an individual of a species $i$ can migrate from the metacommunity. A random site on the border of the grid is selected and if it is empty and habitable the site is occupied.
+(2) With $migration-rate$ an individual of a species $i$ can migrate from the metacommunity.
+A random site on the border of the grid is selected and if it is empty and habitable the site is occupied[d].
 
@@
 
-    where $d(x)$ is the probability that an individual disperses a distance $x$ from the parent. In all cases we used $x_{min} = 1$. We use as parameter the mean $dispersal-distance$.
+    where $d(x)$ is the probability that an individual disperses a distance $x$ from the parent. In all cases, we used $x_{min} = 1$. We use as parameter the mean $dispersal-distance$.
 
@@
 
-    b. Neutral with Habitat Selection:  Propagules in this version selectively target unoccupied patches for reproduction within the distance given by the dispersal kernel, ensuring their survival.
+    b. Neutral with Habitat Selection:  Propagules selectively target unoccupied patches for reproduction within the distance given by the dispersal kernel, ensuring their survival.
 
@@
 
-For all simulations we used a metacommunity of 100 species.
-The simulations commence with the colonization of an empty habitat area.
-After 600 time steps fixed amount of 60% habitat is subsequently subjected to destruction, employing regular or random block degradation to mimic habitat fragmentation due to habitat loss.
+For all simulations, we used a metacommunity of 100 species.
+The simulations start with the colonization of an empty habitat area.
+After 600 time steps fixed amount of 60% of habitat is subsequently subjected to destruction, employing regular or random block degradation to mimic habitat fragmentation due to habitat loss.
 <!-- why we don't use completely random degradation patterns -->
@@
 In this model, the relationship between these parameters, $\lambda = growth-rate / death-rate$, is the main determinant of the fate of an individual of a species.
-In similar non-spatial models, this relationship must exceed 1 for a species to persist, but in spatially explicit models the value needs to be higher, around 1.6.
+In similar non-spatial models, this relationship must exceed 1 for a species to persist.
 Thus we used a minimum $\lambda$ of 1.7, we also fixed the $migration-rate$ at 0.0001.
@@
 For each combination of parameters we ran 30 simulations up to 1500 time steps and recorded the number of species, the Shannon Diversity index [@Magurran2013] and the total number of individuals.
-For some representative combinations of parameters we take snapshots of the model output showing the typical dynamics on time of the 3 recorded quantities and the rank abundance species distribution [@Newman2005; @May1975]
+For some representative combinations of parameters, we take snapshots of the model output showing the typical dynamics on time of the 3 recorded quantities and the rank abundance species distribution [@Newman2005; @May1975]
 
@@
 
-The model was developed in Netlogo 6.3.0 and model source code is available at Github <https://github.com/lsaravia/DynamicForestExtinction>, the Overview, Design concepts, Details (ODD), protocol for describing individual-based models is included within the model source code [@Grimm2010].
+The model was developed in Netlogo 6.3.0 and the model source code is available at Github <https://github.com/lsaravia/DynamicForestExtinction>, the Overview, Design concepts, Details (ODD), protocol for describing individual-based models is included within the model source code [@Grimm2010].
 We performed the simulations and analysis using R version 4.3.1 [@RCoreTeam2023], the R source code to perform simulations and analysis is located at <https://github.com/EcoComplex/multiSLOSS>.
 
-## Results
+## Results[e]
 
-In our neutral model simulations with a migration rate of \(0.0001\) and regular habitat patterns, species richness varied with habitat patch sizes, influenced by dispersal distance and the \(\lambda\) parameter (Figure 1).
-Species richness increased with \(\lambda\), with the highest richness at smaller patch sizes for higher \(\lambda\) values.
-Conversely, a \(\lambda\) of 3 or less showed reduced richness at the smallest patch sizes for lower dispersal distances.
-The interplay of dispersal distances, habitat sizes, and \(\lambda\) impacts community survival and species richness.
+In our neutral model simulations with a migration rate of $0.0001$ and regular habitat patterns, species richness varied with habitat patch sizes, influenced by dispersal distance and the $\lambda$ parameter (Figure 1).
+Species richness increased with $\lambda$, with the highest richness at smaller patch sizes for higher $\lambda$ values.
+Conversely, a $\lambda$ of 3 or less showed reduced richness at the smallest patch sizes for lower dispersal distances.
+In these cases only migrants survive in the smallest patches, and no local populations are established (Figures S13 & S14).
+The interplay of dispersal distances, habitat sizes, and $\lambda$ impacts community survival and species richness.
 For $\lambda < 4$ this suggests a threshold patch size influenced by these 3 parameters, for $\lambda = 4$ patch size decreases species richness.
 The differences between increasing $\lambda$'s almost disappear when dispersal distance is higher.
-Extended dispersal distances could enhance species survival in scenarios with low \(\lambda\) and smaller patches but reduce richness beyond a certain threshold.
+Extended dispersal distances could enhance species survival in scenarios with low $\lambda$ and smaller patches but reduce richness beyond a certain threshold.
 Similar patterns are evident for Shannon diversity (Figure 2).
@@
 
-Our research contributes to the ongoing Single Large vs Several Small (SLOSS) debate in conservation biology, offering new insights based on our model findings.
+Our research contributes to the ongoing SLOSS debate in conservation biology, offering new insights based on our model findings.
 Consistent with previous studies, the results from our simulations primarily support the SS > SL pattern for species retention, as observed across various model scenarios [@Fahrig2020].
-This trend underlines the greater efficacy of multiple smaller habitats in preserving species diversity compared to a single large habitat under similar conditions of habitat loss.
+This trend underlines the greater efficacy of multiple smaller habitats in preserving species diversity compared to a single large habitat under equal conditions of habitat loss.
 A critical factor in this dynamic is the balance between reproduction and extinction rates ($\lambda$ parameter) and dispersal relative to the size of the habitat patches, influencing their role as sources of propagules.
@@
 
-In neutral models one species by chance tends to dominate a patch [@Hubbell2001], so in fragmented landscapes, once the species increments its density is very improbable that another species replaces it so with more patches the probability of having more species is augmented.
+In neutral models one species by chance tends to dominate a patch [@Hubbell2001], so in fragmented landscapes, once the species increments its density it is unlikely that another species replaces it so with more patches the probability of having more species is augmented.
 This effect is enhanced when individuals can select the habitat, because of the increased survival of the propagules, thus fragmentation has a positive effect on the number of species.
@@
 Also, an improved version of the model with different values of the parameters for each species could be developed and compared with the neutral model.
-The effects of fragmentation could be influenced by many other ecological/behavioural characteristics of the species not included in the model, environmental heterogeneity, also the size of the focal landscape compared to dispersal distances, which determines the scale of the dynamics could influence quantitative results [@Rybicki2020].
+The effects of fragmentation could be influenced by many other ecological/behavioural characteristics of the species not considered in the model, including environmental heterogeneity, matrix permeability, individual’s cost of dispersal, among others.
+Further, the size of the focal landscape compared to dispersal distances, which determines the scale of the dynamics, could influence quantitative results [@Rybicki2020].
 Besides that, the approach of using mean values of the parameters of relatively simple multispecies models has been applied and tested to predict species biomass [@Fort2018].
@@
 This could be a more realistic scenario than if all species are present with initially random patterns at an already fragmented habitat [@Rybicki2020].
-This could be more important the more heterogeneous was species spatial distribution pattern before fragmentation.
+This could be more important the more heterogeneous was species spatial distribution pattern before fragmentation, as the cube hypothesis states.
+
+In terms of the cube hypothesis the between patch movement is influenced by the dispersal distance in relation to patch size, as the landscape size and the percentage of habitat destruction are constant, the between-patch distance is higher for bigger patch sizes except for the bigest size that is only one patch, also the number of patches is higher for smaller patch sizes thus between patch movement will be higher.
+Dispersal distance also increase between patch movement, but for higher dispersal distances the sizes of the patches are less important and the relationship tends to be SS = SL.
+The spreading of risk is influenced by the $\lambda$, for smaller lambda spreading the risk would be more important, but to fully incorporate the spreading the risk effect the model should include disturbances.
+Very low dispersal distance favours still favours combined with low $\lambda$ favours SL > SS up to a threshold after wich the patch size is not important.
+The species spatial clumping is influenced by the dispersal distance, higher dispersal distance produce less clumpines but this is cofounded with patch movement, so the effect is difficult to disentagle.
+Thus the cube hypothesis is generally compatible with the results of our model.
 
@@
 Under conditions of adequate patch size and dispersal capabilities, the SS scenario typically retains more species.
-This underscores the ecological significance of smaller habitat patches, such as isolated trees, in maintaining biodiversity and landscape connectivity [@Yan2021; @Manning2006; @Shafer1995].
+This underscores the ecological significance of smaller habitat patches in maintaining biodiversity and landscape connectivity [@Yan2021; @Manning2006; @Shafer1995].
 Our results thus challenge the conventional preference for larger reserves in conservation planning and support the paradigm shift in conservation towards considering sets of many small patches generally having higher biodiversity than sets of few large patches [@Riva2023].
 
-In conclusion, our study adds valuable perspectives to the SLOSS debate, illustrating the complexities of habitat fragmentation and species conservation.
+In conclusion, we believe our study adds valuable perspectives to the SLOSS debate, illustrating the complexities of habitat fragmentation and species conservation.
 While our model has inherent limitations, it provides a useful framework for understanding key ecological processes.
@@
 ## References
+
+[a]esta frase no me convence en general pero me parece que algo de esto es necesario para no saltar tanto desde lo que venimos hablando a el "here we..."
+[b]Lo que trato de decir aca es que consideramos modelos de comunidades de distintos tipos, neutral y competitiva, con distintos traits de dispersion y distintos patrones de habitat loss.
+[c]Cambié bastante la frase para agregar la discusion de la cube hypothesis.
+[d]Hay 2 procesos migracion de la metacomunidad y dispersion local
+[e]matendría en los plots el orden y nombres que están en el texto:
+1) Neutral
+2) Habitat Selection
+3) Competitive Hierarchical

### 2024-02-02 RunSims.Rmd
Updates to ms

@@
   theme(
-    legend.position = c(0.9,0.8)
+    legend.position = c(0.94,0.83)
     )

### 2024-02-02 ms_SLOSS.Rmd
Updates to ms

@@
 ## References
-
-[a]esta frase no me convence en general pero me parece que algo de esto es necesario para no saltar tanto desde lo que venimos hablando a el "here we..."
-[b]Lo que trato de decir aca es que consideramos modelos de comunidades de distintos tipos, neutral y competitiva, con distintos traits de dispersion y distintos patrones de habitat loss.
-[c]Cambié bastante la frase para agregar la discusion de la cube hypothesis.
-[d]Hay 2 procesos migracion de la metacomunidad y dispersion local
-[e]matendría en los plots el orden y nombres que están en el texto:
-1) Neutral
-2) Habitat Selection
-3) Competitive Hierarchical

### 2024-02-21 ms_SLOSS.Rmd
Updated ms

@@
 csl: "ecology-letters.csl"
-output: pdf_document
+
+output:
+  pdf_document: 
+    includes:
+      in_header: margins.sty
 ---
@@
 The SLOSS debate has been approached from several theoretical frameworks, but most of them were based on single-species models [@Fahrig2021].
-Moreover, spatial dimension in several works has been neglected as mere noise around long-term equilibrium dynamics.
-[cita?] However, recent studies using spatially explicit simulations revealed that while high fragmentation leads to more local extinctions due to the reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].
+<!--Moreover, spatial dimension in several works has been neglected as mere noise around long-term equilibrium dynamics [cita?]-->.
+However, recent studies using spatially explicit simulations revealed that while high fragmentation leads to more local extinctions due to the reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].
 
@@
 
-    where $d(x)$ is the probability that an individual disperses a distance $x$ from the parent. In all cases, we used $x_{min} = 1$. We use as parameter the mean $dispersal-distance$.
+    where $d(x)$ is the probability that an individual disperses a distance $x$ from the parent. In all cases, we used $x_{min} = 1$. We use as a parameter the mean $dispersal-distance$.
 
@@
 The simulations start with the colonization of an empty habitat area.
-After 600 time steps fixed amount of 60% of habitat is subsequently subjected to destruction, employing regular or random block degradation to mimic habitat fragmentation due to habitat loss.
+After 600 time steps, a fixed amount of 60% of habitat is subsequently subjected to destruction, employing regular or random block degradation to mimic habitat fragmentation due to habitat loss.
 <!-- why we don't use completely random degradation patterns -->
@@
 
-## Results[e]
+## Results
 
@@
 Conversely, a $\lambda$ of 3 or less showed reduced richness at the smallest patch sizes for lower dispersal distances.
-In these cases only migrants survive in the smallest patches, and no local populations are established (Figures S13 & S14).
+In these cases, only migrants survive in the smallest patches, and no local populations are established (Figures S13 & S14).
 The interplay of dispersal distances, habitat sizes, and $\lambda$ impacts community survival and species richness.
@@
 Examples of simulations with random block and regular habitat patterns are shown in the supplementary material (Figures S9 & S32).
-In these figures we can observe that for the minimum patch size only migrants survive for regular habitat patterns with neutral and hierarchical models.
+In these figures, we can observe that for the minimum patch size only migrants survive for regular habitat patterns with neutral and hierarchical models.
 
@@
 
-Comparing the three versions of the model at $\lambda =2$, for higher dispersal the neutral model tend to have higher richness and Shannon diversity at higher patch sizes, followed by the hierarchical model and the neutral-with-habitat-selection model (Figure 3 & 4).
+Comparing the three versions of the model at $\lambda =2$, for higher dispersal the neutral model tends to have higher richness and Shannon diversity at higher patch sizes, followed by the hierarchical model and the neutral-with-habitat-selection model (Figure 3 & 4).
 At the lowest patch size, only in the neutral-with-habitat-selection have populations that can survive, because in the other model versions the individuals observed are migrants that cannot establish local populations and become extinct (See Figures S5 - S15).
@@
 
-Examples of the final pattern of species and habitat distributions is shown in Figures S9- S32.
+Examples of the final pattern of species and habitat distributions are shown in Figures S9- S32.
 Species spatial distributions closely match the spatial distribution previous to the habitat destruction for neutral and neutral-with-habitat-selection models.
-Is also clearly observed that for the hierarchical model there is a strong dominance of the more competitive species where Shannon Diversity is very low for higher dispersal rate (Figure 4 & S23).
+Is also clearly observed that for the hierarchical model, there is a strong dominance of the more competitive species where Shannon Diversity is very low for higher dispersal rate (Figure 4 & S23).
 
@@
 In the neutral-with-habitat-selection model, the edge effect is not present because the individuals search for empty patches, so the individuals on the border of the patch have the same probability of extinction as the individuals in the interior of the patch.
-Thus when patches are smaller the edge proportion is higher and the edge effect is more important, but this plays a more important role in the minimal patch size for survival of species than for the species richness.
+Thus when patches are smaller the edge proportion is higher and the edge effect is more important, but this plays a more important role in the minimal patch size for the survival of species than for the species richness.
 
@@
 The exceptions are stressed communities---represented in our models with lowest $\lambda$--- with low dispersal may have a continuous increment of richness with patch size, observed only in random block degradation patterns.
-This could be related to the optimal patch size for random block habitat loss patterns tends to be higher than for regular patterns, so we could hypothesize that the optimal was outside the simulated range for these stressed communities.
+This could be related to the optimal patch size for random block habitat loss patterns, which tends to be higher than for regular patterns, so we could hypothesize that the optimal was outside the simulated range for these stressed communities.
 The aggregation of patches in the random block pattern (implies aggregation of the degraded habitat) could be positive if individual habitat patches are very small because it allows some patches to surpass the minimum size required for survival.
@@
 Also, an improved version of the model with different values of the parameters for each species could be developed and compared with the neutral model.
-The effects of fragmentation could be influenced by many other ecological/behavioural characteristics of the species not considered in the model, including environmental heterogeneity, matrix permeability, individual’s cost of dispersal, among others.
+The effects of fragmentation could be influenced by many other ecological/behavioural characteristics of the species not considered in the model, including environmental heterogeneity, matrix permeability, and individual’s cost of dispersal, among others.
 Further, the size of the focal landscape compared to dispersal distances, which determines the scale of the dynamics, could influence quantitative results [@Rybicki2020].
@@
 
-In terms of the cube hypothesis the between patch movement is influenced by the dispersal distance in relation to patch size, as the landscape size and the percentage of habitat destruction are constant, the between-patch distance is higher for bigger patch sizes except for the bigest size that is only one patch, also the number of patches is higher for smaller patch sizes thus between patch movement will be higher.
-Dispersal distance also increase between patch movement, but for higher dispersal distances the sizes of the patches are less important and the relationship tends to be SS = SL.
+In terms of the cube hypothesis, the between patch movement is influenced by the dispersal distance concerning patch size, as the landscape size and the percentage of habitat destruction are constant, the between-patch distance is higher for bigger patch sizes except for the biggest size that is only one patch, also the number of patches is higher for smaller patch sizes thus between patch movement will be higher.
+Dispersal distance also increases between patch movements, but for higher dispersal distances the sizes of the patches are less important and the relationship tends to be SS = SL.
 The spreading of risk is influenced by the $\lambda$, for smaller lambda spreading the risk would be more important, but to fully incorporate the spreading the risk effect the model should include disturbances.
-Very low dispersal distance favours still favours combined with low $\lambda$ favours SL > SS up to a threshold after wich the patch size is not important.
-The species spatial clumping is influenced by the dispersal distance, higher dispersal distance produce less clumpines but this is cofounded with patch movement, so the effect is difficult to disentagle.
+Very low dispersal distance favours still favours combined with low $\lambda$ favours SL > SS up to a threshold after which the patch size is not important.
+The species' spatial clumping is influenced by the dispersal distance, higher dispersal distance produces less clumpines but this is cofounded with patch movement, so the effect is difficult to disentangle.
 Thus the cube hypothesis is generally compatible with the results of our model.

### 2024-02-21 R/trackdown.R

@@ -1,0 +1,4 @@
+
+trackdown::update_file(file = "ms_SLOSS.Rmd", hide_code = TRUE)
+
+trackdown::download_file(file = "ms_SLOSS.Rmd")

### 2024-02-23 ms_SLOSS.Rmd
Updates to ms

@@
 However, to date, most theories regarding this issue were based on single-species models.
-This work aims to explore spatial multi-species models to verify the SS > SL pattern and determine the causes of it.
+This work aims to explore spatial multi-species models to verify the SS > SL pattern and determine its causes.
 We used three types of models: a classical neutral, neutral with habitat selection, and a competitive hierarchical model.
@@
 Starting from a landmark paper [@Diamond1975], conservation biologists engaged in the challenging task of recommending protection measures given the environmental crisis coupled with a lack of knowledge of most of the world’s biodiversity and limited funding for conservation.
-In that context, early recommendations claiming to follow ecological theory derived simple thumb rules decisions to better design protected areas.
+In that context, early recommendations claiming to follow ecological theory derived simple thumb rules to better design protected areas.
 However, as noted rapidly, species/area relationships do not have a clear association with fragmentation per se but mainly with overall habitat loss [@Simberloff1982].
@@
 
-Comparing the three versions of the model at $\lambda =2$, for higher dispersal the neutral model tends to have higher richness and Shannon diversity at higher patch sizes, followed by the hierarchical model and the neutral-with-habitat-selection model (Figure 3 & 4).
+Comparing the three versions of the model at $\lambda =2$, for higher dispersal the neutral model tends to have higher richness and Shannon diversity at higher patch sizes, followed by the hierarchical model and the neutral-with-habitat-selection model (Figures 3 & 4).
 At the lowest patch size, only in the neutral-with-habitat-selection have populations that can survive, because in the other model versions the individuals observed are migrants that cannot establish local populations and become extinct (See Figures S5 - S15).
@@
 Species spatial distributions closely match the spatial distribution previous to the habitat destruction for neutral and neutral-with-habitat-selection models.
-Is also clearly observed that for the hierarchical model, there is a strong dominance of the more competitive species where Shannon Diversity is very low for higher dispersal rate (Figure 4 & S23).
+Is also clearly observed that for the hierarchical model, there is a strong dominance of the more competitive species where Shannon Diversity is very low for a higher dispersal rate (Figure 4 & S23).
 
@@
 
-In previous simulation studies, it was shown that fragmentation was important when the habitat amount is low, around 10% [@Hanski2015; @Rybicki2013], but it has been observed that with levels habitat around 10-20% the effect of fragmentation is positive or have an intermediate maximum [@Loke2019; @With2016], and with lower habitat amount (e.g 7%) the arrangement of patches does not matters, we obtained similar patterns with 40% habitat amount, which is near the amount used in newer simulated studies (32%) that also obtain more species richness with more fragmentation [@Rybicki2020].
+In previous simulation studies, it was shown that fragmentation was important when the habitat amount is low, around 10% [@Hanski2015; @Rybicki2013], but it has been observed that with habitat levels around 10-20% the effect of fragmentation is positive or have an intermediate maximum [@Loke2019; @With2016], and with lower habitat amount (e.g 7%) the arrangement of patches does not matters, we obtained similar patterns with 40% habitat amount, which is near the amount used in newer simulated studies (32%) that also obtain more species richness with more fragmentation [@Rybicki2020].
 
-The novel aspect of our model is that it considers communities that are assembled from a metacommunity in which individuals migrate to an empty habitat, and then after the number of species is stabilized the habitat is fragmented.
-This could be a more realistic scenario than if all species are present with initially random patterns at an already fragmented habitat [@Rybicki2020].
+The novel aspect of our model is that it considers communities that are assembled from a metacommunity in which individuals migrate to an empty habitat.
+After the number of species is stabilized the habitat is fragmented.
+This scenario is deemed more realistic than having all species initially present with random patterns in an already fragmented habitat  [@Rybicki2020].
 This could be more important the more heterogeneous was species spatial distribution pattern before fragmentation, as the cube hypothesis states.
 
-In terms of the cube hypothesis, the between patch movement is influenced by the dispersal distance concerning patch size, as the landscape size and the percentage of habitat destruction are constant, the between-patch distance is higher for bigger patch sizes except for the biggest size that is only one patch, also the number of patches is higher for smaller patch sizes thus between patch movement will be higher.
-Dispersal distance also increases between patch movements, but for higher dispersal distances the sizes of the patches are less important and the relationship tends to be SS = SL.
-The spreading of risk is influenced by the $\lambda$, for smaller lambda spreading the risk would be more important, but to fully incorporate the spreading the risk effect the model should include disturbances.
-Very low dispersal distance favours still favours combined with low $\lambda$ favours SL > SS up to a threshold after which the patch size is not important.
-The species' spatial clumping is influenced by the dispersal distance, higher dispersal distance produces less clumpines but this is cofounded with patch movement, so the effect is difficult to disentangle.
-Thus the cube hypothesis is generally compatible with the results of our model.
+In terms of the cube hypothesis, low between-patch movement, minimal impact of spreading-of-risk on population dynamics, and low species spatial clumping are needed for SL > SS.
+In our model dispersal distance increases between patch movements, but for higher dispersal distances the sizes of the patches are less important and the relationship tends to be SS = SL.
+Spreading-of-risk is influenced by $\lambda$, which represents the relationship between the growth and death rates.
+For smaller $\lambda$, spreading-of-risk becomes more significant because the probability of local extinction is higher.
+A very low dispersal distance combined with low $\lambda$ favors SL > SS up to a threshold in patch size, beyond which SL = SS.
+The spatial clumping of species is influenced by the dispersal distance; a higher dispersal distance produces less clumpiness.
+However, this is confounded with patch movement, making it challenging to disentangle the effect.
+Therefore, the results of our model are generally compatible with the cube hypothesis.
 

### 2024-02-26 ms_SLOSS.Rmd
Ms corrections

@@
 
-For the simulations made with random block habitat patterns, the patches do not have the same sizes but there is a distribution with a minimum size determined by the patch-size parameter and a bigger maximum size (Figures S9 - S11).
-The pattern is similar to the regular case,  with a maximum richness for intermediate patch sizes and values of $\lambda$ higher than 2 the maximum is at the lower habitat patch size (Figure S1 & S2).
-The sizes of patches for the random block simulations are bigger so species richness and diversity are lower, except for the minimum patch size (Figure S7 & S8).
-Examples of simulations with random block and regular habitat patterns are shown in the supplementary material (Figures S9 & S32).
-In these figures, we can observe that for the minimum patch size only migrants survive for regular habitat patterns with neutral and hierarchical models.
+For the simulations conducted with randomly arranged block habitat patterns, the patches exhibit variable sizes, following a distribution with a minimum size determined by the patch-size parameter and a larger maximum size determined by the superposition of several patches (Figures S9 - S11).
+The overall pattern is analogous to the regular case, with a peak in richness occurring for intermediate patch sizes.
+However, for values of $\lambda$ higher than 2, the maximum richness is observed at the smaller habitat patch size (Figure S1 & S2).
+The patches in the random block simulations are generally larger, resulting in lower species richness and diversity, except for the minimum patch size (Figure S7 & S8).
+Supplementary material (Figures S9 & S32) includes examples of simulations with random block and regular habitat patterns.
+In these figures, it is evident that for the minimum patch size, only migrants survive in regular habitat patterns with neutral and hierarchical models.
 
@@
 
-In previous simulation studies, it was shown that fragmentation was important when the habitat amount is low, around 10% [@Hanski2015; @Rybicki2013], but it has been observed that with habitat levels around 10-20% the effect of fragmentation is positive or have an intermediate maximum [@Loke2019; @With2016], and with lower habitat amount (e.g 7%) the arrangement of patches does not matters, we obtained similar patterns with 40% habitat amount, which is near the amount used in newer simulated studies (32%) that also obtain more species richness with more fragmentation [@Rybicki2020].
+In previous simulation studies, the importance of fragmentation was demonstrated to be pronounced when the habitat amount is low, approximately around 10% [@Hanski2015; @Rybicki2013].
+However, experimental studies have observed that with habitat levels around 10-20%, the impact of fragmentation can be positive or show an intermediate maximum [@Loke2019; @With2016].
+When habitat amounts are lower (e.g., 7%), the arrangement of patches appears to have no significant effect.
+Interestingly, our study replicated similar patterns even with a 40% habitat amount.
+Moreover, other studies employing comparable habitat amounts (32%) also reported higher species richness with increased fragmentation [@Rybicki2020].
 

### 2024-02-27 RunSims.Rmd
Updated manuscript with new figure 5

@@
 ```
+
+## Figure of spatial patterns
+
+```{r PlotsSpatialPattern, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+require(cowplot)
+require(ggplot2)
+require(png)
+
+image_paths <- c("Figures/NoSelection_lambda2_dd15_ps3_before_nsp37.png","Figures/NoSelection_lambda2_dd15_ps3_after_nsp32.png","Figures/NoSelection_lambda2_dd15_ps3_after_lt_nsp3.png",
+                 "Figures/NoSelection_lambda2_dd15_ps29_ before_nsp36.png","Figures/NoSelection_lambda2_dd15_ps29_ after_nsp29.png","Figures/NoSelection_lambda2_dd15_ps29_ after_lt_nsp25.png")
+
+plots <- lapply(seq_along(image_paths), function(i) {
+  ggdraw() +  draw_image(image_paths[i],scale=0.9)
+  #ggplot() + annotation_custom(ggplotGrob(readPNG(image_paths[i])), xmin=-Inf, xmax=Inf, ymin=-Inf, ymax=Inf) +
+})
+
+aplots <- plot_grid(plotlist = plots, ncol = 2,
+  labels = c('A.', 'B.', 'C.', 'D.','E.','F.'),
+  label_fontface = "plain",
+  byrow = FALSE
+  )
+aplots
+save_plot("Figures/ModelSpatialPatternComparison.png",aplots,base_width=6,base_height=8,dpi=600)
+
+```

### 2024-02-27 ms_SLOSS.Rmd
Updated manuscript with new figure 5

@@
 Thus an intermediate patch size with a maximum species richness is not observed for this model.
+Is also clearly observed that for the hierarchical model, there is a strong dominance of the more competitive species where Shannon Diversity is very low for a higher dispersal rate (Figure 4 & S23).
 
-Examples of the final pattern of species and habitat distributions are shown in Figures S9- S32.
-Species spatial distributions closely match the spatial distribution previous to the habitat destruction for neutral and neutral-with-habitat-selection models.
-Is also clearly observed that for the hierarchical model, there is a strong dominance of the more competitive species where Shannon Diversity is very low for a higher dispersal rate (Figure 4 & S23).
+Examples of the spatial patterns of species and habitat distributions are illustrated in Figures 5 and S9-S32.
+Simulations with the neutral model depict the species' spatial distribution before fragmentation (Figure 5 A & B) and after fragmentation (Figure 5 C to F), with habitat destruction represented in purple areas.
+Immediately after fragmentation, the species' spatial distributions closely resemble those before habitat destruction, but then the patch compositions diverge.
+The divergence in patch species composition is more pronounced in the hierarchical model, where small patches tend to be dominated by one species (Figures S17 - S24).
+The effect of the minimum patch size for survival is also observed (Figure 5 E).
+
+![Examples of species spatial pattern before and after fragmentation for the neutral model.
+Simulations were performed using a fixed growth rate ($\lambda=2$), migration rate of 0.0001 and dispersal distance 1.5.
+A.
+& B.
+spatial pattern before fragmentation.
+C.
+& D.
+Spatial pattern after fragmentation where 60% of the habitat was destroyed (Purple color), with patch-sizes 3 (C) and 29 (D).
+E.
+& F.
+Spatial pattern after 600 time steps where the extinction of populations is observed at smaller patch-size (E)](Figures/ModelSpatialPatternComparison.png)
 

### 2024-03-08 ms_SLOSS.Rmd
Correciones menores

@@
 
-Leonardo A. Saravia ^1^, Ulises Balza ^1^, Fernando Momo^2^
+Leonardo A. Saravia ^1^^2^, Ulises Balza ^1^, Fernando Momo^3^^4^
 
-1. Centro Austral de Investigaciones Científicas (CADIC-CONICET). Houssay 200. Ushuaia, Argentina
+1. Centro Austral de Investigaciones Científicas (CADIC-CONICET). Houssay 200 (9410), Ushuaia, Argentina.
 
-2. Instituto de Ciencias. Universidad Nacional de General Sarmiento. J.M. Gutierrez 1159 (1613), Los Polvorines.        Buenos Aires, Argentina.
+2. Universidad Nacional de Tierra del Fuego (UNTDF), Yrigoyen 879 (9410), Ushuaia, Argentina
+
+3. Instituto de Ciencias. Universidad Nacional de General Sarmiento. J.M. Gutierrez 1159 (1613), Los Polvorines, Argentina.
+
+4. Departamento de Ciencias Básicas. Universidad Nacional de Luján. Rutas 5 y 7, (6700) Luján. Argentina.
 
@@
 The SLOSS debate has been approached from several theoretical frameworks, but most of them were based on single-species models [@Fahrig2021].
-<!--Moreover, spatial dimension in several works has been neglected as mere noise around long-term equilibrium dynamics [cita?]-->.
 However, recent studies using spatially explicit simulations revealed that while high fragmentation leads to more local extinctions due to the reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].
 
-By examining spatially explicit models with a fixed level of habitat destruction alongside neutral and competitive community dynamics, encompassing diverse dispersal-related traits and various geometries of habitat loss, we aim to align the model parameters with the recently proposed SLOSS cube hypothesis  [@Fahrig2021].
-This hypothesis serves as a comprehensive framework summarizing existing SLOSS predictions and empirical findings.
-According to the SLOSS cube hypothesis, the condition where SL > SS will prevail is contingent upon the simultaneous occurrence of the following conditions: low between-patch movement rate, minimal impact of spreading-of-risk on population dynamics, and low species spatial clumping [@Fahrig2021].
+The recently proposed SLOSS cube hypothesis by @Fahrig2021 serves as a comprehensive framework summarizing existing Single Large or Several Small (SLOSS) predictions and empirical findings.
+This hypothesis posits that the condition where Single Large (SL) patches contribute more to biodiversity conservation than Several Small (SS) patches is contingent upon the simultaneous occurrence of specific conditions: a low between-patch movement rate, minimal impact of spreading-of-risk on population dynamics, and low species spatial clumping [@Fahrig2021].
 
-This work aims to explore spatial multi-species models to verify the conditions when SS > SL pattern holds and determine the causes of it.
-In particular, we simulate species retention after different fragmentation scenarios of the same overall habitat loss, therefore accounting for the interacting effects of habitat loss and fragmentation.
+This research aims to investigate multi-species spatially explicit models, incorporating a fixed level of habitat destruction, neutral and competitive community dynamics, diverse dispersal-related traits, and various geometries of habitat loss.
+The principal objective is to verify the conditions under which the SS > SL pattern holds and discern the underlying causes.
+Through this examination, we seek to contribute insights that can help determine if the conditions outlined in the SLOSS cube hypothesis are supported by our modeling approach..
+
+Specifically, our approach involves simulating species retention under diverse fragmentation scenarios, all characterized by a fixed amount of habitat loss.
+We vary parameters such as patch sizes, community type (competitive or neutral), species dispersal capabilities, and the community's persistence status—whether it is near extinction and degraded or in a healthy state.
+This comprehensive exploration enables us to account for the interacting effects of habitat loss and a range of model parameters on species retention.
 
@@
 
+| $\lambda$ | Dispersal Distance | Habitat Patch size |
+| --------: | ---------:         | ---------:         |
+| 1.7       | 1.1                | 3                  |
+| 2.0       | 1.2                | 9                  |
+| 3.0       | 1.5                | 29                 |
+| 4.0       | 2                  | 61                 |
+|           | 3                  |                    |
+
+Table: Parameter values utilized in simulations, with factorial combinations of these parameters employed to run simulations.
+The symbol $\lambda$ represents the ratio of growth rate to death rate.
+"Dispersal distance" denotes the mean of the power law dispersal kernel, while "Habitat patch size" indicates the side length of habitat patches modeled as squares.
+Both the dispersal distance and habitat patch size are measured in grid cells.
+
 The parameters encompass a range of $growth-rates$ while holding the $death-rate$ fixed at 1.
@@
 For some representative combinations of parameters, we take snapshots of the model output showing the typical dynamics on time of the 3 recorded quantities and the rank abundance species distribution [@Newman2005; @May1975]
-
-| $\lambda$ | Dispersal Distance | Habitat Patch size |
-| --------: | ---------:         | ---------:         |
-| 1.7       | 1.1                | 3                  |
-| 2.0       | 1.2                | 9                  |
-| 3.0       | 1.5                | 29                 |
-| 4.0       | 2                  | 61                 |
-|           | 3                  |                    |
-
-Table: Parameter values used in simulations.
-Factorial combinations of these parameters were used to run simulations.
-$\lambda$ represents the ratio of growth rate to death rate.
-"Dispersal distance" refers to the mean of the power law dispersal kernel.
-"Habitat patch size" indicates the side length of habitat patches modeled as squares.
 
@@
 Further, the size of the focal landscape compared to dispersal distances, which determines the scale of the dynamics, could influence quantitative results [@Rybicki2020].
-Besides that, the approach of using mean values of the parameters of relatively simple multispecies models has been applied and tested to predict species biomass [@Fort2018].
+Additionally, the application of mean parameter values in relatively simple multispecies models has been successfully employed and tested for predicting species biomass, as demonstrated by @Fort2018.
 

### 2024-03-11 ms_SLOSS.Rmd
Updated discussion

@@
 The SLOSS debate has been approached from several theoretical frameworks, but most of them were based on single-species models [@Fahrig2021].
+Despite the long-standing discussion, there remains a notable scarcity of multi-species models explicitly designed to validate fragmentation theories.
 However, recent studies using spatially explicit simulations revealed that while high fragmentation leads to more local extinctions due to the reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].
@@
 
-The novel aspect of our model is that it considers communities that are assembled from a metacommunity in which individuals migrate to an empty habitat.
+One novel aspect of our model is that it considers communities that are assembled from a metacommunity in which individuals migrate to an empty habitat.
 After the number of species is stabilized the habitat is fragmented.
@@
 Our results thus challenge the conventional preference for larger reserves in conservation planning and support the paradigm shift in conservation towards considering sets of many small patches generally having higher biodiversity than sets of few large patches [@Riva2023].
+The model's primary outcomes could be validated using empirical data by assessing the prediction that species with greater dispersal capabilities would exhibit a lower SS effect compared to species with limited dispersal.
 
@@
 While our model has inherent limitations, it provides a useful framework for understanding key ecological processes.
-Future research should aim to extend these findings to more diverse ecological settings and incorporate additional factors influencing biodiversity in fragmented landscapes.
+Future research should aim to extend these findings to more diverse ecological settings and incorporate how heterogeneity between species influence biodiversity in fragmented landscapes.
 

### 2024-03-11 ms_SLOSS.Rmd
Changes to discussion and figure caption

@@
 
-![Examples of species spatial pattern before and after fragmentation for the neutral model.
-Simulations were performed using a fixed growth rate ($\lambda=2$), migration rate of 0.0001 and dispersal distance 1.5.
+![Examples of species spatial patterns before and after fragmentation for the neutral model are illustrated.
+Different shades from dark green to white represent distinct species, with the destroyed habitat depicted in purple.
+Simulations were conducted with a fixed growth rate ($\lambda=2$), a migration rate of 0.0001, and a dispersal distance of 1.5.
 A.
 & B.
-spatial pattern before fragmentation.
+showcase spatial patterns before fragmentation.
 C.
 & D.
-Spatial pattern after fragmentation where 60% of the habitat was destroyed (Purple color), with patch-sizes 3 (C) and 29 (D).
+depict spatial patterns after fragmentation, where 60% of the habitat was destroyed, with patch sizes of 3 (C) and 29 (D).
+In these cases, the significance of the species' spatial distribution before fragmentation becomes evident.
 E.
 & F.
-Spatial pattern after 600 time steps where the extinction of populations is observed at smaller patch-size (E)](Figures/ModelSpatialPatternComparison.png)
+display spatial patterns after 600 time steps, illustrating the extinction of populations at smaller patch sizes (E).](Figures/ModelSpatialPatternComparison.png)
 
@@
 
-In neutral models one species by chance tends to dominate a patch [@Hubbell2001], so in fragmented landscapes, once the species increments its density it is unlikely that another species replaces it so with more patches the probability of having more species is augmented.
-This effect is enhanced when individuals can select the habitat, because of the increased survival of the propagules, thus fragmentation has a positive effect on the number of species.
+In neutral models, where species are considered equivalent, one species, by chance, tends to dominate a patch (Hubbell, 2001).
+Consequently, in fragmented landscapes, once a species increases its density, the likelihood of another species replacing it is low.
+With more patches, the probability of having a greater number of species is heightened.
+This effect is further amplified when individuals can actively select their habitat, leading to increased survival of propagules.
+Consequently, fragmentation exhibits a positive influence on the number of species in such scenarios.
+
 For competitive communities, there is a higher probability that different species dominate different smaller patches, compared with a bigger patch.

### 2024-03-19 ms_SLOSS.Rmd
Changes to submit

@@
 ---
-
-Leonardo A. Saravia ^1^^2^, Ulises Balza ^1^, Fernando Momo^3^^4^
-
-1. Centro Austral de Investigaciones Científicas (CADIC-CONICET). Houssay 200 (9410), Ushuaia, Argentina.
-
-2. Universidad Nacional de Tierra del Fuego (UNTDF), Yrigoyen 879 (9410), Ushuaia, Argentina
-
-3. Instituto de Ciencias. Universidad Nacional de General Sarmiento. J.M. Gutierrez 1159 (1613), Los Polvorines, Argentina.
-
-4. Departamento de Ciencias Básicas. Universidad Nacional de Luján. Rutas 5 y 7, (6700) Luján. Argentina.
 
@@
 ## Keywords
+
+Habitat fragmentation, biodiversity, spatially explicit models, multi-species models, neutral theory, competitive theory, habitat selection, dispersal, extinction, reproduction.
 
@@
 
-The model was developed in Netlogo 6.3.0 and the model source code is available at Github <https://github.com/lsaravia/DynamicForestExtinction>, the Overview, Design concepts, Details (ODD), protocol for describing individual-based models is included within the model source code [@Grimm2010].
-We performed the simulations and analysis using R version 4.3.1 [@RCoreTeam2023], the R source code to perform simulations and analysis is located at <https://github.com/EcoComplex/multiSLOSS>.
+The model was developed in Netlogo 6.3.0 and the model source code is available at Github <https://>, the Overview, Design concepts, Details (ODD), protocol for describing individual-based models is included within the model source code [@Grimm2010].
+We performed the simulations and analysis using R version 4.3.1 [@RCoreTeam2023], the R source code to perform simulations and analysis is located at <https://>.
 

### 2024-03-19 ms_SLOSS_title.Rmd
Changes to submit

---
title: "Why there are more species in several small patches vs few large patches:
  a multi-species modeling approach"
bibliography: sloss_bib.bib
csl: "ecology-letters.csl"

output:
  pdf_document: 
    includes:
      in_header: margins.sty
---

Leonardo A. Saravia ^1^^2^, Ulises Balza ^1^, Fernando Momo^3^^4^

1. Centro Austral de Investigaciones Científicas (CADIC-CONICET). Houssay 200 (9410), Ushuaia, Argentina.

2. Universidad Nacional de Tierra del Fuego (UNTDF), Yrigoyen 879 (9410), Ushuaia, Argentina

3. Instituto de Ciencias. Universidad Nacional de General Sarmiento. J.M. Gutierrez 1159 (1613), Los Polvorines, Argentina.

4. Departamento de Ciencias Básicas. Universidad Nacional de Luján. Rutas 5 y 7, (6700) Luján. Argentina.

## Acknowledgments

We acknowledge the financial support provided by the "Fondo para la Investigación Científica y Tecnológica FONCYT" (Grant No. PICT-2020-SERIEA-02628). Additionally, we extend our gratitude to the National University of General Sarmiento (UNGS), especially Santiago Doyle, for granting access to the computational resources utilized in this study.

### 2024-03-19 ms_SLOSS.Rmd
Minor Changes to submit

@@
 bibliography: sloss_bib.bib
-csl: "ecology-letters.csl"
+csl: "functional-ecology.csl"
 

### 2024-07-01 RunSims.Rmd
New simulations no-selection with habitat destruction of 0.4 and 0.8

@@
 
+```
+
+## Lambda = 1-4 NoSelection = Neutral  Regular prog frag 0.4-0.8
+
+```{r SLOSS_lambda14_pf48_regular,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=TRUE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="NoSelection_pf04-08_hps3-61_dd1-3_brb1-4_regular",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4)),
+                                             "prob-frag" =  list(values=c(0.4,0.8))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"NoSelection\"",
+                                             "replacement-rate" = 0.0
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 20*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
 ```

### 2024-07-01 ms_SLOSS.Rmd
Add Format files
New simulations no-selection with habitat destruction of 0.4 and 0.8

--- 2024-03-19 ms_SLOSS.Rmd
@@
 bibliography: sloss_bib.bib
-csl: "ecology-letters.csl"
-
+csl: "functional-ecology.csl"
 output:
@@
       in_header: margins.sty
+  word_document: default
 ---
@@
 
+1.
 As global biodiversity is rapidly declining due to habitat loss, it is important to determine how to protect it.
@@
 However, to date, most theories regarding this issue were based on single-species models.
+
+2.
 This work aims to explore spatial multi-species models to verify the SS > SL pattern and determine its causes.
@@
 We use 100 species and start simulating the colonization of an empty area then we destroy a fixed amount of habitat using different regular patch sizes.
+
+3.
 The results for all models are almost always that SS retains more species than SL.
@@
 If patch size is above the critical size to maintain local populations the SS > SL pattern is due to the sampling effect of aggregated species distribution combined with the reduction of dispersal ability of species.
+
+4.
 High dispersal rates produce homogenization of species distribution and a reduction of species richness and this is why habitat fragmentation by habitat loss retains more species when patches are smaller.
@@
 
-Habitat fragmentation, biodiversity, spatially explicit models, multi-species models, neutral theory, competitive theory, habitat selection, dispersal, extinction, reproduction.
+Habitat fragmentation, biodiversity, spatially explicit models, multi-species models, neutral theory.
 

### 2024-07-15 RunSims.Rmd
Finished simulations for noSelection prog-frag 0.4 0.6 start simulations hierarchical

@@
 
+# NetLogo installation path (adjust to your needs!):
+#
+netlogopath <- file.path("/home/leonardo/NetLogo")
+outpath <- file.path("Simulations")
+
 #
@@
   simfolder <- "/home/leonardo/Dropbox/Projects/DynamicForestExtinction"                       
+  Sys.setenv(JAVA_HOME = "/usr/lib/jvm/java-17-openjdk-amd64")
 
@@
   simfolder <- "~/Dropbox/Projects/DynamicForestExtinction"
+  netlogopath <- file.path("/Applications/NetLogo\ 6.4.0")                   
+  Sys.setenv(JAVA_HOME = "/Library/Java/JavaVirtualMachines/liberica-jdk-17.jdk/Contents/Home")
+
 }
 
-# Unix default NetLogo installation path (adjust to your needs!):
-#
-netlogopath <- file.path("/home/leonardo/NetLogo")
 modelpath <- file.path(simfolder, "MultipleContactNeutralMigrat.nlogo")
-outpath <- file.path("Simulations")
 
@@
 
-nl <- nl(nlversion = "6.3",
+nl <- nl(nlversion = "6.4",
          nlpath = netlogopath,
@@
 
-```{r SLOSS_lambda14_pf48_regular,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=TRUE}
+```{r SLOSS_lambda14_pf48_regular,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 
@@
 
+## Read Simulations pf 0.4 - 0.8
+
+```{r readSimulations_pf48, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+# Calculate cluster stats
+#
+# Old 10 - 20 simulations 
+# mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_ff.csv"),skip=0) %>% 
+# mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_20_ff.csv"),skip=0) %>%
+
+mdl <- read_sloss_simul(outpath,"NoSelection_pf04-08_hps3-61_dd1-3_brb1-4_regular_ff.csv")
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
+
+```
+
 ## Plots - No selection - regular pattern
@@
 mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
-                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,prob_frag) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
 
@@
 
-mdl1 <- mdl1 %>%  mutate(across(max_patch_mean:max_patch_qh,sqrt))
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
 
@@
 mdl1 %>%  
-  ggplot(aes(x = max_patch_mean, y = habitat_patch_size, color = factor(habitat_patch_size))) + 
-  geom_point(  size = 1.2) + guides(color = "none") #+
-
-mdl1 %>%  
-  ggplot(aes(x = max_patch_mean, y = degraded_mfp_mean, color = factor(habitat_patch_size))) + 
-  geom_point(  size = 1.2) + guides(color = "none") #+
-
-mdl1 %>%  
-  ggplot(aes(x = max_patch_mean, y = non_degraded_mfp_mean, color = factor(habitat_patch_size))) + 
-  geom_point(  size = 1.2) + guides(color = "none") #+
-
-mdl1 %>%  
-  ggplot(aes(x = degraded_mfp_mean, y = non_degraded_mfp_mean, color = factor(habitat_patch_size))) + 
+  ggplot(aes(x = habitat_patch_size, y = mean_patch_mean, color = factor(prob_frag))) + 
+  geom_point(  size = 1.2) + guides(color = "none") + 
+  theme_bw()
+
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = degraded_mfp_mean, color = factor(prob_frag))) + 
+  geom_point(  size = 1.2) + guides(color = "none") +
+  theme_bw()
+
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = non_degraded_mfp_mean, color = factor(prob_frag))) + 
+  geom_point(  size = 1.2) + guides(color = "none") +
+  theme_bw()
+
+mdl1 %>%  
+  ggplot(aes(x = degraded_mfp_mean, y = non_degraded_mfp_mean, color = factor(prob_frag))) + 
   geom_point(  size = 1.2) + guides(color = "none") #+
@@
 
-mdl1 %>%  
+mdl1 %>%  #filter(prob_frag==0.4) %>%
   ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
   labs(color = "Dispersal\n distance") +
@@
   theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/Species_Habitat_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  ) 
+ggsave("Figures/Species_Habitat_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  #facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
   labs(color = "Dispersal\n distance") +
@@
   theme(
-    legend.position = c(0.9,0.2)
-    )
-
-ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
-
-#
-# Number of species by max_patch size 
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
+
+#
+# Number of species by mean_patch size 
+# 
+mdl1 %>%  
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Species_MeanPatch_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
+
+#
+# shannon_diversity by mean_patch size 
 # 
 mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
-  ggplot(aes(x = max_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
-  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
-  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Max patch size") + ylab("Number of species") +
-  theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/Species_MaxPatch_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
-
-#
-# shannon_diversity by max_patch size 
-# 
-mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
-  ggplot(aes(x = max_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)) ,  size = 1.2) +
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Max patch size") + ylab("Shannon Diversity") +
-  theme(
-    legend.position = c(0.9,0.2)
-    )
-
-ggsave("Figures/Shannon_MaxPatch_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Shannon_MeanPatch_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 
-## Read Simulations No Selection Random Block Patterns
+## Read Simulations No Selection Random Block Patterns prob-frag = 0.6
 
@@
 
+## Lambda = 1-4 / Random Block patterns / prob-frag 0.4 - 0.8
+
+```{r SLOSS_lambda14_pf48_block,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=TRUE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="NoSelection_pf04_08_hps3-61_dd1-3_brb1-4_block",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path", "find-cluster-sizes"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4)),
+                                               "prob-frag" =  list(values=c(0.4,0.8))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"random block\"",
+                                             "birds-behavior" = "\"NoSelection\"",
+                                             "replacement-rate" = 0.0
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Read Simulations No Selection Random Block Patterns prob-frag 0.4 - 0.8
+
+```{r readSimulationsBlock_pf04-08, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_sloss_simul(outpath,"NoSelection_pf04_08_hps3-61_dd1-3_brb1-4_block_ff.csv")
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
+
+```
+
 ## Plots Neutral No Selection Random Block Patterns Plots
@@
 mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
-                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(c(birds_at_deforestation:count_birds,non_degraded_mfp,degraded_mfp,max_patch) ,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,prob_frag) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
 
 names(mdl1)
-mdl1 <- mdl1 %>%  mutate(across(max_patch_mean:max_patch_qh,sqrt))
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
 
@@
 mdl1 %>%  
-  ggplot(aes(x = max_patch_mean, y = habitat_patch_size, color = factor(habitat_patch_size))) + 
+  ggplot(aes(x = mean_patch_mean, y = habitat_patch_size, color = factor(habitat_patch_size))) + 
   geom_point(  size = 1.2) + guides(color = "none") #+
@@
 mdl1 %>%  
-  ggplot(aes(x = max_patch_mean, y = degraded_mfp_mean, color = factor(habitat_patch_size))) + 
+  ggplot(aes(x = mean_patch_mean, y = degraded_mfp_mean, color = factor(habitat_patch_size))) + 
   geom_point(  size = 1.2) + guides(color = "none") #+
@@
 mdl1 %>%  
-  ggplot(aes(x = max_patch_mean, y = non_degraded_mfp_mean, color = factor(habitat_patch_size))) + 
+  ggplot(aes(x = mean_patch_mean, y = non_degraded_mfp_mean, color = factor(habitat_patch_size))) + 
   geom_point(  size = 1.2) + guides(color = "none") #+
@@
 #
-mdl1 %>%  
+mdl1 %>%  #filter(prob_frag==0.4) %>%
   ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
@@
   geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
   labs(color = "Dispersal\n distance") +
@@
   theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/Species_Habitat_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  ) 
+ggsave("Figures/Species_Habitat_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
 
@@
   ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
   geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
-  theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
 
@@
 # 
-mdl1  %>% #filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
-  ggplot(aes(x = max_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+mdl1 %>%  
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Max patch size") + ylab("Number of species") +
-  theme(
-    legend.position = c(0.9,0.83)
-    )
-
-ggsave("Figures/Species_MaxPatch_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)
-
-#
-# shannon_diversity by max_patch size 
-# 
-mdl1  %>% # filter(birds_dispersal_distance %in% c(1.1,1.5,3)) %>%
-  ggplot(aes(x = max_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Species_MeanPatch_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
+
+#
+# shannon_diversity by mean_patch size 
+# 
+mdl1 %>%  
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Max patch size") + ylab("Shannon Diversity") +
-  theme(
-    legend.position = c(0.9,0.2)
-    )
-
-ggsave("Figures/Shannon_MaxPatch_Dispersal_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Mean Patch Size") +  ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Shannon_MeanPatch_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
+
+#
+# Number of species by habitat mean free path 
+# 
+mdl1 %>%  
+  ggplot(aes(x = non_degraded_mfp_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Mean Free Path") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Species_habitatMFP_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
 
@@
 
+## Simulations Lambda = 1-4 / Regular patterns - Hierarchical communities / prog-frag 0.4-0.8
+
+```{r Hierarchical_SLOSS_lambda14_pf48_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=TRUE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="Hierarchical_pf04-08_hps3-61_dd1-3_brb1-4_regular",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4)),
+                                             "prob-frag" =  list(values=c(0.4,0.8))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"Hierarchical\"",
+                                             "replacement-rate" = 0.3
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
 # Plot for hierarchical communities
@@
 
-## Plots across neutral hierarchical selection modelos regular fragmentation
-
-```{r PlotsAcrossBehaviorRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+## Plots across neutral hierarchical selection modelos fragmentation type
+
+```{r PlotsAcrossFragType, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 

### 2024-08-02 RunSims.Rmd
New simulations with deforestation-at-time=600

@@
 
-* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 ) 
+* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 600 for lambda=4 ) 
 * Check simulations with Migration0-at-deforestation true or false to take into account that immigration happens at the borders and different kinds of deforestation patterns have different border size e.g.
 regular > randon block
-
-```{r SLOSS_lambda14_sims_regular,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
-
-# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
-# hps = habitat patch size
-# dd = dispersal distance 
-# brb = bird rate birds 
-nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3_brb1-4_regular",
-                            outpath=outpath,
-                            repetition=1,
-                            tickmetrics="false",
-                            idsetup="setup",
-                            idgo="go",
-                            runtime=0,
-                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
-                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
-                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
-                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
-                                             ),
-                            constants = list("world-width" = 199,
-                                             "world-height" = 199,
-                                             "initial-population" = 0,
-                                             "max-birds-species" = 100,
-                                             "death-rate-birds" = 1,
-                                             "migration-rate-birds" = 0.0001,
-                                             "Video" = "false",
-                                             "prob-frag" =  0.6,
-                                             "Deforestation-at-time" = 200,
-                                             "end-time" = 1500,
-                                             "Migration0-at-deforestation" = "false",
-                                             "deforestation-type" = "\"regular\"",
-                                             "birds-behavior" = "\"NoSelection\"",
-                                             "replacement-rate" = 0.0
-                                             ))
-
-#
-# Run nseeds=10 times 
-#
-# Total number of jobs = nseeds * split = 20*2 
-# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
-#
-
-nl@simdesign <- simdesign_ff(nl=nl,
-                               nseeds=30)
-
-# run in Paralell 
-#
-plan(multisession)
-tic()
-results <- run_nl_all(nl,split = 2)
-#results <- run_nl_all(nl,split = 4)
-toc()
-plan(sequential)
-names(results)
-
-#
-# Write the output 
-#
-setsim(nl, "simoutput") <- results 
-write_simoutput(nl)
-```
-
-## Read Simulations
-
-```{r readSimulations, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
-# Read previous simulations
-#
-if( file.exists("Simulations/sloss_sim_ff.rds")) {
-
-  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
-
-} else {
-  sloss_sim <- tibble()
-}
-
-# Read netlogo simulations and 
-# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
-# Calculate cluster stats
-#
-# Old 10 - 20 simulations 
-# mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_ff.csv"),skip=0) %>% 
-# mdl <- read_netlogo_simul(file.path(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_20_ff.csv"),skip=0) %>%
-
-mdl <- read_sloss_simul(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")
-
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
-
-netlogo_evaluate_patch_distr(mdl$find_cluster_sizes[1])
-
-```
 
@@
                                                "birth-rate-birds" = list(values=c(1.7, 2, 3, 4)),
-                                             "prob-frag" =  list(values=c(0.4,0.8))
+                                             "prob-frag" =  list(values=c(0.4,0.6,0.8))
                                              ),
@@
                                              "Video" = "false",
-                                             "Deforestation-at-time" = 200,
+                                             "Deforestation-at-time" = 600,
                                              "end-time" = 1500,
@@
 ggsave("Figures/Shannon_MeanPatch_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
-
-```
-
-## Lambda = 1-4 / Random Block patterns
-
-* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )
-
-```{r SLOSS_lambda14_simulations_block,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
-
-# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
-# hps = habitat patch size
-# dd = dispersal distance 
-# brb = bird rate birds 
-nl@experiment <- experiment(expname="NoSelection_pf06_hps3-61_dd1-3_brb1-4_block",
-                            outpath=outpath,
-                            repetition=1,
-                            tickmetrics="false",
-                            idsetup="setup",
-                            idgo="go",
-                            runtime=0,
-                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path", "find-cluster-sizes"),
-                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
-                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
-                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
-                                             ),
-                            constants = list("world-width" = 199,
-                                             "world-height" = 199,
-                                             "initial-population" = 0,
-                                             "max-birds-species" = 100,
-                                             "death-rate-birds" = 1,
-                                             "migration-rate-birds" = 0.0001,
-                                             "Video" = "false",
-                                             "prob-frag" =  0.6,
-                                             "Deforestation-at-time" = 200,
-                                             "end-time" = 1500,
-                                             "Migration0-at-deforestation" = "false",
-                                             "deforestation-type" = "\"random block\"",
-                                             "birds-behavior" = "\"NoSelection\"",
-                                             "replacement-rate" = 0.0
-                                             ))
-
-#
-# Run nseeds=10 times 
-#
-# Total number of jobs = nseeds * split = 30*2 
-# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
-#
-
-nl@simdesign <- simdesign_ff(nl=nl,
-                               nseeds=30)
-
-# run in Paralell 
-#
-plan(multisession)
-tic()
-results <- run_nl_all(nl,split = 2)
-#results <- run_nl_all(nl,split = 4)
-toc()
-plan(sequential)
-names(results)
-
-#
-# Write the output 
-#
-setsim(nl, "simoutput") <- results 
-write_simoutput(nl)
-```
-
-## Read Simulations No Selection Random Block Patterns prob-frag = 0.6
-
-```{r readSimulationsBlock, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
-# Read previous simulations
-#
-if( file.exists("Simulations/sloss_sim_ff.rds")) {
-
-  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
-
-} else {
-  sloss_sim <- tibble()
-}
-
-# Read netlogo simulations and 
-# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
-#
-mdl <- read_sloss_simul(outpath,"NoSelection_pf06_hps3-61_dd1-3_brb1-4_block_ff.csv")
-
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
 
@@
                                                "birth-rate-birds" = list(values=c(1.7, 2, 3, 4)),
-                                               "prob-frag" =  list(values=c(0.4,0.8))
+                                               "prob-frag" =  list(values=c(0.4,0.6,0.8))
                                              ),
@@
 
-```{r Hierarchical_SLOSS_lambda14_pf48_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=TRUE}
+```{r Hierarchical_SLOSS_lambda14_pf48_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
 
@@
 
-# Plot for hierarchical communities
+## Read Simulations hierarchical regular Patterns prob_frag 0.4 - 0.8
+
+```{r readSimulationsHierarchicalRegular_pf48, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_sloss_simul(outpath,"Hierarchical_pf04-08_hps3-61_dd1-3_brb1-4_regular_ff.csv")
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
+
+```
+
+## Simulations Lambda = 1-4 / Random Block - Hierarchical communities / prog-frag 0.4-0.6-0.8
+
+```{r Hierarchical_SLOSS_lambda14_pf48_simulations_block,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="Hierarchical_pf04-08_hps3-61_dd1-3_brb1-4_block",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4)),
+                                             "prob-frag" =  list(values=c(0.4,0.6,0.8))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"random block\"",
+                                             "birds-behavior" = "\"Hierarchical\"",
+                                             "replacement-rate" = 0.3
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Read Simulations hierarchical random-block Patterns prob_frag 0.4 - 0.8
+
+```{r readSimulationsHierarchicalBlock_pf48, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+# Read previous simulations
+#
+if( file.exists("Simulations/sloss_sim_ff.rds")) {
+
+  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
+
+} else {
+  sloss_sim <- tibble()
+}
+
+# Read netlogo simulations and 
+# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
+#
+mdl <- read_sloss_simul(outpath,"Hierarchical_pf04-08_hps3-61_dd1-3_brb1-4_block_ff.csv")
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
+
+```
+
+# Plot for hierarchical communities Regular pattern
 
@@
 mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
-                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(birds_at_deforestation:count_birds,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,prob_frag) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
+
 names(mdl1)
@@
   geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
-  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
   labs(color = "Dispersal\n distance") +
@@
   theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/Hierarchical_Species_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Hierarchical_Species_Dispersal_pf_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
   geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
-  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
   labs(color = "Dispersal\n distance") +
@@
   theme(
-    legend.position = c(0.94,0.83)
-    )
-
-ggsave("Figures/Hierarchical_Shannon_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Hierarchical_Shannon_Dispersal_pf_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+
+```
+
+# Plot for hierarchical communities random block pattern
+
+```{r ExploratoryPlotsHierarchicalBlock, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+
+mdl <- sloss_sim %>% filter(deforestation_type == "random block", birds_behavior=="Hierarchical", habitat_patch_size != 199)
+names(mdl)
+#
+# summarize data 
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,prob_frag) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
+
+names(mdl1)
+
+# require(plotly)
+# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))
+
+mdl1 %>% 
+  ggplot(aes(x=habitat_patch_size, y=calc_number_of_species_mean, color=birds_dispersal_distance)) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds))) +
+  geom_ribbon(
+    aes(
+      ymin = calc_number_of_species_ql,
+      ymax = calc_number_of_species_qh,
+      fill = birds_dispersal_distance
+    ),
+    alpha = 0.3,
+    colour = NA,
+    show.legend = FALSE
+
+  ) +
+  scale_color_viridis_c(guide = FALSE) + 
+  scale_fill_viridis_c() +
+  facet_grid(birds_dispersal_distance ~ birth_rate_birds) + 
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species")
+
+mdl1 %>%  #filter(prob_frag==0.4) %>%
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Hierarchical_Species_Dispersal_pf_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Hierarchical_Shannon_Dispersal_pf_Lambda_block.png",width=8,height=6,units="in",dpi=600)
 
@@
 
-## Read Simulations Birth Selection  regular Patterns
+## Simulations Lambda = 1-4 / Regular patterns - Birth Selection communities / prog frag 0.4-0.8
+
+```{r Selection_SLOSS_lambda14_pf48_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+
+# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="Selection_pf04-08_hps3-61_dd1-3_brb1-4_regular",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4)),
+                                               "prob-frag" =  list(values=c(0.4,0.8))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"regular\"",
+                                             "birds-behavior" = "\"BirthSelection\"",
+                                             "replacement-rate" = 0.3
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Simulations Lambda = 1-4 / random block patterns - Birth Selection communities / prog frag 0.4-0.8
+
+```{r Selection_SLOSS_lambda14_pf48_block,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+
+# pf = probability of fragmentation
+# hps = habitat patch size
+# dd = dispersal distance 
+# brb = bird rate birds 
+nl@experiment <- experiment(expname="Selection_pf04-08_hps3-61_dd1-3_brb1-4_block",
+                            outpath=outpath,
+                            repetition=1,
+                            tickmetrics="false",
+                            idsetup="setup",
+                            idgo="go",
+                            runtime=0,
+                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
+                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
+                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
+                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4)),
+                                               "prob-frag" =  list(values=c(0.4,0.6,0.8))
+                                             ),
+                            constants = list("world-width" = 199,
+                                             "world-height" = 199,
+                                             "initial-population" = 0,
+                                             "max-birds-species" = 100,
+                                             "death-rate-birds" = 1,
+                                             "migration-rate-birds" = 0.0001,
+                                             "Video" = "false",
+                                             "Deforestation-at-time" = 200,
+                                             "end-time" = 1500,
+                                             "Migration0-at-deforestation" = "false",
+                                             "deforestation-type" = "\"random block\"",
+                                             "birds-behavior" = "\"BirthSelection\"",
+                                             "replacement-rate" = 0.3
+                                             ))
+
+#
+# Run nseeds=10 times 
+#
+# Total number of jobs = nseeds * split = 30*2 
+# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
+#
+
+nl@simdesign <- simdesign_ff(nl=nl,
+                               nseeds=30)
+
+# run in Paralell 
+#
+plan(multisession)
+tic()
+results <- run_nl_all(nl,split = 2)
+#results <- run_nl_all(nl,split = 4)
+toc()
+plan(sequential)
+names(results)
+
+#
+# Write the output 
+#
+setsim(nl, "simoutput") <- results 
+write_simoutput(nl)
+```
+
+## Read Simulations Birth Selection  regular & random block Patterns
 
@@
 #
-mdl <- read_sloss_simul(outpath,"Selection_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")
+# Ya leimos este ->
+# mdl <- read_sloss_simul(outpath,"Selection_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")
+#
+# sloss_sim <- bind_rows(sloss_sim,mdl)
+
+mdl <- read_sloss_simul(outpath,"Selection_pf04-08_hps3-61_dd1-3_brb1-4_regular_ff.csv")
+
+sloss_sim <- bind_rows(sloss_sim,mdl)
+
+mdl <- read_sloss_simul(outpath,"Selection_pf04-08_hps3-61_dd1-3_brb1-4_block_ff.csv")
+
+# Check simulations in mdl
+mdl %>% distinct(prob_frag,deforestation_type,birds_behavior)
+
+# Check simulations in sloss_sim
+sloss_sim %>% distinct(prob_frag,deforestation_type,birds_behavior) %>% arrange(prob_frag,deforestation_type,birds_behavior)
 
@@
 mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
-                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds) %>% summarise(across(birds_at_deforestation:count_birds,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,prob_frag) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
+
 names(mdl1)
 
-# require(plotly)
-# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))
-
-mdl1 %>%  
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
+
+#
+mdl1 %>%  #filter(prob_frag==0.4) %>%
   ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
@@
   theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/Selection_Species_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Selection_Species_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
   theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/Selection_Shannon_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Selection_Shannon_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
+
+```
+
+# Plot for birth selection communities random block patterns
+
+```{r ExploratoryPlotsSelectionBlock, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+
+mdl <- sloss_sim %>% filter(deforestation_type == "random block", birds_behavior=="BirthSelection", habitat_patch_size != 199)
+names(mdl)
+#
+# summarize data 
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,prob_frag) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
+
+names(mdl1)
+
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
+
+#
+mdl1 %>%  #filter(prob_frag==0.4) %>%
+  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Selection_Species_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
+mdl1 %>%  
+  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Selection_Shannon_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
 

### 2024-08-02 ms_SLOSS.Rmd
New simulations with deforestation-at-time=600

@@
 several small areas of the same size (SS), despite empirical support favouring SS.
-However, to date, most theories regarding this issue were based on single-species models.
+However, to date, while many studies have explored this issue using model communities, the results are mixed.
+In competitive community models, fragmentation has a negative effect at low habitat amounts, whereas neutral community models suggest positive effects.
 
@@
 
-The SLOSS debate has been approached from several theoretical frameworks, but most of them were based on single-species models [@Fahrig2021].
+The SLOSS debate has been approached from several theoretical frameworks, but many cases have been based based on single-species models [@Fahrig2021].
 Despite the long-standing discussion, there remains a notable scarcity of multi-species models explicitly designed to validate fragmentation theories.
-However, recent studies using spatially explicit simulations revealed that while high fragmentation leads to more local extinctions due to the reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].
+However, recent studies using spatially explicit simulations of competitive communities have revealed that while high fragmentation leads to more local extinctions due to reduced habitat size, at low levels of fragmentation, competitive exclusion is more prevalent, resulting in decreased diversity [@Rybicki2020].
+Conversely, studies based on neutral models have shown that fragmentation can have a positive effect on species richness, particularly when the habitat amount is low [@Smith2024].
+Additionally, spatially implicit models with competition-colonization trade-offs result in nonlinear oscillatory responses to both habitat loss and fragmentation [@Zhang2023].
 
@@
 This comprehensive exploration enables us to account for the interacting effects of habitat loss and a range of model parameters on species retention.
-
-<!-- In that context, spatial-explicit multi-species modelling could be meaningful as several organisms retain dispersal distance and other dispersal-related life traits far consistently among species (quotes).-->
 
@@
 For all simulations, we used a metacommunity of 100 species.
-The simulations start with the colonization of an empty habitat area.
-After 600 time steps, a fixed amount of 60% of habitat is subsequently subjected to destruction, employing regular or random block degradation to mimic habitat fragmentation due to habitat loss.
+The simulations begin with the colonization of an empty habitat area.
+After 600 time steps, a fixed amount of habitat (40%, 60%, or 80%) is subsequently subjected to destruction, employing either regular or random block degradation to mimic habitat fragmentation due to habitat loss.
+The results were collected after 1500 time steps.
 <!-- why we don't use completely random degradation patterns -->

### 2024-09-25 RunSims.Rmd
Updated to submit revision 1

@@
 
-mdl1 %>%  #filter(prob_frag==0.4) %>%
+mdl1 %>%  filter(prob_frag==0.6) %>%
   ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
-  facet_grid(prob_frag ~ birth_rate_birds) +
+  facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  #facet_grid(prob_frag ~ birth_rate_birds) +
   labs(color = "Dispersal\n distance") +
@@
   ) 
-ggsave("Figures/Species_Habitat_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
+ggsave("Figures/Species_Habitat_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 #
-mdl1 %>%  
+mdl1 %>%  filter(prob_frag==0.6) %>%
   ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  #facet_wrap(~ birth_rate_birds, ncol = 4) + 
-  facet_grid(prob_frag ~ birth_rate_birds) +
+  facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  #facet_grid(prob_frag ~ birth_rate_birds) +
   labs(color = "Dispersal\n distance") +
@@
 
-ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
+ggsave("Figures/Shannon_Habitat_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
   facet_grid(prob_frag ~ birth_rate_birds) +
   labs(color = "Dispersal\n distance") + 
+  #,title = "Clasical neutral model with regular habitat degradation") +
   theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
@@
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
+  labs(color = "Dispersal\n distance",
+       title = "Clasical neutral model with regular habitat degradation") +
   theme_bw() + xlab("Mean Patch Size") + ylab("Shannon Diversity") +
@@
 
-ggsave("Figures/Shannon_MeanPatch_Dispersal_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+ggsave("Figures/Shannon_MeanPatch_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 
-```{r SLOSS_lambda14_pf48_block,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=TRUE}
+```{r SLOSS_lambda14_pf48_block,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
 
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance"
+       #,title = "Clasical neutral model with random block habitat degradation"
+       ) +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Species_MeanPatch_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
+
+#
+# shannon_diversity by mean_patch size 
+# 
+mdl1 %>%  
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
   #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
-  theme(
-    legend.position = "bottom",
-    legend.direction = "horizontal"
-  )
-
-ggsave("Figures/Species_MeanPatch_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
-
-#
-# shannon_diversity by mean_patch size 
-# 
-mdl1 %>%  
-  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
-  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
-  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
-  scale_color_viridis_d() + scale_fill_viridis_d() +
-  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
-  facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
+  labs(color = "Dispersal\n distance",
+       title = "Clasical neutral model with random block habitat degradation") +
   theme_bw() + xlab("Mean Patch Size") +  ylab("Shannon Diversity") +
@@
 ggsave("Figures/Species_habitatMFP_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
-
-```
-
-## Simulations Lambda = 1-4 / Regular patterns - Hierarchical communities
-
-* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )
-
-```{r Hierarchical_SLOSS_lambda14_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
-
-# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
-# hps = habitat patch size
-# dd = dispersal distance 
-# brb = bird rate birds 
-nl@experiment <- experiment(expname="Hierarchical_pf06_hps3-61_dd1-3_brb1-4_regular",
-                            outpath=outpath,
-                            repetition=1,
-                            tickmetrics="false",
-                            idsetup="setup",
-                            idgo="go",
-                            runtime=0,
-                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
-                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
-                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
-                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
-                                             ),
-                            constants = list("world-width" = 199,
-                                             "world-height" = 199,
-                                             "initial-population" = 0,
-                                             "max-birds-species" = 100,
-                                             "death-rate-birds" = 1,
-                                             "migration-rate-birds" = 0.0001,
-                                             "Video" = "false",
-                                             "prob-frag" =  0.6,
-                                             "Deforestation-at-time" = 200,
-                                             "end-time" = 1500,
-                                             "Migration0-at-deforestation" = "false",
-                                             "deforestation-type" = "\"regular\"",
-                                             "birds-behavior" = "\"Hierarchical\"",
-                                             "replacement-rate" = 0.3
-                                             ))
-
-#
-# Run nseeds=10 times 
-#
-# Total number of jobs = nseeds * split = 30*2 
-# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
-#
-
-nl@simdesign <- simdesign_ff(nl=nl,
-                               nseeds=30)
-
-# run in Paralell 
-#
-plan(multisession)
-tic()
-results <- run_nl_all(nl,split = 2)
-#results <- run_nl_all(nl,split = 4)
-toc()
-plan(sequential)
-names(results)
-
-#
-# Write the output 
-#
-setsim(nl, "simoutput") <- results 
-write_simoutput(nl)
-```
-
-## Read Simulations hierarchical regular Patterns
-
-```{r readSimulationsHierarchicalRegular, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
-# Read previous simulations
-#
-if( file.exists("Simulations/sloss_sim_ff.rds")) {
-
-  sloss_sim <- readRDS("Simulations/sloss_sim_ff.rds")
-
-} else {
-  sloss_sim <- tibble()
-}
-
-# Read netlogo simulations and 
-# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
-#
-mdl <- read_sloss_simul(outpath,"Hierarchical_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")
-
-sloss_sim <- bind_rows(sloss_sim,mdl)
-
-saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
 
@@
 mdl <- read_sloss_simul(outpath,"Hierarchical_pf04-08_hps3-61_dd1-3_brb1-4_regular_ff.csv")
+
+mdl %>% distinct(prob_frag,deforestation_type,birds_behavior)
+
+sloss_sim %>% distinct(prob_frag,deforestation_type,birds_behavior) %>% arrange(prob_frag,deforestation_type,birds_behavior)
 
@@
 
+sloss_sim %>% distinct(prob_frag,deforestation_type,birds_behavior) %>% arrange(prob_frag,deforestation_type,birds_behavior)
+
 saveRDS(sloss_sim, "Simulations/sloss_sim_ff.rds")
@@
 names(mdl1)
-
-# require(plotly)
-# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))
-
-mdl1 %>% 
-  ggplot(aes(x=habitat_patch_size, y=calc_number_of_species_mean, color=birds_dispersal_distance)) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds))) +
-  geom_ribbon(
-    aes(
-      ymin = calc_number_of_species_ql,
-      ymax = calc_number_of_species_qh,
-      fill = birds_dispersal_distance
-    ),
-    alpha = 0.3,
-    colour = NA,
-    show.legend = FALSE
-
-  ) +
-  scale_color_viridis_c(guide = FALSE) + 
-  scale_fill_viridis_c() +
-  facet_grid(birds_dispersal_distance ~ birth_rate_birds) + 
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species")
-
-mdl1 %>%  #filter(prob_frag==0.4) %>%
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
+
+mdl1 %>%
   ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
@@
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
+  labs(color = "Dispersal\n distance",
+       title = "Competitive hierarchical model with regular habitat degradation") +
   theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
@@
 
-ggsave("Figures/Hierarchical_Species_Dispersal_pf_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+ggsave("Figures/Hierarchical_Species_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
+  labs(color = "Dispersal\n distance",
+       title = "Competitive hierarchical model with regular habitat degradation") +
   theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
@@
 
-ggsave("Figures/Hierarchical_Shannon_Dispersal_pf_Lambda_regular.png",width=8,height=6,units="in",dpi=600)
+ggsave("Figures/Hierarchical_Shannon_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
+
+#
+#
+#
+
+mdl1 %>%
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap( ~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance",
+       title = "Competitive hierarchical model with regular habitat degradation") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Hierarchical_MeanPatch_Species_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
+mdl1 %>%  
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birth_rate_birds), fill = factor(birds_dispersal_distance)), alpha = 0.2,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  #facet_wrap(~ birth_rate_birds, ncol = 4) + 
+  facet_grid(prob_frag ~ birth_rate_birds) +
+  labs(color = "Dispersal\n distance",
+       title = "Competitive hierarchical model with regular habitat degradation") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Hierarchical_MeanPatch_Shannon_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 names(mdl1)
-
-# require(plotly)
-# plot_ly(data=mdl1, x=~birds_dispersal_distance, y=~habitat_patch_size, z=~calc_number_of_species, type="scatter3d", mode="markers", color=~calc_number_of_species, marker = list(size = 1))
-
-mdl1 %>% 
-  ggplot(aes(x=habitat_patch_size, y=calc_number_of_species_mean, color=birds_dispersal_distance)) + 
-  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds))) +
-  geom_ribbon(
-    aes(
-      ymin = calc_number_of_species_ql,
-      ymax = calc_number_of_species_qh,
-      fill = birds_dispersal_distance
-    ),
-    alpha = 0.3,
-    colour = NA,
-    show.legend = FALSE
-
-  ) +
-  scale_color_viridis_c(guide = FALSE) + 
-  scale_fill_viridis_c() +
-  facet_grid(birds_dispersal_distance ~ birth_rate_birds) + 
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species")
-
-mdl1 %>%  #filter(prob_frag==0.4) %>%
-  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
+
+mdl1 %>%  
+#  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
-  theme(
-    legend.position = "bottom",
-    legend.direction = "horizontal"
-  )
-
-ggsave("Figures/Hierarchical_Species_Dispersal_pf_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+  labs(color = "Dispersal\n distance",
+       title = "Competitive hierarchical model with random block habitat degradation") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Hierarchical_MeanPatch_Species_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
 
@@
 mdl1 %>%  
-  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+#  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
-  theme(
-    legend.position = "bottom",
-    legend.direction = "horizontal"
-  )
-
-ggsave("Figures/Hierarchical_Shannon_Dispersal_pf_Lambda_block.png",width=8,height=6,units="in",dpi=600)
+  labs(color = "Dispersal\n distance",
+       title = "Competitive hierarchical model with random block habitat degradation") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Hierarchical_MeanPatch_Shannon_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
 
@@
 
-* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 200 for lambda=4 )
+* Initial conditions: migration from the regional community and after the area is filled of individuals, then 60% of the area degraded with regular patches (t= 600 )
 
 ```{r Selection_SLOSS_lambda14_simulations_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
-
-# pf = probability of fragmentation, hps = habitat patch size , dd=dispersal distance 
-# hps = habitat patch size
-# dd = dispersal distance 
-# brb = bird rate birds 
-nl@experiment <- experiment(expname="Selection_pf06_hps3-61_dd1-3_brb1-4_regular",
-                            outpath=outpath,
-                            repetition=1,
-                            tickmetrics="false",
-                            idsetup="setup",
-                            idgo="go",
-                            runtime=0,
-                            metrics=c("birds-at-deforestation", "calc-number-of-species","calc-shannon-diversity","count birds", "mean-free-path","find-cluster-sizes"),
-                            variables = list("birds-dispersal-distance" = list(values=c(1.1, 1.2, 1.5, 2, 3)),
-                                              "habitat-patch-size" = list(values=c(3, 9, 29, 61)),
-                                               "birth-rate-birds" = list(values=c(1.7, 2, 3, 4))
-                                             ),
-                            constants = list("world-width" = 199,
-                                             "world-height" = 199,
-                                             "initial-population" = 0,
-                                             "max-birds-species" = 100,
-                                             "death-rate-birds" = 1,
-                                             "migration-rate-birds" = 0.0001,
-                                             "Video" = "false",
-                                             "prob-frag" =  0.6,
-                                             "Deforestation-at-time" = 200,
-                                             "end-time" = 1500,
-                                             "Migration0-at-deforestation" = "false",
-                                             "deforestation-type" = "\"regular\"",
-                                             "birds-behavior" = "\"BirthSelection\"",
-                                             "replacement-rate" = 0.3
-                                             ))
-
-#
-# Run nseeds=10 times 
-#
-# Total number of jobs = nseeds * split = 30*2 
-# Each job runs parameter input matrix / split = nrow(nl@simdesign@siminput) / split =  80 / 2 
-#
-
-nl@simdesign <- simdesign_ff(nl=nl,
-                               nseeds=30)
-
-# run in Paralell 
-#
-plan(multisession)
-tic()
-results <- run_nl_all(nl,split = 2)
-#results <- run_nl_all(nl,split = 4)
-toc()
-plan(sequential)
-names(results)
-
-#
-# Write the output 
-#
-setsim(nl, "simoutput") <- results 
-write_simoutput(nl)
-```
-
-## Simulations Lambda = 1-4 / Regular patterns - Birth Selection communities / prog frag 0.4-0.8
-
-```{r Selection_SLOSS_lambda14_pf48_regular,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
 
@@
                                                "birth-rate-birds" = list(values=c(1.7, 2, 3, 4)),
-                                               "prob-frag" =  list(values=c(0.4,0.8))
+                                              "prob-frag" =  list(values=c(0.4,0.6,0.8))
                                              ),
@@
 
-```{r Selection_SLOSS_lambda14_pf48_block,echo=FALSE,message=FALSE,warning=FALSE,eval=FALSE}
+```{r Selection_SLOSS_lambda14_pf48_block,echo=FALSE,message=FALSE,warning=FALSE,eval=TRUE}
 
@@
 # Read netlogo simulations and
-# Convert the list of mean_free_path to degraded_mfp and non_degraded_mfp
-#
-# Ya leimos este ->
-# mdl <- read_sloss_simul(outpath,"Selection_pf06_hps3-61_dd1-3_brb1-4_regular_ff.csv")
-#
-# sloss_sim <- bind_rows(sloss_sim,mdl)
 
@@
 #
-mdl1 %>%  #filter(prob_frag==0.4) %>%
-  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+mdl1 %>%  
+#  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
-  theme(
-    legend.position = "bottom",
-    legend.direction = "horizontal"
-  )
-
-ggsave("Figures/Selection_Species_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
+  labs(color = "Dispersal\n distance",
+       title = "Neutral w/Habitat selection model and regular habitat degradation") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Selection_MeanPatch_Species_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 mdl1 %>%  
-  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  #ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
-  theme(
-    legend.position = "bottom",
-    legend.direction = "horizontal"
-  )
-
-ggsave("Figures/Selection_Shannon_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
+  labs(color = "Dispersal\n distance",
+       title = "Neutral w/Habitat selection model and regular habitat degradation") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Selection_MeanPatch_Shannon_Dispersal_Lambda_pf_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 #
-mdl1 %>%  #filter(prob_frag==0.4) %>%
-  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+mdl1 %>% 
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
-  theme(
-    legend.position = "bottom",
-    legend.direction = "horizontal"
-  )
-
-ggsave("Figures/Selection_Species_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
+  labs(color = "Dispersal\n distance",
+       title = "Neutral w/Habitat selection model and random block habitat degradation") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Selection_MeanPatch_Species_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
 
@@
 mdl1 %>%  
-  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   facet_grid(prob_frag ~ birth_rate_birds) +
-  labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
-  theme(
-    legend.position = "bottom",
-    legend.direction = "horizontal"
-  )
-
-ggsave("Figures/Selection_Shannon_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
-
-```
-
-## Plots across neutral hierarchical selection modelos regular fragmentation
+  labs(color = "Dispersal\n distance",
+       title = "Neutral w/Habitat selection model and random block habitat degradation") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/Selection_MeanPatch_Shannon_Dispersal_Lambda_pf_block.png",width=8,height=6,units="in",dpi=600)
+
+```
+
+## Plots across neutral hierarchical selection model regular fragmentation
 
@@
 mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
-                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,birds_behavior) %>% summarise(across(birds_at_deforestation:count_birds,list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,birds_behavior,prob_frag) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) )))
+
 names(mdl1)
@@
 mdl1 <- mdl1 %>% mutate(birds_behavior= factor(birds_behavior, levels = behavior_order, labels = behavior_labels))
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
 
 mdl1 %>%  filter(birth_rate_birds == 2) %>% 
-  ggplot(aes(x = habitat_patch_size, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ birds_behavior, ncol = 4) + 
+  facet_grid(prob_frag ~ birds_behavior) +
   labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Number of species") +
-  theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/AcrossBehav_Species_Dispersal_Lambda2_regular.png",width=8,height=6,units="in",dpi=600)
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+ggsave("Figures/AcrossBehav_MeanPatch_Species_Dispersal_Lambda2_pf_regular.png",width=8,height=6,units="in",dpi=600)
 
@@
 mdl1 %>%  filter(birth_rate_birds == 2) %>% 
-  ggplot(aes(x = habitat_patch_size, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ birds_behavior, ncol = 4) + 
+  facet_grid(prob_frag ~ birds_behavior) +
   labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Habitat Patch Size") + ylab("Shannon Diversity") +
-  theme(
-    legend.position = c(0.85,0.18)
-    )
-
-ggsave("Figures/AcrossBehav_Shannon_Dispersal_Lambda2_regular.png",width=8,height=6,units="in",dpi=600)
+  theme_bw() + xlab("Mean Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/AcrossBehav_MeanPatch_Shannon_Dispersal_Lambda2_pf_regular.png",width=8,height=6,units="in",dpi=600)
+
+```
+
+## Plots across neutral hierarchical selection modelos random block fragmentation
+
+```{r PlotsAcrossBehaviorblock, echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',eval=FALSE}
+
+mdl <- sloss_sim %>% filter(deforestation_type == "random block", habitat_patch_size != 199)
+names(mdl)
+#
+# summarize data 
+#
+mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,birds_behavior,prob_frag) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+names(mdl1)
+behavior_order <- c("NoSelection", "Hierarchical", "BirthSelection")
+behavior_labels <- c("Neutral", "Hierarchical", "Habitat Selection")
+
+mdl1 <- mdl1 %>% mutate(birds_behavior= factor(birds_behavior, levels = behavior_order, labels = behavior_labels))
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
+
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_number_of_species_ql, ymax = calc_number_of_species_qh, group = interaction(birds_dispersal_distance, birds_behavior), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_grid(prob_frag ~ birds_behavior) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/AcrossBehav_MeanPatch_Species_Dispersal_Lambda2_pf_block.png",width=8,height=6,units="in",dpi=600)
+
+# Shannon-Diversity  
+#
+mdl1 %>%  filter(birth_rate_birds == 2) %>% 
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
+  geom_point(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = 1.2) +
+  geom_ribbon(aes(ymin = calc_shannon_diversity_ql, ymax = calc_shannon_diversity_qh, group = interaction(birds_dispersal_distance, birds_behavior), fill = factor(birds_dispersal_distance)), alpha = 0.3,colour=NA, show.legend = FALSE) +
+  scale_color_viridis_d() + scale_fill_viridis_d() +
+  facet_grid(prob_frag ~ birds_behavior) +
+  labs(color = "Dispersal\n distance") +
+  theme_bw() + xlab("Mean Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/AcrossBehav_MeanPatch_Shannon_Dispersal_Lambda2_pf_block.png",width=8,height=6,units="in",dpi=600)
 
@@
 mdl1 <- mdl %>% mutate(habitat_patch_size = habitat_patch_size+1) %>%
-                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,deforestation_type) %>% summarise(across(c(birds_at_deforestation:count_birds,max_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
+                         group_by(birds_dispersal_distance, habitat_patch_size,birth_rate_birds,deforestation_type,prob_frag) %>% summarise(across(c(birds_at_deforestation:count_birds,mean_patch,degraded_mfp,non_degraded_mfp),list(mean=mean, sd=sd, ql=~quantile(.x,probs=c((1-0.95)/2)), qh=~quantile(.x,probs=c(1-(1-0.95)/2)) ))) #  %>% mutate(deltaPop = (count_birds - birds_at_deforestation )/ birds_at_deforestation  )
 names(mdl1)
 
-mdl1 <- mdl1 %>%  mutate(across(max_patch_mean:max_patch_qh,sqrt))
+mdl1 <- mdl1 %>%  mutate(across(mean_patch_mean:mean_patch_qh,sqrt))
 
 mdl1 %>%  filter(birth_rate_birds == 2) %>% 
-  ggplot(aes(x = max_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
+  ggplot(aes(x = mean_patch_mean, y = calc_number_of_species_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ deforestation_type, ncol = 2) + 
+  facet_grid(prob_frag ~ deforestation_type) +
+  #facet_wrap(~ deforestation_type, ncol = 2) + 
   labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Max Habitat Patch Size") + ylab("Number of species") +
-  theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/AcrossLoss_Species_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
+  theme_bw() + xlab("Mean Patch Size") + ylab("Number of species") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/AcrossLoss_Species_Dispersal_Lambda2_pf.png",width=8,height=6,units="in",dpi=600)
 
@@
 mdl1 %>%  filter(birth_rate_birds == 2) %>% 
-  ggplot(aes(x = max_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
+  ggplot(aes(x = mean_patch_mean, y = calc_shannon_diversity_mean, color = factor(birds_dispersal_distance))) + 
   geom_line(aes(group = interaction(birds_dispersal_distance, birth_rate_birds)),  size = .5) +
@@
   scale_color_viridis_d() + scale_fill_viridis_d() +
-  facet_wrap(~ deforestation_type, ncol = 4) + 
+  facet_grid(prob_frag ~ deforestation_type) +
   labs(color = "Dispersal\n distance") +
-  theme_bw() + xlab("Max Habitat Patch Size") + ylab("Shannon Diversity") +
-  theme(
-    legend.position = c(0.9,0.8)
-    )
-
-ggsave("Figures/AcrossLoss_Shannon_Dispersal_Lambda2.png",width=8,height=6,units="in",dpi=600)
+  theme_bw() + xlab("Mean Habitat Patch Size") + ylab("Shannon Diversity") +
+  theme(
+    legend.position = "bottom",
+    legend.direction = "horizontal"
+  )
+
+ggsave("Figures/AcrossLoss_Shannon_Dispersal_Lambda2_pf.png",width=8,height=6,units="in",dpi=600)
 
@@
 
-image_paths <- c("Figures/NoSelection_lambda2_dd15_ps3_before_nsp37.png","Figures/NoSelection_lambda2_dd15_ps3_after_nsp32.png","Figures/NoSelection_lambda2_dd15_ps3_after_lt_nsp3.png",
-                 "Figures/NoSelection_lambda2_dd15_ps29_ before_nsp36.png","Figures/NoSelection_lambda2_dd15_ps29_ after_nsp29.png","Figures/NoSelection_lambda2_dd15_ps29_ after_lt_nsp25.png")
+image_paths <- c("Figures/NoSelection_lambda2_dd15_ps3_before_nsp37.png",
+                 "Figures/NoSelection_lambda2_dd15_ps3_after_nsp32.png",
+                 "Figures/NoSelection_lambda2_dd15_ps3_after_lt_nsp3.png",
+                 "Figures/NoSelection_lambda2_dd15_ps29_ after_lt_nsp25.png",
+                 "Figures/NoSelection_lambda2_dd15_ps3_before_block_nsp36.png",
+                 "Figures/NoSelection_lambda2_dd15_ps3_after_block_nsp35.png",
+                 "Figures/NoSelection_lambda2_dd15_ps3_after_block_lt_nsp16.png",
+                 "Figures/NoSelection_lambda2_dd15_ps29_ after_block_lt_nsp12.png")
 
@@
 aplots <- plot_grid(plotlist = plots, ncol = 2,
-  labels = c('A.', 'B.', 'C.', 'D.','E.','F.'),
+  labels = c('A.', 'B.', 'C.', 'D.','E.','F.','G.','H.'),
   label_fontface = "plain",

### 2024-09-25 ms_SLOSS.Rmd
Updated to submit revision 1

@@
   pdf_document:
+    keep_tex: true
     includes:
       in_header: margins.sty
-  word_document: default
 ---
@@
 In the classical neutral model, species are equivalent and have the same parameters; when there is a reproduction event they send the propagule to a surrounding patch blindly, and if the patch is already occupied the propagule dies.
-In the neutral-with-habitat selection model, the propagules select an empty patch so their survival is ensured.
+In the neutral with habitat selection model, the propagules select an empty patch so their survival is ensured.
 In the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant with a replacement rate parameter.
@@
 
-    Regular degradation: Habitat is removed so that habitable patches form a regular pattern of squares with sides given by a habitat-size parameter.
-As the percentage of degraded habitat is constant varying the habitat-size parameter results in either a few large habitat patches (SL) or several small (SS).
+    Regular degradation: Habitat is removed in a way that results in habitable patches forming a regular pattern of squares, with side lengths determined by a habitat-size parameter.
+While the percentage of degraded habitat remains constant, varying the habitat-size parameter leads to either a few large patches (SL) or many smaller ones (SS).
+The algorithm adjusts either the number or the size of patches to maintain the constant percentage of degradation, especially when only one patch fits within the grid.
+We quantify this pattern using the mean habitat patch size.
 
@@
 The sizes are identical to those used for regular degradation simulations, but it results in habitat patches that have a minimum size corresponding to the habitat-size but a maximum size that is bigger because habitat patches can overlap between them producing a larger patch.
-Then, we characterize this with the maximum habitat patch size.
+Then, we characterize this with the mean habitat patch size.
 
@@
 In similar non-spatial models, this relationship must exceed 1 for a species to persist.
-Thus we used a minimum $\lambda$ of 1.7, we also fixed the $migration-rate$ at 0.0001.
-We simulate each model variant using the parameters shown in Table 1.
-For the neutral classical version only, we also simulate random block habitat patterns.
+Thus we used a minimum $\lambda$ of 1.7 and fixed the migration rate at 0.0001.
+We simulated each model variant using the parameters shown in Table 1, with both regular and random block habitat patterns.
 For each combination of parameters we ran 30 simulations up to 1500 time steps and recorded the number of species, the Shannon Diversity index [@Magurran2013] and the total number of individuals.
@@
 
-In our neutral model simulations with a migration rate of $0.0001$ and regular habitat patterns, species richness varied with habitat patch sizes, influenced by dispersal distance and the $\lambda$ parameter (Figure 1).
+The spatial pattern of habitat loss is illustrated in Figure 1.
+Simulations with regular degradation (Figures 1C, 1E, and 1G) produce habitat patches of uniform size, evenly spaced across the landscape.
+In contrast, simulations using randomly arranged block patterns result in patches of variable sizes, where the habitat-size parameter defines the minimum size, and superposition of patches creates a larger maximum size (Figures 1D, 1F, and 1H).
+Consequently, the mean patch size is higher for random block degradation than for regular degradation.
+In some cases, to achieve the targeted percentage of habitat loss, the effective patch size in regular degradation exceeds the defined habitat-size parameter (Figure 2).
+Simulations with the neutral model illustrate species' spatial distributions before fragmentation (Figures 1A & 1B) and after fragmentation (Figures 1C–F), with habitat destruction depicted in purple.
+Immediately after fragmentation, species distributions closely resemble pre-fragmentation patterns, but diverge over time.
+
+![Examples of species spatial patterns before and after fragmentation for the neutral model under different habitat loss patterns.
+Different shades from dark green to white represent distinct species, while destroyed habitat is shown in purple.
+Simulations were run with a fixed growth rate ($\lambda=2$), a migration rate of 0.0001, and a dispersal distance of 1.5.
+Panels A & B show spatial patterns before fragmentation.
+Panels C & D depict patterns immediately after 60% habitat destruction, with patch sizes of 3 and regular degradation (C) or random block degradation (D).
+Panels E & F illustrate spatial patterns after 1500 time steps, showing population extinction with regular degradation (E).
+Panels G & H display the spatial pattern after 1500 time steps with a habitat patch size of 29.](Figures/ModelSpatialPatternComparison.png)
+
+In our neutral model simulations with a migration rate of $0.0001$ and regular habitat degradation patterns, species richness varied with habitat patch sizes, influenced by dispersal distance and the $\lambda$ parameter (Figure 2).
 Species richness increased with $\lambda$, with the highest richness at smaller patch sizes for higher $\lambda$ values.
-Conversely, a $\lambda$ of 3 or less showed reduced richness at the smallest patch sizes for lower dispersal distances.
-In these cases, only migrants survive in the smallest patches, and no local populations are established (Figures S13 & S14).
-The interplay of dispersal distances, habitat sizes, and $\lambda$ impacts community survival and species richness.
+Conversely, a $\lambda$ less than 3 showed reduced richness at the smallest patch sizes for lower dispersal distances.
+In these cases, only migrants survive in the smallest patches, and no local populations are established (Figures 1E, and S19 & S20).
+
+The interplay of dispersal distances, mean patch sizes, $\lambda$ and percent of habitat destruction impacts community survival and species richness.
 For $\lambda < 4$ this suggests a threshold patch size influenced by these 3 parameters, for $\lambda = 4$ patch size decreases species richness.
@@
 Extended dispersal distances could enhance species survival in scenarios with low $\lambda$ and smaller patches but reduce richness beyond a certain threshold.
-Similar patterns are evident for Shannon diversity (Figure 2).
+Similar patterns are evident for Shannon diversity (Figure S1 & S2).
 
-![Relationship between the number of species and habitat patch size for varying dispersal distances and $\lambda$ parameter (across panels).
+![Relationship between the number of species and mean habitat patch size across varying dispersal distances.
+The horizontal panels show the relationship between growth rate and death rate, parameter $\lambda$, while the vertical panels represent different percentages of habitat destruction.
 Simulations were performed using a neutral model with a constant migration rate (0.0001).
-Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
-The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.](Figures/Species_Habitat_Dispersal_Lambda_regular.png)
+After 600 time steps, the specified percentage of the habitat was destroyed, making it unavailable for species colonization.
+The habitat destruction follows a regular pattern, with all patches being of equal size and uniformly spaced from one another.](Figures/Species_MeanPatch_Dispersal_Lambda_pf_regular.png)
 
-![Relationship between Shannon Diversity and habitat patch size for varying dispersal distances and $\lambda$ parameter (across panels).
-Simulations were performed using a neutral model with a constant migration rate (0.0001).
-Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
-The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.](Figures/Shannon_Habitat_Dispersal_Lambda_regular.png)
+![Relationship between the number of species and mean habitat patch size across varying dispersal distances for the neutral model.
+The horizontal panels show the relationship between growth rate and death rate, parameter $\lambda$, while the vertical panels represent different percentages of habitat destruction.
+Simulations were performed with a constant migration rate (0.0001).
+After 600 time steps, the specified percentage of the habitat was destroyed, making it unavailable for species colonization.
+Habitat blocks of the same size (habitat patch size) were randomly distributed over the simulation grid.
+As a result, some patches overlapped, leading to a distribution of patch sizes.](Figures/Species_MeanPatch_Dispersal_Lambda_pf_block.png)
 
-For the simulations conducted with randomly arranged block habitat patterns, the patches exhibit variable sizes, following a distribution with a minimum size determined by the patch-size parameter and a larger maximum size determined by the superposition of several patches (Figures S9 - S11).
-The overall pattern is analogous to the regular case, with a peak in richness occurring for intermediate patch sizes.
-However, for values of $\lambda$ higher than 2, the maximum richness is observed at the smaller habitat patch size (Figure S1 & S2).
-The patches in the random block simulations are generally larger, resulting in lower species richness and diversity, except for the minimum patch size (Figure S7 & S8).
-Supplementary material (Figures S9 & S32) includes examples of simulations with random block and regular habitat patterns.
-In these figures, it is evident that for the minimum patch size, only migrants survive in regular habitat patterns with neutral and hierarchical models.
+For random block habitat degradation the overall pattern is analogous to the regular case (Figure 3), with a peak in richness occurring for the second patch sizes, for $\lambda$ less than 3.
+However, for values of $\lambda$ greater than 2, maximum richness shifts to the smallest habitat patch size.
+Patches in the random block simulations are generally larger than in regular degradation, leading to lower species richness and diversity, except at the minimum patch size (Figure 3).
+Simulations comparing random block and regular habitat patterns (Figure 1) show that, for the smallest patch size, only migrant species persist in regular patterns under neutral and hierarchical models (Figures S19, S20, S27, S28), but not in the neutral model with habitat selection.
 
-![Relationship between the number of species and habitat patch size for different versions of the model (across panels) and varying dispersal distances.
-The neutral is the classical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat-selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
-Simulations were performed using a fixed growth rate (2), migration and replacement rate of 0.0001 and 0.3 respectively.
-Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
-The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.](Figures/AcrossBehav_Species_Dispersal_Lambda2_regular.png)
+![Relationship between the number of species and habitat patch size across panels representing different model versions and varying dispersal distances.
+The vertical panels correspond to different percentages of habitat destruction.
+In the neutral model, species are equivalent and share the same parameters; in the hierarchical model, species exhibit competitive dominance, with more dominant species replacing less dominant ones; and in the neutral with habitat selection model, species are neutral but propagules search for empty patches, ensuring survival.
+Simulations were performed using a fixed growth rate ($\lambda = 2$), with migration and replacement rates of 0.0001 and 0.3, respectively.
+After 600 time steps, the specified percentage of the habitat was destroyed, making it unavailable for species.
+The habitat destruction follows a regular pattern, with all patches being of equal size and uniformly spaced from one another.](Figures/AcrossBehav_MeanPatch_Species_Dispersal_Lambda2_pf_regular.png)
 
-![Relationship between Shannon Diversity and habitat patch size for different versions of the model (across panels) and varying dispersal distances.
-The neutral is the classical neutral model, species are equivalent and have the same parameters; in the hierarchical model, species have competitive dominance and more dominant ones can replace the less dominant; the neutral-with-habitat-selection model, is a neutral model but propagules search an empty patch so their survival is ensured.
-Simulations were performed using a fixed growth rate (2), migration and replacement rate of 0.0001 and 0.3 respectively.
-Following 200 time steps of the model, 60% of the habitat was destroyed, rendering it unavailable for the species.
-The destruction of habitat resulted in the formation of regular patterns, where all remaining patches exhibited equal sizes.](Figures/AcrossBehav_Shannon_Dispersal_Lambda2_regular.png)
+![Relationship between the number of species and habitat patch size across panels representing different model versions and varying dispersal distances.
+The vertical panels correspond to different percentages of habitat destruction.
+In the neutral model, species are equivalent and share the same parameters; in the hierarchical model, species exhibit competitive dominance, with more dominant species replacing less dominant ones; and in the neutral with habitat selection model, species are neutral but propagules search for empty patches, ensuring survival.
+Simulations were performed using a fixed growth rate ($\lambda = 2$), with migration and replacement rates of 0.0001 and 0.3, respectively.
+After 600 time steps, the specified percentage of the habitat was destroyed, making it unavailable for species.
+Habitat blocks of the same size (habitat patch size) were randomly distributed over the simulation grid.
+As a result, some patches overlapped, leading to a distribution of patch sizes.](Figures/AcrossBehav_MeanPatch_Species_Dispersal_Lambda2_pf_block.png)
 
-Comparing the three versions of the model at $\lambda =2$, for higher dispersal the neutral model tends to have higher richness and Shannon diversity at higher patch sizes, followed by the hierarchical model and the neutral-with-habitat-selection model (Figures 3 & 4).
-At the lowest patch size, only in the neutral-with-habitat-selection have populations that can survive, because in the other model versions the individuals observed are migrants that cannot establish local populations and become extinct (See Figures S5 - S15).
-Thus an intermediate patch size with a maximum species richness is not observed for this model.
-Is also clearly observed that for the hierarchical model, there is a strong dominance of the more competitive species where Shannon Diversity is very low for a higher dispersal rate (Figure 4 & S23).
+When comparing the three model versions at $\lambda = 2$, the neutral model exhibits higher species richness at larger patch sizes, followed by the neutral with habitat selection model and the hierarchical model (Figures 4 & 5).
+A similar pattern is seen for Shannon diversity (Figures S11 & S12).
+At smaller patch sizes, the neutral with habitat selection model supports populations in all cases, whereas in the other models with low dispersal and $\lambda$ below 3, only migrants are present, unable to establish local populations and leading to extinction (Figures 1, S3 & S5).
+Consequently, intermediate patch sizes with maximum species richness are not observed in the habitat selection model.
+Additionally, the hierarchical model shows a strong dominance of more competitive species, resulting in very low Shannon diversity at higher dispersal rates (Figures S4 & S6).
 
-Examples of the spatial patterns of species and habitat distributions are illustrated in Figures 5 and S9-S32.
-Simulations with the neutral model depict the species' spatial distribution before fragmentation (Figure 5 A & B) and after fragmentation (Figure 5 C to F), with habitat destruction represented in purple areas.
-Immediately after fragmentation, the species' spatial distributions closely resemble those before habitat destruction, but then the patch compositions diverge.
-The divergence in patch species composition is more pronounced in the hierarchical model, where small patches tend to be dominated by one species (Figures S17 - S24).
-The effect of the minimum patch size for survival is also observed (Figure 5 E).
+More examples of species and habitat spatial patterns are illustrated in Figures S15–S38.
+In the hierarchical model, differences between simulations with varying dispersal distances are evident.
+Individual patches are often dominated by a single species, and with short dispersal distances, regions where different species dominate can be established (Figures S23–S30).
+The impact of minimum patch size on species survival is also clear (Figure 5E, S19, S20, S27, & S28).
 
-![Examples of species spatial patterns before and after fragmentation for the neutral model are illustrated.
-Different shades from dark green to white represent distinct species, with the destroyed habitat depicted in purple.
-Simulations were conducted with a fixed growth rate ($\lambda=2$), a migration rate of 0.0001, and a dispersal distance of 1.5.
-A.
-& B.
-showcase spatial patterns before fragmentation.
-C.
-& D.
-depict spatial patterns after fragmentation, where 60% of the habitat was destroyed, with patch sizes of 3 (C) and 29 (D).
-In these cases, the significance of the species' spatial distribution before fragmentation becomes evident.
-E.
-& F.
-display spatial patterns after 600 time steps, illustrating the extinction of populations at smaller patch sizes (E).](Figures/ModelSpatialPatternComparison.png)
+The primary difference between varying percentages of habitat destruction is that lower percentages result in a higher number of species in the smallest patch sizes, particularly at higher dispersal distances.
+In the regular habitat destruction pattern, the increased number of patches shortens the distance between them, facilitating the persistence of local populations.
+In the random block pattern, patches are generally larger due to overlap, which increases the minimum patch size, allowing local populations to persist as well (Figures 2, 3, 4 & 5).
 
@@
 A critical factor in this dynamic is the balance between reproduction and extinction rates ($\lambda$ parameter) and dispersal relative to the size of the habitat patches, influencing their role as sources of propagules.
-Notably, the SL > SS scenario emerges only when dispersal limitations prevent patch connectivity, and patch sizes are above a certain threshold to prevent local extinctions, aligning with Rybicki's [-@Rybicki2020] findings.
+Notably, the SL > SS scenario emerges only when dispersal limitations prevent patch connectivity, and $\lambda$ is relatively small so patch sizes above a certain threshold are needed to prevent local extinctions, aligning with Rybicki's [-@Rybicki2020] findings.
+In these cases, an unimodal relationship between species diversity and patch size is observed, with maximum species diversity occurring at larger patch sizes for shorter dispersal distances.
+Notably, habitat destruction levels do not substantially alter this pattern.
 
-In neutral models, where species are considered equivalent, one species, by chance, tends to dominate a patch (Hubbell, 2001).
+In neutral models, where species are considered equivalent, one species, by chance, tends to dominate a patch [@Hubbell2001].
 Consequently, in fragmented landscapes, once a species increases its density, the likelihood of another species replacing it is low.
@@
 
-Edge effects are present in the neutral and hierarchical models, but not in the neutral-with-habitat-selection model.
+Edge effects are present in the neutral and hierarchical models, but not in the neutral with habitat selection model.
 In the first two models, the edge effect is because individuals in the patch's border have fewer neighbors and the probability of extinction is higher.
-In the neutral-with-habitat-selection model, the edge effect is not present because the individuals search for empty patches, so the individuals on the border of the patch have the same probability of extinction as the individuals in the interior of the patch.
+In the neutral with habitat selection model, the edge effect is not present because the individuals search for empty patches, so the individuals on the border of the patch have the same probability of extinction as the individuals in the interior of the patch.
 Thus when patches are smaller the edge proportion is higher and the edge effect is more important, but this plays a more important role in the minimal patch size for the survival of species than for the species richness.
 
-The spatial arrangement of habitat patches, contrary to some expectations, did not produce different species richness patterns regarding habitat patch size.
-The exceptions are stressed communities---represented in our models with lowest $\lambda$--- with low dispersal may have a continuous increment of richness with patch size, observed only in random block degradation patterns.
-This could be related to the optimal patch size for random block habitat loss patterns, which tends to be higher than for regular patterns, so we could hypothesize that the optimal was outside the simulated range for these stressed communities.
+The spatial arrangement of habitat patches, contrary to some expectations, did not produce different species richness patterns when habitat patch size is accounted.
+The exceptions are stressed communities---represented in our models with lowest $\lambda$--- with low dispersal may have a continuous increment of richness with wider increment of patch sizes, observed only in random block degradation patterns.
 The aggregation of patches in the random block pattern (implies aggregation of the degraded habitat) could be positive if individual habitat patches are very small because it allows some patches to surpass the minimum size required for survival.
@@
 When habitat amounts are lower (e.g., 7%), the arrangement of patches appears to have no significant effect.
-Interestingly, our study replicated similar patterns even with a 40% habitat amount.
+Interestingly, our study replicated similar patterns even within a range of 60 to 20% habitat amount.
 Moreover, other studies employing comparable habitat amounts (32%) also reported higher species richness with increased fragmentation [@Rybicki2020].

### 2024-09-26 ms_SLOSS.Rmd
Submited version

@@
 
-While the overall habitat loss is widely accepted as a process leading to long-term biodiversity loss [@Keil2015;@Hodgson2011], the proposition that single large reserves retain more species than several small reserves of the same area (SL > SS) remain at a similar hierarchical state, even after several criticisms (reviewed in  @Fahrig2017).
+While the overall habitat loss is widely accepted as a process leading to long-term biodiversity loss [@Keil2015;@Hodgson2011], the proposition that single large reserves retain more species than several small reserves of the same area (SL > SS) remain at a similar hierarchical state, even after several criticisms [reviewed in  @Fahrig2017].
 Empirical observational and experimental data shows the predominance of the opposite pattern (SS > SL) ever since the 80s and this fact has been reinforced to this day [@Simberloff1982; @Fahrig2020; @Hammill2020].
@@
 The principal objective is to verify the conditions under which the SS > SL pattern holds and discern the underlying causes.
-Through this examination, we seek to contribute insights that can help determine if the conditions outlined in the SLOSS cube hypothesis are supported by our modeling approach..
+Through this examination, we seek to contribute insights that can help determine if the conditions outlined in the SLOSS cube hypothesis are supported by our modeling approach.
 
@@
 In neutral models, where species are considered equivalent, one species, by chance, tends to dominate a patch [@Hubbell2001].
-Consequently, in fragmented landscapes, once a species increases its density, the likelihood of another species replacing it is low.
+Consequently, in fragmented landscapes, once a species increases its density, the likelihood of another species replacing is low.
 With more patches, the probability of having a greater number of species is heightened.
@@
 Naturally, there are limitations to our modeling work.
-Firstly, it is important to note that more research is needed to apply our to determine the effect of fragmentation in real-world communities.
+Firstly, it is important to note that more research is needed to apply our model to determine the effect of fragmentation in real-world communities.
 We assume that all species have the same parameters, thus hypothetically determining these parameters for all the species, and using the mean to determine the responses to fragmentation would render an approximate value for the effects and thresholds.

### 2024-10-09 ms_SLOSS.Rmd
Updated title and Readme with zenodo

@@
 ---
+
+Leonardo A. Saravia ^1^^2^, Ulises Balza ^1^, Fernando Momo^3^^4^
+
+1. Centro Austral de Investigaciones Científicas (CADIC-CONICET). Houssay 200 (9410), Ushuaia, Argentina.
+
+2. Universidad Nacional de Tierra del Fuego (UNTDF), Yrigoyen 879 (9410), Ushuaia, Argentina
+
+3. Instituto de Ciencias. Universidad Nacional de General Sarmiento. J.M. Gutierrez 1159 (1613), Los Polvorines, Argentina.
+
+4. Departamento de Ciencias Básicas. Universidad Nacional de Luján. Rutas 5 y 7, (6700) Luján. Argentina.
 
@@
 
-The model was developed in Netlogo 6.3.0 and the model source code is available at Github <https://>, the Overview, Design concepts, Details (ODD), protocol for describing individual-based models is included within the model source code [@Grimm2010].
-We performed the simulations and analysis using R version 4.3.1 [@RCoreTeam2023], the R source code to perform simulations and analysis is located at <https://>.
+The model was developed in Netlogo 6.3.0 and the model source code is available at Github <https://github.com/lsaravia/DynamicForestExtinction>, the Overview, Design concepts, Details (ODD), protocol for describing individual-based models is included within the model source code [@Grimm2010].
+We performed the simulations and analysis using R version 4.3.1 [@RCoreTeam2023], the R source code to perform simulations and analysis is located at <https://github.com/EcoComplex/multiSLOSS>.
 
@@
 
+## Author Contributions
+
+Leonardo Saravia, Ulises Balsa, and Fernando Momo conceived the project.
+Leonardo Saravia designed, developed, and performed the simulations.
+He analyzed the data in close consultation with all co-authors.
+Leonardo Saravia and Ulises Balsa wrote the first draft of the manuscript.
+All authors contributed to the manuscript’s revision and finalization.
+
 ## Acknowledgments
 
+We acknowledge the financial support provided by the "Fondo para la Investigación Científica y Tecnológica FONCYT" (Grant No.
+PICT-2020-SERIEA-02628).
+Additionally, we extend our gratitude to the National University of General Sarmiento (UNGS), especially to Dr.
+Santiago Doyle, for granting access to the computational resources utilized in this study.
+
+## Conflict Of Interest Statement
+
+We declare no conflict of interest.
+
+## Data Availability Statement
+
+The NetLogo model is available for download on GitHub: https://github.com/lsaravia/DynamicForestExtinction.
+Data and R scripts can also be accessed on GitHub: https://github.com/EcoComplex/multiSLOSS and @Saravia2024.
+
 ## References

### 2024-10-09 ms_SLOSS_title.Rmd
Updated title and Readme with zenodo

@@
 
+## Author Contributions
+
+Leonardo Saravia, Ulises Balsa, and Fernando Momo conceived the project.
+Leonardo Saravia designed, developed, and performed the simulations.
+He analyzed the data in close consultation with all co-authors.
+Leonardo Saravia and Ulises Balsa wrote the first draft of the manuscript.
+All authors contributed to the manuscript’s revision and finalization.
+
 ## Acknowledgments
@@
 PICT-2020-SERIEA-02628).
-Additionally, we extend our gratitude to the National University of General Sarmiento (UNGS), especially Santiago Doyle, for granting access to the computational resources utilized in this study.
+Additionally, we extend our gratitude to the National University of General Sarmiento (UNGS), especially to Dr.
+Santiago Doyle, for granting access to the computational resources utilized in this study.
+
+## Conflict Of Interest Statement
+
+We declare no conflict of interest.
+
+## Data Availability Statement
+
+The NetLogo model is available for download on GitHub: https://github.com/lsaravia/DynamicForestExtinction and Zenodo https://doi.org/10.5281/zenodo.8157152.
+Data and R scripts can also be accessed on GitHub: https://github.com/EcoComplex/multiSLOSS and Zenodo https://doi.org/10.5281/zenodo.13891484.

### 2024-10-25 ms_SLOSS_title.Rmd
Update PLS

@@
 csl: "ecology-letters.csl"
-
 output:
+  word_document: default
   pdf_document:

### 2024-11-11 ms_SLOSS.Rmd
Minor modifications add SI captions

@@
 
-![Relationship between the number of species and habitat patch size across panels representing different model versions and varying dispersal distances.
+![Relationship between the number of species and habitat patch size across panels representing different model versions and varying dispersal distances under regular habitat destruction.
 The vertical panels correspond to different percentages of habitat destruction.
@@
 
-![Relationship between the number of species and habitat patch size across panels representing different model versions and varying dispersal distances.
+![Relationship between the number of species and habitat patch size across panels representing different model versions and varying dispersal distances ``````Ω¸under random block habitat destruction.
 The vertical panels correspond to different percentages of habitat destruction.
